### 2022-08-20 data/example.tex
Add example.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{mdframed}
\usepackage{minted}

\newenvironment{lcode}
  {\VerbatimEnvironment
   \begin{mdframed}
   \begin{minted}[breaklines, fontsize=\footnotesize]{python}}
  {\end{minted}
   \end{mdframed}}
\newenvironment{lresult}
  {\begin{mdframed}}
  {\end{mdframed}}

\graphicspath{ {./} }

\begin{document}

\section{Environment tags}

LitREPL for latex recognizes specifically named environments as code and result
sections. It doesn't really evaluate Tex commands so renaming those environments
wouldn't work. But we still need to introduce it to Latex so we start with a
newenvironment declarations.

\begin{verbatim}
\newenvironment{lcode}
  {\VerbatimEnvironment
   \begin{mdframed}
   \begin{minted}[breaklines, fontsize=\footnotesize]{python}}
  {\end{minted}
   \end{mdframed}}
\newenvironment{lresult}
  {\begin{mdframed}[leftline=false,rightline=false]}
  {\end{mdframed}}
\end{verbatim}

\section{Executing code snippets}

Executable code snippet is the content of the \texttt{lcode} environment.
Putting the cursor on it and typing the \texttt{:LitEval1} command runs it
it in a background Python interpreter.

\begin{lcode}
W='Hello, world!'
print(W)
\end{lcode}

\texttt{lresult} section next to the executable section is a result container.
LitREPL would paste here the output of above code snippet. The original content
of the section will be replaced.

\begin{lresult}
Hello, world!
\end{lresult}

Commented \texttt{lresult} environmet tags are still recognized as a section for
verbatim results. This way users can generate parts of the latex document.

\begin{lcode}
print('Hello, LitREPL.vim!')
\end{lcode}

\begin{mdframed}
%\begin{lresult}
Hello LitREPL.vim
%\end{lresult}
\end{mdframed}

\section{Invoking Mathplotlib code}

\begin{lcode}
import numpy as np
import matplotlib.pyplot as plt
X=np.array(range(10))
_=plt.plot(X, np.sin(X))
def texplot(f):
  plt.savefig(f)
  print("\\includegraphics{./"+f+"}")

texplot('_img.png')
\end{lcode}

\newcommand{\linplace}[2]{#2}
%\begin{lresult}
\includegraphics{\linplace{VAR}{_img.png}}
%\end{lresult}

\end{document}

### 2022-08-20 python/litrepl.py
Adjust filetype names

@@ -234,12 +234,7 @@ def parse_(grammar):
   # print(tree.pretty())
   return tree
 
-GRAMMARS={'markdown':grammar_md,'latex':grammar_latex}
-
-def parse_md():
-  return parse_(grammar_md)
-def parse_latex():
-  return parse_(grammar_latex)
+GRAMMARS={'markdown':grammar_md,'tex':grammar_latex,'latex':grammar_latex}
 
 def print_(tree,symbols):
   class C(Interpreter):
@@ -259,9 +254,7 @@ def print_(tree,symbols):
       print(f"{symbols.verbeginmarker}{tree.children[1].children[0].value}{symbols.verendmarker}", end='')
   C().visit(tree)
 
-print_md=partial(print_,symbols=symbols_md)
-print_latex=partial(print_,symbols=symbols_latex)
-SYMBOLS={'markdown':symbols_md,'latex':symbols_latex}
+SYMBOLS={'markdown':symbols_md,'tex':symbols_latex,'latex':symbols_latex}
 
 def cursor_within(pos, posA, posB)->bool:
   if pos[0]>posA[0] and pos[0]<posB[0]:

### 2022-08-20 python/litrepl.py
Support linline tags for Latex

@@ -144,7 +144,7 @@ def start(a):
     else:
       start_(partial(fork_python,name='python'))
   else:
-    raise ValueError(f"Unsupported interpreter {a.interpreter}")
+    raise ValueError(f"Unsupported interpreter '{a.interpreter}'")
 
 def running()->bool:
   """ Checks if the background session was run or not. """
@@ -198,6 +198,7 @@ class SymbolsLatex:
   ocodendmarker=r"\end{lresult}"
   verbeginmarker=r"%\begin{lresult}"
   verendmarker=r"%\end{lresult}"
+  inlinemarker=r"\linline"
 
 symbols_latex=SymbolsLatex()
 icodebeginmarkerE=r"\\begin\{lcode\}"
@@ -206,25 +207,34 @@ ocodebeginmarkerE=r"\\begin\{lresult\}"
 ocodendmarkerE=r"\\end\{lresult\}"
 verbeginmarkerE=r"\%\\begin\{lresult\}"
 verendmarkerE=r"\%\\end\{lresult\}"
+inlinemarkerE=r"\\linline\{"
 
+OBR="{"
+CBR="}"
+BCBR="\\}"
 grammar_latex = fr"""
 start: (snippet)*
 snippet : icodesection -> e_icodesection
         | ocodesection -> e_ocodesection
         | oversection -> e_versection
         | topleveltext -> e_text
-        // | inlinesection -> e_inline
+        | inlinesection -> e_inline
 icodesection.1 : icodebeginmarker innertext icodendmarker
 ocodesection.1 : ocodebeginmarker innertext ocodendmarker
 oversection.1 : verbeginmarker innertext verendmarker
+inlinesection.1 : inlinemarker "{OBR}" inltext "{CBR}" spaces_obr inltext cbr
+inlinemarker : "{symbols_latex.inlinemarker}"
 icodebeginmarker : "{symbols_latex.icodebeginmarker}"
 icodendmarker : "{symbols_latex.icodendmarker}"
 ocodebeginmarker : "{symbols_latex.ocodebeginmarker}"
 ocodendmarker : "{symbols_latex.ocodendmarker}"
 verbeginmarker : "{symbols_latex.verbeginmarker}"
 verendmarker : "{symbols_latex.verendmarker}"
-topleveltext : /(.(?!{icodebeginmarkerE}|{ocodebeginmarkerE}|{verbeginmarkerE}))*./s
[…]

### 2022-08-20 sh/test.sh

@@ -1,11 +1,9 @@
 #!/bin/sh
 
-CWD=$(cd `dirname $0`/..; pwd;)
-
-set -e -x
-
 mktest() {
-  T="$CWD/_test/$1"
+  set -e -x
+  test -d "$LITREPL_ROOT"
+  T="$LITREPL_ROOT/_test/$1"
   rm -rf  "$T" || true
   mkdir -p "$T"
   cd "$T"
@@ -103,7 +101,7 @@ EOF
 )} #}}}
 
 
-test_eval_md_1() {
+test_eval_md_1() {(
 mktest "_test_eval_md_1"
 runlitrepl start
 runlitrepl eval-section --line 1 --col 1 >out.md <<"EOF"
@@ -120,9 +118,9 @@ EOF
 grep -q "Hello, World!" out.md
 grep -v -q "PLACEHOLDER" out.md
 runlitrepl stop
-}
+)}
 
-test_eval_md_2() {
+test_eval_md_2() {(
 mktest "_test_eval_md_2"
 runlitrepl start
 runlitrepl eval-section --line 1 --col 1 >out.md <<"EOF"
@@ -136,9 +134,9 @@ EOF
 grep -q "ABCDEF" out.md
 grep -v -q "PLACEHOLDER" out.md
 runlitrepl stop
-}
+)}
 
-test_eval_tex_1() {
+test_eval_tex_1() {(
 mktest "_test_eval_tex_1"
 runlitrepl start
 cat >source.tex <<"EOF"
@@ -154,17 +152,51 @@ cat source.tex | runlitrepl --filetype=latex eval-section --line 1 --col 1 >out.
 grep -q "AB" out.tex
 grep -v -q "PLACEHOLDER" out.tex
 runlitrepl stop
-}
+)}
 
-trap "echo FAIL" EXIT
-for I in python ipython ; do
-  echo "Checking $I"
-  LITREPL_INTERPRETER=$I
-  test_parse_print
-  test_eval_md_1
-  test_eval_md_2
-  test_eval_tex_1
-done
-trap "" EXIT
-echo OK
+test_parse_print_tex_inline_1() {(
+mktest "_test_parse_print_tex_inline_1"
+# runlitrepl start
+cat >source.tex <<"EOF"
+111
+\linline{"A"+"B"}
+{XX}
+222
+EOF
+cat source.tex | runlitrepl --filetype=latex parse-print >out.tex
+diff -u source.tex out.tex
+# runlitrepl stop
+)}
+
[…]

### 2022-08-20 python/litrepl.py
Fix an empty inline bug

@@ -232,7 +232,7 @@ verbeginmarker : "{symbols_latex.verbeginmarker}"
 verendmarker : "{symbols_latex.verendmarker}"
 topleveltext : /(.(?!{icodebeginmarkerE}|{ocodebeginmarkerE}|{verbeginmarkerE}|{inlinemarkerE}))*./s
 innertext : /(.(?!{icodendmarkerE}|{ocodendmarkerE}|{verendmarkerE}))*./s
-inltext : /(.(?!{CBR}))*./s
+inltext : ( /(.(?!{CBR}))*./s )?
 spaces_obr : /[ \t\r\n]*{OBR}/s
 cbr : "{CBR}"
 """

### 2022-08-20 sh/test.sh

@@ -183,6 +183,19 @@ grep -v -q "XX" out.tex
 runlitrepl stop
 )}
 
+test_eval_tex_inline_2() {(
+mktest "_test_eval_tex_inline_2"
+runlitrepl start
+cat >source.tex <<"EOF"
+111
+\linline{"A"+"B"}{}
+222
+EOF
+cat source.tex | runlitrepl --filetype=latex eval-section --line 2 --col 1 >out.tex
+grep -q "AB" out.tex
+runlitrepl stop
+)}
+
 if test "$(basename $0)" = "test.sh" ; then
   set -e -x
   trap "echo FAIL" EXIT
@@ -195,6 +208,7 @@ if test "$(basename $0)" = "test.sh" ; then
     test_eval_tex_1
     test_parse_print_tex_inline_1
     test_eval_tex_inline_1
+    test_eval_tex_inline_2
   done
   trap "" EXIT
   echo OK

### 2022-08-20 data/example.tex
Support linline tags for Latex
Update readme

@@
   {\end{mdframed}}
+\newcommand{\linline}[2]{#2}
 
@@
 
-LitREPL for latex recognizes specifically named environments as code and result
-sections. It doesn't really evaluate Tex commands so renaming those environments
-wouldn't work. But we still need to introduce it to Latex so we start with a
-newenvironment declarations.
+LitREPL treats specially named Latex environments as code and result sections.
+It doesn't really evaluate \TeX\ commands so the tags should present in the source
+file explicitly rather than be a result of other commands such as
+\texttt{newcommand} or \texttt{newenvironment}.
 
@@
   {\end{mdframed}}
+\newcommand{\linplace}[2]{#2}
 \end{verbatim}
@@
 
-Executable code snippet is the content of the \texttt{lcode} environment.
-Putting the cursor on it and typing the \texttt{:LitEval1} command runs it
-it in a background Python interpreter.
+Executable code is obtained from the content of the \texttt{lcode} environment.
+Putting the cursor on this environment and typing the \texttt{:LitEval1} Vim
+command sends the code to the Python interpreter persistently running in the
+background.
+
+\begin{verbatim}
+\begin{lcode}
+W='Hello, world!'
+print(W)
+\end{lcode}
+\end{verbatim}
 
@@
 \texttt{lresult} section next to the executable section is a result container.
-LitREPL would paste here the output of above code snippet. The original content
-of the section will be replaced.
+When executed, LitREPL replaces its content with the executable code output.
+
+\begin{verbatim}
+\begin{lresult}
+??
+\end{lresult}
+\end{verbatim}
 
@@
 
-Commented \texttt{lresult} environmet tags are still recognized as a section for
-verbatim results. This way users can generate parts of the latex document.
+VimREPL also recognises commented \texttt{lresult} tags as a valid result
+container. This we can generate custom parts of the document.
 
@@
 
+\begin{verbatim}
+%\begin{lresult}
+??
+%\end{lresult}
+\end{verbatim}
+
 \begin{mdframed}
 %\begin{lresult}
-Hello LitREPL.vim
+Hello, LitREPL.vim!
 %\end{lresult}
@@
 
-\section{Invoking Mathplotlib code}
+A special two-argument \texttt{linline} tag works both as an executable section
+and as a result container. We expect it to be used to print the contents of
+Python variables, so the VimREPL wraps the executable code with the additional
+\texttt{print(...)} operator. On the Latex side, we define the tag to ignore its
+first argument but pass the second one as-is.
+
+\begin{verbatim}
+\begin{mdframed}\linline{W}{??}\end{mdframed}
+\end{verbatim}
+
+\begin{mdframed}\linline{W}{Hello, world!}\end{mdframed}
+
+\section{Mathplotlib plots and other graphics}
 
@@
 X=np.array(range(10))
-_=plt.plot(X, np.sin(X))
-def texplot(f):
-  plt.savefig(f)
-  print("\\includegraphics{./"+f+"}")
-
-texplot('_img.png')
+plt.plot(X, np.sin(X))
+path='_img.png'
+plt.savefig(path)
 \end{lcode}
 
-\newcommand{\linplace}[2]{#2}
-%\begin{lresult}
-\includegraphics{\linplace{VAR}{_img.png}}
-%\end{lresult}
+\includegraphics{\linline{path}{_img.png}}
 

### 2022-08-20 python/litrepl.py
Add (eval-sections STR) command

@@ -2,7 +2,7 @@
 
 # from pylightnix import *
 
-from typing import List, Optional, Tuple
+from typing import List, Optional, Tuple, Set, Dict
 from re import search, match as re_match
 import re
 import os
@@ -245,29 +245,6 @@ def parse_(grammar):
   return tree
 
 GRAMMARS={'markdown':grammar_md,'tex':grammar_latex,'latex':grammar_latex}
-
-def print_(tree,symbols):
-  class C(Interpreter):
-    def text(self,tree):
-      print(tree.children[0].value, end='')
-    def topleveltext(self,tree):
-      return self.text(tree)
-    def innertext(self,tree):
-      return self.text(tree)
-    def icodesection(self,tree):
-      print(f"{symbols.icodebeginmarker}{tree.children[1].children[0].value}{symbols.icodendmarker}", end='')
-    def ocodesection(self,tree):
-      print(f"{symbols.ocodebeginmarker}{tree.children[1].children[0].value}{symbols.ocodendmarker}", end='')
-    def oversection(self,tree):
-      print(f"{symbols.verbeginmarker}{tree.children[1].children[0].value}{symbols.verendmarker}", end='')
-    def inlinesection(self,tree):
-      code=tree.children[1].children[0].value
-      result=tree.children[3].children[0].value
-      spaces_obs=tree.children[2].children[0].value
-      im=symbols.inlinemarker
-      print(f"{im}{OBR}{code}{CBR}{spaces_obs}{result}{CBR}", end='')
-  C().visit(tree)
-
 SYMBOLS={'markdown':symbols_md,'tex':symbols_latex,'latex':symbols_latex}
 
 def cursor_within(pos, posA, posB)->bool:
@@ -289,17 +266,15 @@ def unindent(col:int,lines:str)->str:
 def indent(col,lines:str)->str:
   return '\n'.join([' '*col+l for l in lines.split('\n')])
 
-def eval_section_(a, tree, symbols, cpos:Optional[Tuple[int,int]]=None,
-                  nsec:Optional[int]=None):
-  # TODO: nsec
-  assert cpos is not None
-  line,col=cpos if cpos is not None else (None,None)
+def eval_section_(a, tree, symbols, nsecs:Set[int]):
   if not running():
     start(a)
+  ssrc:Dict[int,str]={}
+  sres:Dict[int,str]={}
   class C(Interpreter):
     def __init__(self):
-      self.task=None
-      self.result=None
+      self.nsec=-1
[…]

### 2022-08-20 sh/test.sh

@@ -101,98 +101,93 @@ EOF
 )} #}}}
 
 
-test_eval_md_1() {(
-mktest "_test_eval_md_1"
+test_eval_md() {(
+mktest "_test_eval_md"
 runlitrepl start
-runlitrepl eval-section --line 1 --col 1 >out.md <<"EOF"
+cat >source.md <<"EOF"
 ```python
 def hello(name):
   print(f"Hello, {name}!")
 
 hello('World')
 ```
+TXT
 ```
 PLACEHOLDER
 ```
+TXT
+* ```python
+  print("ABC"+"DEF")
+  ```
+* <!--litrepl-->
+  PLACEHOLDER
+  <!--litrepl-->
 EOF
-grep -q "Hello, World!" out.md
-grep -v -q "PLACEHOLDER" out.md
-runlitrepl stop
-)}
-
-test_eval_md_2() {(
-mktest "_test_eval_md_2"
-runlitrepl start
-runlitrepl eval-section --line 1 --col 1 >out.md <<"EOF"
+cat source.md | runlitrepl --filetype=markdown parse-print >out.md
+diff -u source.md out.md
+cat source.md | runlitrepl --filetype=markdown eval-sections '0..$' >out.md
+diff -u out.md - <<"EOF"
 ```python
-print("ABC"+"DEF")
+def hello(name):
+  print(f"Hello, {name}!")
+
+hello('World')
 ```
-<!--litrepl-->
-PLACEHOLDER
-<!--litrepl-->
+TXT
+```
+Hello, World!
+```
+TXT
+* ```python
+  print("ABC"+"DEF")
+  ```
+* <!--litrepl-->
+  ABCDEF
+  <!--litrepl-->
 EOF
-grep -q "ABCDEF" out.md
-grep -v -q "PLACEHOLDER" out.md
 runlitrepl stop
 )}
 
-test_eval_tex_1() {(
-mktest "_test_eval_tex_1"
+test_eval_tex() {(
+mktest "_test_eval_tex"
 runlitrepl start
 cat >source.tex <<"EOF"
 \begin{lcode}
 print("A"+"B")
 \end{lcode}
 Text
-%\begin{lresult}
-PLACEHOLDER
-%\end{lresult}
-EOF
-cat source.tex | runlitrepl --filetype=latex eval-section --line 1 --col 1 >out.tex
-grep -q "AB" out.tex
-grep -v -q "PLACEHOLDER" out.tex
-runlitrepl stop
-)}
-
[…]

### 2022-08-20 python/litrepl.py
Forbid consequitive text sections in parsers

@@ -164,16 +164,20 @@ class SymbolsMarkdown:
   ocodendmarker="```"
   verbeginmarker="<!--litrepl-->"
   verendmarker="<!--litrepl-->"
+  combeginmarker=r"<!--lignore-->"
+  comendmarker=r"<!--lnoignore-->"
 
 symbols_md=SymbolsMarkdown()
 
 grammar_md = fr"""
-start: (snippet)*
+start: (text)? (snippet (text)?)*
 snippet : icodesection -> e_icodesection
         | ocodesection -> e_ocodesection
         | oversection -> e_versection
-        | text -> e_text
-//        | inlinesection -> e_inline
+        | comsection -> e_comsection
+        // | text -> e_text
+        // | inlinesection -> e_inline
+comsection.2 : combeginmarker comtext comendmarker
 icodesection.1 : icodebeginmarker text icodendmarker
 ocodesection.1 : ocodebeginmarker text ocodendmarker
 oversection.1 : verbeginmarker text verendmarker
@@ -186,7 +190,10 @@ verbeginmarker : "<!--litrepl-->"
 verendmarker : "<!--litrepl-->"
 inlinebeginmarker : "`"
 inlinendmarker : "`"
-text : /(.(?!```|<\!--litrepl-->))*./s
+combeginmarker : "{symbols_md.combeginmarker}"
+comendmarker : "{symbols_md.comendmarker}"
+text : /(.(?!```|<\!--litrepl-->|{symbols_md.combeginmarker}|{symbols_md.comendmarker}))*./s
+comtext : /(.(?!{symbols_md.comendmarker}))*./s
 """
 
 
@@ -199,6 +206,8 @@ class SymbolsLatex:
   verbeginmarker=r"%\begin{lresult}"
   verendmarker=r"%\end{lresult}"
   inlinemarker=r"\linline"
+  combeginmarker=r"%lignore"
+  comendmarker=r"%lnoignore"
 
 symbols_latex=SymbolsLatex()
 icodebeginmarkerE=r"\\begin\{lcode\}"
@@ -208,17 +217,25 @@ ocodendmarkerE=r"\\end\{lresult\}"
 verbeginmarkerE=r"\%\\begin\{lresult\}"
 verendmarkerE=r"\%\\end\{lresult\}"
 inlinemarkerE=r"\\linline\{"
+combeginmarkerE=r"\%lignore"
+comendmarkerE=r"\%lnoignore"
 
+allmarkersE='|'.join([icodebeginmarkerE,icodendmarkerE,
[…]

### 2022-08-20 sh/test.sh

@@ -122,6 +122,18 @@ TXT
 * <!--litrepl-->
   PLACEHOLDER
   <!--litrepl-->
+== Ignore ==
+<!--lignore-->
+```python
+print('A'+'B')
+```
+```
+PLACEHOLDER
+```
+<!--lnoignore-->
+```
+NOEVAL
+```
 EOF
 cat source.md | runlitrepl --filetype=markdown parse-print >out.md
 diff -u source.md out.md
@@ -144,6 +156,18 @@ TXT
 * <!--litrepl-->
   ABCDEF
   <!--litrepl-->
+== Ignore ==
+<!--lignore-->
+```python
+print('A'+'B')
+```
+```
+PLACEHOLDER
+```
+<!--lnoignore-->
+```
+NOEVAL
+```
 EOF
 runlitrepl stop
 )}
@@ -165,8 +189,21 @@ Text
 \begin{lresult}
 NOEVAL
 \end{lresult}
+== Inline ==
 \linline{"A"+"B"}
 {XX}\linline{"C"+"D"}{}
+== Ignore ==
+%lignore
+\begin{lcode}
+print("X"+"Y")
+\end{lcode}
+%\begin{lresult}
+ZZZZZ
+%\end{lresult}
+%lnoignore
+\begin{lresult}
+PLACEHOLDER
+\end{lresult}
 EOF
 cat source.tex | runlitrepl --filetype=latex parse-print >out.tex
 diff -u source.tex out.tex
@@ -185,11 +222,53 @@ Text
 \begin{lresult}
 NOEVAL
 \end{lresult}
+== Inline ==
 \linline{"A"+"B"}
 {AB}\linline{"C"+"D"}{CD}
+== Ignore ==
+%lignore
+\begin{lcode}
+print("X"+"Y")
+\end{lcode}
+%\begin{lresult}
+ZZZZZ
+%\end{lresult}
+%lnoignore
+\begin{lresult}
+PLACEHOLDER
+\end{lresult}
 EOF
 runlitrepl stop
 )}
+
+# test_eval_ignore() {(
+# mktest "_test_eval_ignore"
+# runlitrepl start
+# cat >source.tex <<"EOF"
+# %lignore
+# \begin{lcode}
+# print("X"+"W")
+# \end{lcode}
+# %\begin{lresult}
+# ZZZZZ
+# %\end{lresult}
+# %lnoignore
+# EOF
+# # cat source.tex | runlitrepl --filetype=latex parse #>out.tex
[…]

### 2022-08-21 python/litrepl.py
Update marker names, allow 1-nested parens in Latex inline

@@ -173,15 +173,11 @@ grammar_md = fr"""
 start: (text)? (snippet (text)?)*
 snippet : icodesection -> e_icodesection
         | ocodesection -> e_ocodesection
-        | oversection -> e_versection
         | comsection -> e_comsection
-        // | text -> e_text
-        // | inlinesection -> e_inline
 comsection.2 : combeginmarker comtext comendmarker
 icodesection.1 : icodebeginmarker text icodendmarker
 ocodesection.1 : ocodebeginmarker text ocodendmarker
-oversection.1 : verbeginmarker text verendmarker
-// inlinesection : inlinebeginmarker text inlinendmarker
+               | verbeginmarker text verendmarker
 icodebeginmarker : "{symbols_md.icodebeginmarker}"
 icodendmarker : "{symbols_md.icodendmarker}"
 ocodebeginmarker : "{symbols_md.ocodebeginmarker}"
@@ -201,10 +197,12 @@ comtext : /(.(?!{symbols_md.comendmarker}))*./s
 class SymbolsLatex:
   icodebeginmarker=r"\begin{lcode}"
   icodendmarker=r"\end{lcode}"
+  icodebeginmarker2=r"%lcode"
+  icodendmarker2=r"%lnocode"
   ocodebeginmarker=r"\begin{lresult}"
   ocodendmarker=r"\end{lresult}"
-  verbeginmarker=r"%\begin{lresult}"
-  verendmarker=r"%\end{lresult}"
+  verbeginmarker=r"%lresult"
+  verendmarker=r"%lnoresult"
   inlinemarker=r"\linline"
   combeginmarker=r"%lignore"
   comendmarker=r"%lnoignore"
@@ -212,15 +210,18 @@ class SymbolsLatex:
 symbols_latex=SymbolsLatex()
 icodebeginmarkerE=r"\\begin\{lcode\}"
 icodendmarkerE=r"\\end\{lcode\}"
+icodebeginmarker2E=r"\%lcode"
+icodendmarker2E=r"\%lnocode"
 ocodebeginmarkerE=r"\\begin\{lresult\}"
 ocodendmarkerE=r"\\end\{lresult\}"
-verbeginmarkerE=r"\%\\begin\{lresult\}"
-verendmarkerE=r"\%\\end\{lresult\}"
+verbeginmarkerE=r"\%lresult"
+verendmarkerE=r"\%lnoresult"
[…]

### 2022-08-21 sh/test.sh

@@ -91,9 +91,9 @@ Bartext
   \end{lcode}
 
   \item
-  %\begin{lresult}
+  %lresult
   hello("Verbose")
-  %\end{lresult}
+  %lnoresult
 \end{itemize}
 \end{document}
 EOF
@@ -166,7 +166,7 @@ PLACEHOLDER
 ```
 <!--lnoignore-->
 ```
-NOEVAL
+ABCDEF
 ```
 EOF
 runlitrepl stop
@@ -182,9 +182,9 @@ print("A"+"B")
 Text
 \begin{enumerate}
 \item
-  %\begin{lresult}
+  %lresult
   PLACEHOLDER
-  %\end{lresult}
+  %lnoresult
 \end{enumerate}
 \begin{lresult}
 NOEVAL
@@ -192,14 +192,18 @@ NOEVAL
 == Inline ==
 \linline{"A"+"B"}
 {XX}\linline{"C"+"D"}{}
+\begin{lcode}
+tag='\\textbf{bold}'
+\end{lcode}
+\linline{tag+tag}{\textbf{old}\textbf{old}}
 == Ignore ==
 %lignore
 \begin{lcode}
 print("X"+"Y")
 \end{lcode}
-%\begin{lresult}
+%lresult
 ZZZZZ
-%\end{lresult}
+%lnoresult
 %lnoignore
 \begin{lresult}
 PLACEHOLDER
@@ -215,60 +219,63 @@ print("A"+"B")
 Text
 \begin{enumerate}
 \item
-  %\begin{lresult}
+  %lresult
   AB
-  %\end{lresult}
+  %lnoresult
 \end{enumerate}
 \begin{lresult}
-NOEVAL
+AB
 \end{lresult}
 == Inline ==
 \linline{"A"+"B"}
 {AB}\linline{"C"+"D"}{CD}
+\begin{lcode}
+tag='\\textbf{bold}'
+\end{lcode}
+\linline{tag+tag}{\textbf{bold}\textbf{bold}}
 == Ignore ==
 %lignore
 \begin{lcode}
 print("X"+"Y")
 \end{lcode}
-%\begin{lresult}
+%lresult
 ZZZZZ
-%\end{lresult}
+%lnoresult
 %lnoignore
 \begin{lresult}
-PLACEHOLDER
 \end{lresult}
 EOF
 runlitrepl stop
 )}
 
-# test_eval_ignore() {(
-# mktest "_test_eval_ignore"
[…]

### 2022-08-21 python/litrepl.py
Fix a regression in solve_cpos

@@ -291,7 +291,7 @@ def unindent(col:int,lines:str)->str:
 def indent(col,lines:str)->str:
   return '\n'.join([' '*col+l for l in lines.split('\n')])
 
-def eval_section_(a, tree, symbols, nsecs:Set[int]):
+def eval_section_(a, tree, symbols, nsecs:Set[int])->None:
   if not running():
     start(a)
   ssrc:Dict[int,str]={}
@@ -299,7 +299,6 @@ def eval_section_(a, tree, symbols, nsecs:Set[int]):
   class C(Interpreter):
     def __init__(self):
       self.nsec=-1
-      # self.nosec=0
     def text(self,tree):
       print(tree.children[0].value, end='')
     def topleveltext(self,tree):
@@ -308,7 +307,6 @@ def eval_section_(a, tree, symbols, nsecs:Set[int]):
       return self.text(tree)
     def icodesection(self,tree):
       self.nsec+=1
-      # self.nosec=0
       t=tree.children[1].children[0].value
       bmarker=getattr(symbols,tree.children[0].data)
       emarker=getattr(symbols,tree.children[2].data)
@@ -326,7 +324,6 @@ def eval_section_(a, tree, symbols, nsecs:Set[int]):
         assert self.nsec in sres
         print(bmarker+"\n"+indent(bm.column-1,(f"{sres[self.nsec]}"+emarker)),
               end='')
-        # self.nosec+=1
       else:
         print(f"{bmarker}{tree.children[1].children[0].value}{emarker}", end='')
     def ocodesection(self,tree):
@@ -368,8 +365,7 @@ def solve_cpos(tree,cs:List[Tuple[int,int]]
     def oversection(self,tree):
       self._count(tree.children[0].meta,tree.children[2].meta)
     def inlinesection(self,tree):
-      self.nsec+=1
-      self._count(tree.children[0].meta,tree.children[4].meta)
+      self._count(tree.children[0].meta,tree.children[5].meta)
   c=C()
   c.visit(tree)
   return c.nsec,acc

### 2022-08-21 python/litrepl.py
Add setup.py and Nix build rule for litrepl

@@ -430,6 +430,7 @@ def solve_sloc(s:str,tree)->Set[int]:
 
 if __name__=='__main__':
   ap=ArgumentParser(prog='litrepl.py')
+  ap.add_argument('--version',action='version',version='999.999',help='print version')
   ap.add_argument('--filetype',metavar='STR',default='markdown',help='ft help')
   ap.add_argument('--interpreter',metavar='EXE',default='auto',help='python|ipython|auto')
   sps=ap.add_subparsers(help='command help', dest='command')

### 2022-08-21 python/setup.py

@@ -1,0 +1,13 @@
+from setuptools import setup, find_packages
+from distutils.spawn import find_executable
+
+setup(
+  name="litrepl",
+  zip_safe=False, # https://mypy.readthedocs.io/en/latest/installed_packages.html
+  package_dir={'':'.'},
+  packages=find_packages(where='.'),
+  install_requires=['lark'],
+  scripts=['litrepl.py'],
+  python_requires='>=3.6',
+)
+

### 2022-08-21 sh/test.sh

@@ -11,7 +11,7 @@ mktest() {
 
 runlitrepl() {
   test -n "$LITREPL_INTERPRETER"
-  litrepl.py --interpreter="$LITREPL_INTERPRETER" "$@"
+  $LITREPL --interpreter="$LITREPL_INTERPRETER" "$@"
 }
 
 test_parse_print() {( # {{{
@@ -277,8 +277,11 @@ EOF
 runlitrepl stop
 )}
 
-if test "$(basename $0)" = "test.sh" ; then
+if test -n '$LITREPL_TEST' || echo "$(basename $0)" | grep -q "test.sh" ; then
   set -e -x
+  if test -z "$LITREPL"; then
+    LITREPL=$LITREPL_ROOT/python/litrepl.py
+  fi
   trap "echo FAIL" EXIT
   for I in python ipython ; do
     echo "Checking $I"

### 2022-08-21 env.sh
Update buildsystem; Add Makefile and wheel

@@ -1,5 +1,8 @@
 CWD=$(pwd)
-export PATH=$CWD/sh:$CWD/python:$PATH
+export PATH=\
+$CWD/sh:\
+$CWD/python/bin:\
+$PATH
 export LITREPL_ROOT="$CWD"
 export PYLIGHTNIX_ROOT="$CWD/_pylightnix"
 export PYTHONPATH=\
@@ -17,5 +20,5 @@ unset VIMRUNTIME
 python3 -c 'import lark' || echo "Warning: 'lark' python library not found" >&2
 python3 -c 'import dataclasses' || echo "Warning: 'dataclasses' python library not found" >&2
 # python -c 'import pylightnix' || echo "Warning: 'lark' python library not found" >&2
-litrepl.py --version >/dev/null || echo "litrepl.py is not in PATH" >&2
+litrepl --version >/dev/null || echo "litrepl is not in PATH" >&2
 socat -V >/dev/null || echo "GNU socat not found" >&2

### 2022-08-21 python/litrepl/__init__.py

@@ -1,0 +1,17 @@
+from logging import getLogger
+logger=getLogger(__name__)
+info=logger.info
+warning=logger.warning
+
+from litrepl.base import *
+
+try:
+  from litrepl.version import __version__
+except ImportError:
+  try:
+    from setuptools_scm import get_version
+    from os.path import join
+    __version__ = get_version(root=join('..','..'), relative_to=__file__)
+  except Exception:
+    warning("LitREPL failed to read its version.")
+    __version__ = None

### 2022-08-21 python/litrepl/base.py

@@ -1,13 +1,10 @@
-#!/usr/bin/env python3
-
-# from pylightnix import *
-
-from typing import List, Optional, Tuple, Set, Dict
-from re import search, match as re_match
 import re
 import os
 import sys
 import fcntl
+
+from typing import List, Optional, Tuple, Set, Dict, Callable
+from re import search, match as re_match
 from select import select
 from os import environ, system
 from lark import Lark, Visitor, Transformer, Token, Tree
@@ -90,14 +87,14 @@ def fork_ipython(name):
   open('_config.py','w').write(
     'from IPython.terminal.prompts import Prompts, Token\n'
     'class EmptyPrompts(Prompts):\n'
-    '    def in_prompt_tokens(self):\n'
-    '        return [(Token.Prompt, "")]\n'
-    '    def continuation_prompt_tokens(self, width=None):\n'
-    '        return [(Token.Prompt, "") ]\n'
-    '    def rewrite_prompt_tokens(self):\n'
-    '        return []\n'
-    '    def out_prompt_tokens(self):\n'
-    '        return []\n'
+    '  def in_prompt_tokens(self):\n'
+    '    return [(Token.Prompt, "")]\n'
+    '  def continuation_prompt_tokens(self, width=None):\n'
+    '    return [(Token.Prompt, "") ]\n'
+    '  def rewrite_prompt_tokens(self):\n'
+    '    return []\n'
+    '  def out_prompt_tokens(self):\n'
+    '    return []\n'
     'c.TerminalInteractiveShell.prompts_class = EmptyPrompts\n'
     'c.TerminalInteractiveShell.separate_in = ""\n'
     'c.TerminalInteractiveShell.separate_out = ""\n'
@@ -115,7 +112,7 @@ def fork_ipython(name):
   )
   exit(0)
 
-def start_(fork_handler:callable):
+def start_(fork_handler:Callable[...,None])->None:
   """ Starts the background Python interpreter. Kill an existing interpreter if
   any. Creates files `_inp.pipe`, `_out.pipt`, `_pid.txt`."""
   if isfile('_pid.txt'):
@@ -427,46 +424,4 @@ def solve_sloc(s:str,tree)->Set[int]:
       return set()
   return set.union(*[_safeset(lambda:range(_get(q[0]),_get(q[1])+1)) if len(q)==2
                      else _safeset(lambda:[_get(q[0])]) for q in qs])
[…]

### 2022-08-21 python/litrepl/version.py

@@ -1,0 +1,2 @@
+# AUTOGENERATED!
+__version__ = '0.1.dev42+ge4ff4bb.d19800101'

### 2022-08-21 python/setup.py

@@ -1,13 +1,0 @@
-from setuptools import setup, find_packages
-from distutils.spawn import find_executable
-
-setup(
-  name="litrepl",
-  zip_safe=False, # https://mypy.readthedocs.io/en/latest/installed_packages.html
-  package_dir={'':'.'},
-  packages=find_packages(where='.'),
-  install_requires=['lark'],
-  scripts=['litrepl.py'],
-  python_requires='>=3.6',
-)
-

### 2022-08-21 setup.py

@@ -1,0 +1,43 @@
+from setuptools import setup, find_packages
+from distutils.spawn import find_executable
+from setuptools_scm import get_version
+from os.path import isfile, join
+from os import environ
+
+if environ.get("LITREPL_ROOT",None):
+  LITREPL_ROOT=environ["LITREPL_ROOT"]
+else:
+  LITREPL_ROOT='.'
+
+LITREPL_VERSION=get_version(root=LITREPL_ROOT)
+
+with open(join('python','litrepl','version.py'), 'w') as f:
+  f.write("# AUTOGENERATED!\n")
+  f.write(f"__version__ = '{LITREPL_VERSION}'\n")
+
+setup(
+  name="litrepl",
+  zip_safe=False, # https://mypy.readthedocs.io/en/latest/installed_packages.html
+  use_scm_version={
+    "root": LITREPL_ROOT,
+  },
+  package_dir={'':'python'},
+  packages=find_packages(where='python'),
+  install_requires=['lark'],
+  setup_requires=['setuptools_scm'],
+  scripts=['./python/bin/litrepl'],
+  python_requires='>=3.6',
+  author="Sergei Mironov",
+  author_email="grrwlf@gmail.com",
+  description="LitREPL a macroprocessing Python library for Litrate "\
+              "programming and code execution.",
+  classifiers=[
+    "Programming Language :: Python :: 3",
+    "License :: OSI Approved :: Apache Software License",
+    "Operating System :: POSIX :: Linux",
+    "Topic :: Software Development :: Build Tools",
+    "Intended Audience :: Developers",
+    "Development Status :: 3 - Alpha",
+  ],
+)
+

### 2022-08-21 sh/test.sh

@@ -280,7 +280,7 @@ runlitrepl stop
 if test -n '$LITREPL_TEST' || echo "$(basename $0)" | grep -q "test.sh" ; then
   set -e -x
   if test -z "$LITREPL"; then
-    LITREPL=$LITREPL_ROOT/python/litrepl.py
+    LITREPL=$LITREPL_ROOT/python/bin/litrepl
   fi
   trap "echo FAIL" EXIT
   for I in python ipython ; do

### 2022-08-21 python/litrepl/version.py
Remove version.py from git

@@ -1,2 +1,0 @@
-# AUTOGENERATED!
-__version__ = '0.1.dev42+ge4ff4bb.d19800101'

### 2022-08-21 setup.py
Disable local versions for Pypi

@@ -15,11 +15,15 @@ with open(join('python','litrepl','version.py'), 'w') as f:
   f.write("# AUTOGENERATED!\n")
   f.write(f"__version__ = '{LITREPL_VERSION}'\n")
 
+def local_scheme(version):
+  return ""
+
 setup(
   name="litrepl",
   zip_safe=False, # https://mypy.readthedocs.io/en/latest/installed_packages.html
   use_scm_version={
     "root": LITREPL_ROOT,
+    "local_scheme": local_scheme,
   },
   package_dir={'':'python'},
   packages=find_packages(where='python'),

### 2022-08-22 doc/example.tex
Updated docs

\documentclass{article}
\usepackage[utf8]{inputenc}
\begin{document}

LitREPL for latex recognizes \texttt{lcode} environments as code and
\texttt{lresult} as result sections. The tag names is currently hardcoded into
the simple parser the tool is using, so we need to additionally introduce it to
the Latex system. Here we do it in a most simple way.

\newenvironment{lcode}{\begin{texttt}}{\end{texttt}}
\newenvironment{lresult}{\begin{texttt}}{\end{texttt}}
\newcommand{\linline}[2]{#2}

Executable section is the text between the \texttt{lcode} begin/end tags.
Putting the cursor on it and typing the \texttt{:LitEval1} executes it in the
background Python interpreter.

\begin{lcode}
W='Hello, World!'
print(W)
\end{lcode}

\texttt{lresult} tags next to the executable section mark the result section.
The output of the executable section will be pasted here. The
original content of the section will be replaced.

\begin{lresult}
Hello, World!
\end{lresult}

Commented \texttt{lresult}/\texttt{lnoresult} tags also marks result sections.
This way we could customise the Latex markup for every particular section.

\begin{lcode}
print("Hi!")
\end{lcode}

%lresult
Hi!
%lnoresult

Additionally, VimREPL for Latex recognises \texttt{linline} tags for which it
prints its first argument and pastes the result in place of the second argument.

\linline{W}{Hello, World!}

\end{document}

### 2022-08-22 doc/quickstart.tex
Add support for commented markdown tags

--- 2022-08-20 data/example.tex
@@
 \usepackage{minted}
+\usepackage{verbatim}
 
@@
   {\VerbatimEnvironment
-   \begin{mdframed}
-   \begin{minted}[breaklines, fontsize=\footnotesize]{python}}
-  {\end{minted}
-   \end{mdframed}}
-\newenvironment{lresult}
-  {\begin{mdframed}}
-  {\end{mdframed}}
+   \begin{minted}[breaklines,fontsize=\footnotesize]{python}}
+  {\end{minted}}
+\BeforeBeginEnvironment{lcode}{\begin{mdframed}}
+\AfterEndEnvironment{lcode}{\end{mdframed}}
+
+% \newenvironment{lcode}{\verbatim}{\endverbatim}
+% \BeforeBeginEnvironment{lcode}{\begin{minted}[breaklines,fontsize=\footnotesize]{python} }
+% \AfterEndEnvironment{lcode}{\end{minted}}
+% \BeforeBeginEnvironment{lcode}{\begin{mdframed}}
+% \AfterEndEnvironment{lcode}{\end{mdframed}}
+
+\newenvironment{lresult}{\verbatim}{\endverbatim}
+\BeforeBeginEnvironment{lresult}{\begin{mdframed}}
+\AfterEndEnvironment{lresult}{\end{mdframed}}
+
 \newcommand{\linline}[2]{#2}
@@
 \newenvironment{lresult}
-  {\begin{mdframed}[leftline=false,rightline=false]}
+  {\begin{mdframed}}
   {\end{mdframed}}
-\newcommand{\linplace}[2]{#2}
+\newcommand{\linline}[2]{#2}
 \end{verbatim}
@@
 \section{Executing code snippets}
+
+\subsection{Regular output}
 
@@
 
+%lignore
 \begin{verbatim}
 \begin{lcode}
-W='Hello, world!'
+W='Value of \\textbf{W}!'
 print(W)
@@
 \end{verbatim}
+%lnoignore
 
 \begin{lcode}
-W='Hello, world!'
+W='Value of \\textbf{W}!'
 print(W)
@@
 
+%lignore
 \begin{verbatim}
 \begin{lresult}
-??
+PLACEHOLDER
 \end{lresult}
 \end{verbatim}
+%lnoignore
 
 \begin{lresult}
-Hello, world!
+Value of \textbf{W}!
 \end{lresult}
+
+\subsection{Verbatim input and output}
 
@@
 
+%lignore
+\begin{verbatim}
 \begin{lcode}
-print('Hello, LitREPL.vim!')
+print(W)
 \end{lcode}
+\begin{mdframed}
+%lresult
+PLACEHOLDER
+%lnoresult
+\end{mdframed}
+\end{verbatim}
+%lnoignore
 
+\begin{lcode}
+print(W)
+\end{lcode}
+\begin{mdframed}
+%lresult
+Value of \textbf{W}!
+%lnoresult
+\end{mdframed}
+
+A less often used option of verbatim input is also supported
+
+%lignore
 \begin{verbatim}
-%\begin{lresult}
-??
-%\end{lresult}
+%lcode
+print(r'Value of \textbf{W}!')
+%lnocode
+\begin{lresult}
+PLACEHOLDER
+\end{lresult}
 \end{verbatim}
+%lnoignore
 
-\begin{mdframed}
-%\begin{lresult}
-Hello, LitREPL.vim!
-%\end{lresult}
-\end{mdframed}
+%lcode
+print(r'Value of \textbf{W}!')
+%lnocode
+\begin{lresult}
+Value of \textbf{W}!
+\end{lresult}
+
+\subsection{Inline code}
 
@@
 and as a result container. We expect it to be used to print the contents of
-Python variables, so the VimREPL wraps the executable code with the additional
-\texttt{print(...)} operator. On the Latex side, we define the tag to ignore its
+Python variables, so VimREPL wraps the executable portion with the additional
+\texttt{print(\%s)} operator. On the Latex side, we define the tag to ignore its
 first argument but pass the second one as-is.
 
+%lignore
 \begin{verbatim}
@@
 \end{verbatim}
+%lnoignore
 
-\begin{mdframed}\linline{W}{Hello, world!}\end{mdframed}
+\begin{mdframed}\linline{W}{Value of \textbf{W}!}\end{mdframed}
 
@@
 import matplotlib.pyplot as plt
-X=np.array(range(10))
-plt.plot(X, np.sin(X))
+plt.close()
+X=np.array(range(100))
+plt.plot(X, np.sin(X/10))
 path='_img.png'
@@
 
-\includegraphics{\linline{path}{_img.png}}
+%lignore
+\begin{verbatim}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{\linline{path}{?}}
+\end{center}
+\end{verbatim}
+%lnoignore
+
+\begin{center}
+\includegraphics[width=0.8\textwidth]{\linline{path}{_img.png}}
+\end{center}
 

### 2022-08-22 python/litrepl/base.py

@@ -157,10 +157,12 @@ def stop():
 class SymbolsMarkdown:
   icodebeginmarker="```python"
   icodendmarker="```"
-  ocodebeginmarker="```"
+  icodebeginmarker2="<!--lcode"
+  icodendmarker2="lnocode-->"
+  ocodebeginmarker="```lresult"
   ocodendmarker="```"
-  verbeginmarker="<!--litrepl-->"
-  verendmarker="<!--litrepl-->"
+  verbeginmarker="<!--lresult-->"
+  verendmarker="<!--lnoresult-->"
   combeginmarker=r"<!--lignore-->"
   comendmarker=r"<!--lnoignore-->"
 
@@ -172,20 +174,24 @@ snippet : icodesection -> e_icodesection
         | ocodesection -> e_ocodesection
         | comsection -> e_comsection
 comsection.2 : combeginmarker comtext comendmarker
-icodesection.1 : icodebeginmarker text icodendmarker
-ocodesection.1 : ocodebeginmarker text ocodendmarker
-               | verbeginmarker text verendmarker
+icodesection.1 : icodebeginmarker ctext icodendmarker
+               | icodebeginmarker2 ctext icodendmarker2
+ocodesection.1 : ocodebeginmarker ctext ocodendmarker
+               | verbeginmarker ctext verendmarker
 icodebeginmarker : "{symbols_md.icodebeginmarker}"
 icodendmarker : "{symbols_md.icodendmarker}"
+icodebeginmarker2 : "{symbols_md.icodebeginmarker2}"
+icodendmarker2 : "{symbols_md.icodendmarker2}"
 ocodebeginmarker : "{symbols_md.ocodebeginmarker}"
 ocodendmarker : "{symbols_md.ocodendmarker}"
-verbeginmarker : "<!--litrepl-->"
-verendmarker : "<!--litrepl-->"
+verbeginmarker : "{symbols_md.verbeginmarker}"
+verendmarker : "{symbols_md.verendmarker}"
 inlinebeginmarker : "`"
 inlinendmarker : "`"
 combeginmarker : "{symbols_md.combeginmarker}"
 comendmarker : "{symbols_md.comendmarker}"
-text : /(.(?!```|<\!--litrepl-->|{symbols_md.combeginmarker}|{symbols_md.comendmarker}))*./s
[…]

### 2022-08-22 sh/test.sh

@@ -34,7 +34,7 @@ def hello(name):
 
 hello("LitREPL")
 ```
-```
+```lresult
 Hello, LitREPL!
 ```
 ```python
@@ -44,9 +44,9 @@ hello("World")
 Hello, World!
 ```
 * List
-* <!--litrepl-->
+* <!--lresult-->
   BLABLA
-  <!--litrepl-->
+  <!--lnoresult-->
 EOF
 
 run_latex <<"EOF"
@@ -112,16 +112,16 @@ def hello(name):
 hello('World')
 ```
 TXT
-```
+```lresult
 PLACEHOLDER
 ```
 TXT
 * ```python
   print("ABC"+"DEF")
   ```
-* <!--litrepl-->
+* <!--lresult-->
   PLACEHOLDER
-  <!--litrepl-->
+  <!--lnoresult-->
 == Ignore ==
 <!--lignore-->
 ```python
@@ -133,6 +133,15 @@ PLACEHOLDER
 <!--lnoignore-->
 ```
 NOEVAL
+```
+```lresult
+EVAL
+```
+<!--lcode
+print('FOO')
+lnocode-->
+```lresult
+XX
 ```
 EOF
 cat source.md | runlitrepl --filetype=markdown parse-print >out.md
@@ -146,16 +155,16 @@ def hello(name):
 hello('World')
 ```
 TXT
-```
+```lresult
 Hello, World!
 ```
 TXT
 * ```python
   print("ABC"+"DEF")
   ```
-* <!--litrepl-->
+* <!--lresult-->
   ABCDEF
-  <!--litrepl-->
+  <!--lnoresult-->
 == Ignore ==
 <!--lignore-->
 ```python
@@ -166,7 +175,16 @@ PLACEHOLDER
 ```
 <!--lnoignore-->
 ```
+NOEVAL
+```
+```lresult
 ABCDEF
+```
+<!--lcode
+print('FOO')
+lnocode-->
+```lresult
+FOO
 ```
 EOF
 runlitrepl stop

### 2022-08-22 doc/quickstart.tex
Recorded a latex screencast

--- 2022-08-20 data/example.tex
@@
 \usepackage{minted}
+\renewcommand{\MintedPygmentize}{/nix/store/1r3k7gbfc8bjc4mr5f3bc3j7g7ws24cy-python3-3.8.12-env/bin/pygmentize}
+\usepackage{verbatim}
+
+\newenvironment{myverbatim}
+  {\VerbatimEnvironment
+   \begin{minted}[breaklines,fontsize=\footnotesize]{latex}}
+  {\end{minted}}
+\BeforeBeginEnvironment{myverbatim}{\begin{mdframed}[nobreak=true,frametitle=\tiny{Source}]}
+\AfterEndEnvironment{myverbatim}{\end{mdframed}}
 
@@
   {\VerbatimEnvironment
-   \begin{mdframed}
-   \begin{minted}[breaklines, fontsize=\footnotesize]{python}}
-  {\end{minted}
-   \end{mdframed}}
-\newenvironment{lresult}
-  {\begin{mdframed}}
-  {\end{mdframed}}
+   \begin{minted}[breaklines,fontsize=\footnotesize]{python}}
+  {\end{minted}}
+\BeforeBeginEnvironment{lcode}{\begin{mdframed}[frametitle=\tiny{Rendered}]}
+\AfterEndEnvironment{lcode}{\end{mdframed}}
+
+\newenvironment{lresult}{\verbatim}{\endverbatim}
+\BeforeBeginEnvironment{lresult}{\begin{mdframed}\footnotesize}
+\AfterEndEnvironment{lresult}{\end{mdframed}}
+
+\newcommand{\linline}[2]{#2}
 
@@
 
-\section{Environment tags}
+\section{Organizing the document}
 
-LitREPL for latex recognizes specifically named environments as code and result
-sections. It doesn't really evaluate Tex commands so renaming those environments
-wouldn't work. But we still need to introduce it to Latex so we start with a
-newenvironment declarations.
+\subsection{Preamble}
 
-\begin{verbatim}
+The following Latex code provides a reference Latex definitions of
+\texttt{lcode}, \texttt{lresult} and \texttt{linline} tags. These names are
+currently hardcoded into VimREPL. Users may define them arbitrary.
+
+\begin{myverbatim}
 \newenvironment{lcode}
   {\VerbatimEnvironment
-   \begin{mdframed}
-   \begin{minted}[breaklines, fontsize=\footnotesize]{python}}
-  {\end{minted}
-   \end{mdframed}}
-\newenvironment{lresult}
-  {\begin{mdframed}[leftline=false,rightline=false]}
-  {\end{mdframed}}
-\end{verbatim}
+   \begin{minted}[breaklines,fontsize=\footnotesize]{python}}
+  {\end{minted}}
+\BeforeBeginEnvironment{lcode}{\begin{mdframed}[frametitle=\tiny{Rendered}]}
+\AfterEndEnvironment{lcode}{\end{mdframed}}
 
-\section{Executing code snippets}
+\newenvironment{lresult}{\verbatim}{\endverbatim}
+\BeforeBeginEnvironment{lresult}{\begin{mdframed}\footnotesize}
+\AfterEndEnvironment{lresult}{\end{mdframed}}
 
-Executable code snippet is the content of the \texttt{lcode} environment.
-Putting the cursor on it and typing the \texttt{:LitEval1} command runs it
-it in a background Python interpreter.
+\newcommand{\linline}[2]{#2}
+\end{myverbatim}
 
+\subsection{Code and result sections}
+
+%lignore
+\begin{myverbatim}
 \begin{lcode}
-W='Hello, world!'
+W='Value of \\textbf{W}!'
 print(W)
@@
 
-\texttt{lresult} section next to the executable section is a result container.
-LitREPL would paste here the output of above code snippet. The original content
-of the section will be replaced.
+\begin{lresult}
+\end{lresult}
+\end{myverbatim}
+%lnoignore
+
+\begin{lcode}
+W='Value of \\textbf{W}!'
+print(W)
+\end{lcode}
 
 \begin{lresult}
-Hello, world!
+Value of \textbf{W}!
 \end{lresult}
 
-Commented \texttt{lresult} environmet tags are still recognized as a section for
-verbatim results. This way users can generate parts of the latex document.
+\subsection{Verbatim input and output}
+
+\subsubsection{Verbatim output}
+
+%lignore
+\begin{myverbatim}
+\begin{lcode}
+print(W)
+\end{lcode}
+\begin{mdframed}
+%lresult
+%lnoresult
+\end{mdframed}
+\end{myverbatim}
+%lnoignore
 
 \begin{lcode}
-print('Hello, LitREPL.vim!')
+W='Value of \\textbf{W}!'
+print(W)
+\end{lcode}
+\begin{mdframed}
+%lresult
+Value of \textbf{W}!
+%lnoresult
+\end{mdframed}
+
+\subsubsection{Verbatim input}
+A less often used option of verbatim input is also supported
+
+%lignore
+\begin{myverbatim}
+%lcode
+print('Value of \\textbf{W}!')
+%lnocode
+\begin{lresult}
+\end{lresult}
+\end{myverbatim}
+%lnoignore
+
+%lcode
+print(r'Value of \textbf{W}!')
+%lnocode
+\begin{lresult}
+Value of \textbf{W}!
+\end{lresult}
+
+\subsection{Inline code}
+
+%lignore
+\begin{myverbatim}
+\begin{lcode}
+W='\\textbf{W}'
+\end{lcode}
+\begin{mdframed}Here is an inline value of \linline{W}{}.\end{mdframed}
+\end{myverbatim}
+%lnoignore
+
+\begin{lcode}
+W='\\textbf{W}'
+\end{lcode}
+\begin{mdframed}Here is an inline value of \linline{W}{\textbf{W}}.\end{mdframed}
+
+\break
+\section{Mathplotlib and other graphics}
+
+%lignore
+\begin{myverbatim}
+\begin{lcode}
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+plt.close()
+X=np.array(range(100))
+plt.plot(X, np.sin(X/10))
+os.makedirs('img',exist_ok=True)
+path='img/plot.png'
+plt.savefig(path)
 \end{lcode}
 
-\begin{mdframed}
-%\begin{lresult}
-Hello LitREPL.vim
-%\end{lresult}
-\end{mdframed}
-
-\section{Invoking Mathplotlib code}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{\linline{path}{img/plot.png}}
+\end{center}
+\end{myverbatim}
+%lnoignore
 
 \begin{lcode}
+import os
 import numpy as np
 import matplotlib.pyplot as plt
-X=np.array(range(10))
-_=plt.plot(X, np.sin(X))
-def texplot(f):
-  plt.savefig(f)
-  print("\\includegraphics{./"+f+"}")
-
-texplot('_img.png')
+plt.close()
+X=np.array(range(100))
+plt.plot(X, np.sin(X/10))
+os.makedirs('img',exist_ok=True)
+path='img/plot.png'
+plt.savefig(path)
 \end{lcode}
 
-\newcommand{\linplace}[2]{#2}
-%\begin{lresult}
-\includegraphics{\linplace{VAR}{_img.png}}
-%\end{lresult}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{\linline{path}{img/plot.png}}
+\end{center}
+
+\section{Live Demo}
+
+\begin{lcode}
+W='Hello'
+print(W)
+\end{lcode}
+
+\begin{lresult}
+Hello
+\end{lresult}
 

### 2022-08-22 python/litrepl/base.py

@@ -99,7 +99,7 @@ def fork_ipython(name):
     'c.TerminalInteractiveShell.separate_in = ""\n'
     'c.TerminalInteractiveShell.separate_out = ""\n'
     )
-  system((f'{name} --config=_config.py --debug --logfile=_ipython.log -c '
+  system((f'{name} --config=_config.py --colors=NoColor --logfile=_ipython.log -c '
           '"import os; import sys; sys.ps1=\'\'; sys.ps2=\'\';'
           'os.open(\'_inp.pipe\',os.O_RDWR);'
           'os.open(\'_out.pipe\',os.O_RDWR);"'

### 2022-08-23 python/litrepl/base.py
Add minimal support for tqdm

@@ -3,6 +3,7 @@ import os
 import sys
 import fcntl
 
+from copy import deepcopy
 from typing import List, Optional, Tuple, Set, Dict, Callable
 from re import search, match as re_match
 from select import select
@@ -23,13 +24,69 @@ def mkre(prompt:str):
   return re.compile(f"(.*)(?={prompt})|{prompt}".encode('utf-8'),
                     re.A|re.MULTILINE|re.DOTALL)
 
-def readout(fdr, prompt=mkre('>>>'),timeout:Optional[int]=10)->str:
+def merge_basic(acc,r)->bytes:
+  return acc+r
+
+def merge_rn(acc,r)->bytes:
+  sz=len(acc)
+  sz_r=len(r)
+  acc+=r
+  sz1=sz+sz_r
+  i=sz
+  i_n=sz-1
+  while i<sz1:
+    if acc[i]==10:
+      i_n=i
+      i+=1
+    elif acc[i]==13:
+      acc=acc[:i_n+1]+acc[i+1:]
+      sz1-=(i-i_n)
+      i=i_n+1
+    else:
+      i+=1
+  sz=sz1
+  return acc
+
+def merge_basic2(acc,r,xxx)->Tuple[bytes,int]:
+  return (acc+r,xxx)
+
+def merge_rn2(buf_,r,i_n=-1)->Tuple[bytes,int]:
+  buf=deepcopy(buf_)
+  sz=len(buf)
+  i_n=-(sz-i_n)
+  start=0
+  for i in range(len(r)):
+    if r[i]==10:
+      i_n=i
+      buf+=r[start:i+1]
+      start=i+1
+    elif r[i]==13:
+      if i_n<0:
+        buf=buf[:sz+i_n+1]
+        sz=len(buf)
+        i_n=-1
+      start=i+1
+  buf+=r[start:]
+  if i_n>=0:
+    i_n-=start
+  assert b'\r' not in buf
+  if i_n>=0:
+    assert i_n<len(buf), f"{len(buf)}, {i_n}"
+    assert buf[i_n]==(b'\n'[0]), f"{buf}, {i_n}, {buf[i_n]}"
+  return buf,sz+i_n if i_n<0 else sz+i_n
+
+def readout(fdr,
+            prompt=mkre('>>>'),
+            timeout:Optional[int]=10,
+            merge=merge_basic2)->str:
   acc:bytes=b''
+  i_n=-1
   while select([fdr],[],[],timeout)[0] != []:
[…]

### 2022-08-23 sh/test.sh

@@ -295,6 +295,25 @@ EOF
 runlitrepl stop
 )}
 
+test_tqdm() {(
+mktest "_test_eval_md"
+runlitrepl start
+cat >source.md <<"EOF"
+```python
+from tqdm import tqdm
+for i in tqdm(range(100)):
+  pass
+```
+```lresult
+```
+EOF
+cat source.md | runlitrepl --filetype=markdown parse-print >out.md
+diff -u source.md out.md
+cat source.md | runlitrepl --filetype=markdown eval-sections '0..$' >out.md
+test "$(cat out.md | grep '100%' | wc -l)" = "1"
+runlitrepl stop
+)}
+
 if test -n '$LITREPL_TEST' || echo "$(basename $0)" | grep -q "test.sh" ; then
   set -e -x
   if test -z "$LITREPL"; then
@@ -306,6 +325,7 @@ if test -n '$LITREPL_TEST' || echo "$(basename $0)" | grep -q "test.sh" ; then
     LITREPL_INTERPRETER=$I
     test_parse_print
     test_eval_md
+    test_tqdm
     test_eval_tex
   done
   trap "" EXIT

### 2023-03-11 python/litrepl/base.py
Fix tex linline parser issue

@@ -171,6 +171,11 @@ class SymbolsLatex:
 
 symbols_latex=SymbolsLatex()
 
+OBR="{"
+CBR="}"
+BCBR="\\}"
+BOBR="\\{"
+
 def toplevel_markers_latex():
   sl=symbols_latex
   return '|'.join([sl.icodebeginmarker,sl.icodendmarker,
@@ -178,10 +183,7 @@ def toplevel_markers_latex():
                    sl.ocodebeginmarker,sl.ocodendmarker,
                    sl.verbeginmarker,sl.verendmarker,
                    sl.combeginmarker,sl.comendmarker,
-                   sl.inlinemarker])
-OBR="{"
-CBR="}"
-BCBR="\\}"
+                   sl.inlinemarker+BOBR])
 grammar_latex = fr"""
 start: (topleveltext)? (snippet (topleveltext)? )*
 snippet : icodesection -> e_icodesection

### 2023-03-11 sh/test.sh

@@ -229,6 +229,9 @@ test_eval_tex() {( #{{{
 mktest "_test_eval_tex"
 runlitrepl start
 cat >source.tex <<"EOF"
+% == Ignore non-tags ==
+\newcommand{\linline}[2]{#2}
+% == Normal evaluation ==
 \begin{lcode}
 print("A"+"B")
 \end{lcode}
@@ -240,16 +243,16 @@ Text
   %lnoresult
 \end{enumerate}
 \begin{lresult}
-NOEVAL
+PLACEHOLDER
 \end{lresult}
-== Inline ==
+% == Inline evaluation ==
 \linline{"A"+"B"}
 {XX}\linline{"C"+"D"}{}
 \begin{lcode}
 tag='\\textbf{bold}'
 \end{lcode}
 \linline{tag+tag}{\textbf{old}\textbf{old}}
-== Ignore ==
+% == Meta-comments ==
 %lignore
 \begin{lcode}
 print("X"+"Y")
@@ -266,6 +269,9 @@ cat source.tex | runlitrepl --filetype=latex parse-print >out.tex
 diff -u source.tex out.tex
 cat source.tex | runlitrepl --filetype=latex eval-sections '0..$' >out.tex
 diff -u out.tex - <<"EOF"
+% == Ignore non-tags ==
+\newcommand{\linline}[2]{#2}
+% == Normal evaluation ==
 \begin{lcode}
 print("A"+"B")
 \end{lcode}
@@ -279,14 +285,14 @@ Text
 \begin{lresult}
 AB
 \end{lresult}
-== Inline ==
+% == Inline evaluation ==
 \linline{"A"+"B"}
 {AB}\linline{"C"+"D"}{CD}
 \begin{lcode}
 tag='\\textbf{bold}'
 \end{lcode}
 \linline{tag+tag}{\textbf{bold}\textbf{bold}}
-== Ignore ==
+% == Meta-comments ==
 %lignore
 \begin{lcode}
 print("X"+"Y")

### 2023-03-11 sh/test.sh
Re-design test runner script

@@ -1,7 +1,6 @@
 #!/bin/sh
 
 mktest() {
-  set -e -x
   test -d "$LITREPL_ROOT"
   T="$LITREPL_ROOT/_test/$1"
   rm -rf  "$T" || true
@@ -9,13 +8,9 @@ mktest() {
   cd "$T"
 }
 
-runlitrepl() {
-  test -n "$LITREPL_INTERPRETER"
-  $LITREPL --interpreter="$LITREPL_INTERPRETER" "$@"
-}
-
 test_parse_print() {( # {{{
 mktest "_test_parse_print"
+runlitrepl start
 run_md() {
   cat > "source.md"
   cat "source.md" | runlitrepl --filetype=markdown parse-print > "parsed.md"
@@ -375,22 +370,56 @@ grep -q 'BG' out2.md
 runlitrepl stop
 )} #}}}
 
-if test -n "$LITREPL_TEST" || echo "$(basename $0)" | grep -q "test.sh" ; then
-  set -e -x
-  if test -z "$LITREPL"; then
-    LITREPL=$LITREPL_ROOT/python/bin/litrepl
-  fi
-  trap "echo FAIL" EXIT
-  for I in python ipython ; do
-    echo "Checking $I"
-    LITREPL_INTERPRETER=$I
-    test_parse_print
-    time test_eval_md
-    test_tqdm
-    time test_eval_tex
-    test_async
-  done
-  trap "" EXIT
-  echo OK
+
+interpreters() {
+  echo "$(which python)"
+  echo "$(which ipython)"
+}
+
+tests() {
+  echo test_parse_print
+  echo test_eval_md
+  echo test_tqdm
+  echo test_eval_tex
+  echo test_async
+}
+
+runlitrepl() {
+  test -n "$LITREPL_INTERPRETER"
+  test -n "$LITREPL_BIN"
+  $LITREPL_BIN --interpreter="$LITREPL_INTERPRETER" "$@"
+}
+
+set -e
+
+INTERPS=`interpreters`
+TESTS=`tests`
+while test -n "$1" ; do
+  case "$1" in
+    -i) INTERPS="$2"; shift ;;
+    --interpreters=*) INTERPS=$(echo "$1" | sed 's/.*=//g') ;;
+    -t) TESTS="$2"; shift ;;
+    --tests) TESTS=$(echo "$1" | sed 's/.*=//g') ;;
[…]

### 2023-03-11 python/litrepl/__init__.py
Read version properly

@@ -1,3 +1,5 @@
+from typing import Optional
+from os.path import join
 from logging import getLogger
 logger=getLogger(__name__)
 info=logger.info
@@ -7,14 +9,29 @@ from .types import *
 from .eval import *
 from .base import *
 
+# Determine the package version using the following source priorities:
+# 1) version.txt+git 2) version.txt 3) version.py
+__version__:Optional[str]
 try:
-  from litrepl.version import __version__
-except ImportError:
+  LITREPL_ROOT=environ['LITREPL_ROOT']
+  ver=open(join(LITREPL_ROOT,'version.txt')).read().strip()
+  LITREPL_REVISION:Optional[str]
   try:
-    from setuptools_scm import get_version
-    from os.path import join
-    __version__ = get_version(root=join('..','..'), relative_to=__file__)
+    from subprocess import check_output
+    LITREPL_REVISION=check_output(['git', 'rev-parse', 'HEAD'],
+                                  cwd=LITREPL_ROOT).decode().strip()
   except Exception:
+    try:
+      LITREPL_REVISION=environ["LITREPL_REVISION"]
+    except Exception:
+      LITREPL_REVISION=None
+
+  rev=f"+g{LITREPL_REVISION[:7]}" if LITREPL_REVISION is not None else ""
+  __version__ = f"{ver}{rev}"
+except Exception:
+  try:
+    from litrepl.version import __version__
+  except ImportError:
     warning("Neither `litrepl/version.py` module was not generated during the "
             "setup, nor the Git metadata is available. Re-install litrepl with "
             "the `setuptools_scm` package available to fix")

### 2023-03-11 setup.py

@@ -6,6 +6,10 @@ from typing import Optional
 logger=getLogger(__name__)
 warning=logger.warning
 
+# Determine the revision using the following sourcr priorities:
+# 1) version.txt+git 2) version.txt
+# and save it to version.py
+ver=open(join(dirname(__file__),'version.txt')).read().strip()
 LITREPL_REVISION:Optional[str]
 try:
   LITREPL_REVISION=environ["LITREPL_REVISION"]
@@ -13,13 +17,14 @@ except Exception:
   warning("Couldn't read LITREPL_REVISION, trying `git rev-parse`")
   try:
     from subprocess import check_output
-    LITREPL_REVISION=check_output(['git', 'rev-parse', 'HEAD']).decode().strip()
-  except Exception:
+    import sys
+    LITREPL_REVISION=check_output(['git', 'rev-parse', 'HEAD'],
+                                  cwd=environ['LITREPL_ROOT']).decode().strip()
+  except Exception as e:
     warning("Couldn't use `git rev-parse`, no revision metadata will be set")
     LITREPL_REVISION=None
 
 with open(join('python','litrepl','version.py'), 'w') as f:
-  ver=open(join(dirname(__file__),'version.txt')).read().strip()
   rev=f"+g{LITREPL_REVISION[:7]}" if LITREPL_REVISION is not None else ""
   f.write("# AUTOGENERATED by setup.exe!\n")
   f.write(f"__version__ = '{ver}{rev}'\n")

### 2023-03-11 setup.py
Minor fixes; Version 2.0.5

@@ -40,7 +40,7 @@ setup(
   python_requires='>=3.6',
   author="Sergei Mironov",
   author_email="grrwlf@gmail.com",
-  description="LitREPL a macroprocessing Python library for Litrate "\
+  description="LitREPL is a macroprocessing Python library for Litrate "\
               "programming and code execution.",
   classifiers=[
     "Programming Language :: Python :: 3",

### 2023-03-11 python/litrepl/__init__.py
Redesign versioning system; Version 2.0.6

@@ -11,28 +11,34 @@ from .base import *
 
 # Determine the package version using the following source priorities:
 # 1) version.txt+git 2) version.txt 3) version.py
-__version__:Optional[str]
+__revision__:Optional[str]
 try:
-  LITREPL_ROOT=environ['LITREPL_ROOT']
-  ver=open(join(LITREPL_ROOT,'version.txt')).read().strip()
-  LITREPL_REVISION:Optional[str]
-  try:
-    from subprocess import check_output
-    LITREPL_REVISION=check_output(['git', 'rev-parse', 'HEAD'],
-                                  cwd=LITREPL_ROOT).decode().strip()
-  except Exception:
-    try:
-      LITREPL_REVISION=environ["LITREPL_REVISION"]
-    except Exception:
-      LITREPL_REVISION=None
-
-  rev=f"+g{LITREPL_REVISION[:7]}" if LITREPL_REVISION is not None else ""
-  __version__ = f"{ver}{rev}"
+  __revision__=environ["LITREPL_REVISION"]
 except Exception:
   try:
-    from litrepl.version import __version__
-  except ImportError:
-    warning("Neither `litrepl/version.py` module was not generated during the "
-            "setup, nor the Git metadata is available. Re-install litrepl with "
-            "the `setuptools_scm` package available to fix")
-    __version__ = None
+    from subprocess import check_output
+    import sys
+    __revision__=check_output(['git', 'rev-parse', 'HEAD'],
+                            cwd=environ['LITREPL_ROOT']).decode().strip()
+  except Exception:
+    try:
+      from litrepl.revision import __revision__ as __rv__
+      __revision__=__rv__
+    except ImportError:
+      __revision__=None
+
+
+__semver__:Optional[str]
+try:
+  __semver__=open(join(environ['LITREPL_ROOT'],'version.txt')).read().strip()
+except Exception:
+    try:
+      from litrepl.semver import __semver__ as __sv__
+      __semver__= __sv__
+    except ImportError:
+      __semver__=None
+
+
+__version__:Optional[str]
+__version__=(__semver__ + (f"+g{__revision__[:7]}" \
+                           if __revision__ else "")) if __semver__ else None

### 2023-03-11 setup.py

@@ -6,10 +6,6 @@ from typing import Optional
 logger=getLogger(__name__)
 warning=logger.warning
 
-# Determine the revision using the following sourcr priorities:
-# 1) version.txt+git 2) version.txt
-# and save it to version.py
-ver=open(join(dirname(__file__),'version.txt')).read().strip()
 LITREPL_REVISION:Optional[str]
 try:
   LITREPL_REVISION=environ["LITREPL_REVISION"]
@@ -20,19 +16,37 @@ except Exception:
     import sys
     LITREPL_REVISION=check_output(['git', 'rev-parse', 'HEAD'],
                                   cwd=environ['LITREPL_ROOT']).decode().strip()
-  except Exception as e:
+  except Exception:
     warning("Couldn't use `git rev-parse`, no revision metadata will be set")
     LITREPL_REVISION=None
 
-with open(join('python','litrepl','version.py'), 'w') as f:
-  rev=f"+g{LITREPL_REVISION[:7]}" if LITREPL_REVISION is not None else ""
-  f.write("# AUTOGENERATED by setup.exe!\n")
-  f.write(f"__version__ = '{ver}{rev}'\n")
+
+LITREPL_SEMVER:Optional[str]
+try:
+  LITREPL_SEMVER=open(join(dirname(__file__),'semver.txt')).read().strip()
+except Exception:
+  warning("Couldn't read 'version.txt', no metadata will be set")
+  LITREPL_SEMVER=None
+
+
+if LITREPL_REVISION:
+  with open(join('python','litrepl','revision.py'), 'w') as f:
+    f.write("# AUTOGENERATED by setup.exe!\n")
+    f.write(f"__revision__ = '{LITREPL_REVISION}'\n")
+
+if LITREPL_SEMVER:
+  with open(join('python','litrepl','semver.py'), 'w') as f:
+    f.write("# AUTOGENERATED by setup.exe!\n")
+    f.write(f"__semver__ = '{LITREPL_SEMVER}'\n")
+
+LITREPL_VERSION=((LITREPL_SEMVER + (f"+g{LITREPL_REVISION[:7]}" if
+                                    LITREPL_REVISION else ""))
+                 if LITREPL_SEMVER else "")
 
 setup(
   name="litrepl",
   zip_safe=False, # https://mypy.readthedocs.io/en/latest/installed_packages.html
-  version=f"{ver}",
+  version=LITREPL_VERSION,
   package_dir={'':'python'},
   packages=find_packages(where='python'),
   install_requires=['lark'],

### 2023-03-11 setup.py
Use semver for pypi version; Version 2.0.7

@@ -39,14 +39,10 @@ if LITREPL_SEMVER:
     f.write("# AUTOGENERATED by setup.exe!\n")
     f.write(f"__semver__ = '{LITREPL_SEMVER}'\n")
 
-LITREPL_VERSION=((LITREPL_SEMVER + (f"+g{LITREPL_REVISION[:7]}" if
-                                    LITREPL_REVISION else ""))
-                 if LITREPL_SEMVER else "")
-
 setup(
   name="litrepl",
   zip_safe=False, # https://mypy.readthedocs.io/en/latest/installed_packages.html
-  version=LITREPL_VERSION,
+  version=LITREPL_SEMVER if LITREPL_SEMVER else "999",
   package_dir={'':'python'},
   packages=find_packages(where='python'),
   install_requires=['lark'],

### 2023-03-11 python/litrepl/base.py
litrepl: introduce --workdir option

@@ -10,34 +10,49 @@ from select import select
 from os import environ, system
 from lark import Lark, Visitor, Transformer, Token, Tree
 from lark.visitors import Interpreter
-from os.path import isfile
+from os.path import isfile, join
 from signal import signal, SIGINT
 from time import sleep
-from dataclasses import dataclass
+from dataclasses import dataclass, astuple
 from functools import partial
 from argparse import ArgumentParser
 from collections import defaultdict
+from os import makedirs,getuid,getcwd
+from tempfile import gettempdir
+from hashlib import sha256
 
-from .types import PrepInfo, RunResult, NSec, FileName, SecRec
+from .types import PrepInfo, RunResult, NSec, FileName, SecRec, FileNames
 from .eval import (process, pstderr, rresultLoad, rresultSave, processAdapt,
                    processCont)
 
-def fork_python(name):
+def pipenames(a)->FileNames:
+  """ Return file names of in.pipe, out.pip and log """
+  workdir=a.workdir if a.workdir is not None else \
+    join(gettempdir(),"litrepl_"+sha256(getcwd().encode('utf-8')).hexdigest()[:6])
+  makedirs(workdir, exist_ok=True)
+  return FileNames(workdir, join(workdir,"_in.pipe"), join(workdir,"_out.pipt"),
+                   join(workdir,"_pid.txt"))
+
+def fork_python(a, name):
   assert 'python' in name
+  wd,inp,outp,pid=astuple(pipenames(a))
   system((f'{name} -uic "import os; import sys; sys.ps1=\'\'; sys.ps2=\'\';'
-          'os.open(\'_inp.pipe\',os.O_RDWR|os.O_SYNC);'
-          'os.open(\'_out.pipe\',os.O_RDWR|os.O_SYNC);"'
-          '<_inp.pipe >_out.pipe 2>&1 & echo $! >_pid.txt'))
-  open('_inp.pipe','w').write(
+          f'os.open(\'{inp}\',os.O_RDWR|os.O_SYNC);'
+          f'os.open(\'{outp}\',os.O_RDWR|os.O_SYNC);"'
+          f'<"{inp}" >"{outp}" 2>&1 & echo $! >"{pid}"'))
+  open(inp,'w').write(
     'import signal\n'
     'def _handler(signum,frame):\n'
     '  raise KeyboardInterrupt()\n\n'
[…]

### 2023-03-11 python/litrepl/eval.py

@@ -11,15 +11,15 @@ from select import select
 from os import environ, system
 from lark import Lark, Visitor, Transformer, Token, Tree
 from lark.visitors import Interpreter
-from os.path import isfile
+from os.path import isfile, join
 from signal import signal, SIGINT, SIGALRM, setitimer, ITIMER_REAL
 from time import sleep
-from dataclasses import dataclass
+from dataclasses import dataclass, astuple
 from functools import partial
 from argparse import ArgumentParser
 from contextlib import contextmanager
 
-from .types import RunResult, ReadResult, FileName
+from .types import RunResult, ReadResult, FileNames
 
 def pstderr(*args,**kwargs):
   print(*args, file=sys.stderr, **kwargs, flush=True)
@@ -113,10 +113,11 @@ def pusererror(fname,err)->None:
   with open(fname,"w") as f:
     f.write(err)
 
-def processAsync(lines:str)->RunResult:
+def processAsync(fns:FileNames, lines:str)->RunResult:
   """ Send `lines` to the interpreter and fork the response reader """
+  wd,inp,outp,pidf=astuple(fns)
   codehash=abs(hash(lines))
-  fname=f"/tmp/litrepl-eval-{codehash}.txt"
+  fname=join(wd,f"litrepl_eval_{codehash}.txt")
   pattern=PATTERN
   fo=os.open(fname,os.O_WRONLY|os.O_SYNC|os.O_TRUNC|os.O_CREAT)
   assert fo>0
@@ -128,8 +129,8 @@ def processAsync(lines:str)->RunResult:
     # Child
     fdr=0; fdw=0
     try:
-      fdw=os.open('_inp.pipe', os.O_WRONLY|os.O_SYNC)
-      fdr=os.open('_out.pipe', os.O_RDONLY|os.O_SYNC)
+      fdw=os.open(inp, os.O_WRONLY|os.O_SYNC)
+      fdr=os.open(outp, os.O_RDONLY|os.O_SYNC)
       if fdw<0 or fdr<0:
         pusererror(fname,f"ERROR: litrepl.py couldn't open session pipes\n")
       fcntl.flock(fdw,fcntl.LOCK_EX|fcntl.LOCK_NB)
@@ -157,8 +158,8 @@ def processAsync(lines:str)->RunResult:
     return RunResult(fname,pattern)
 
 @contextmanager
-def with_sigint(brk=False,ipid:Optional[int]=None):
[…]

### 2023-03-11 python/litrepl/types.py

@@ -30,3 +30,10 @@ class SecRec:
   """ Request for section evaluation """
   nsecs:Set[NSec]
   pending:Dict[NSec,RunResult]
+
+@dataclass
+class FileNames:
+  wd:str
+  inp:str
+  outp:str
+  pidf:str

### 2023-03-11 python/litrepl/base.py
litrepl: Debug-print work dir

@@ -17,7 +17,7 @@ from dataclasses import dataclass, astuple
 from functools import partial
 from argparse import ArgumentParser
 from collections import defaultdict
-from os import makedirs,getuid,getcwd
+from os import makedirs, getuid, getcwd
 from tempfile import gettempdir
 from hashlib import sha256
 
@@ -25,10 +25,18 @@ from .types import PrepInfo, RunResult, NSec, FileName, SecRec, FileNames
 from .eval import (process, pstderr, rresultLoad, rresultSave, processAdapt,
                    processCont)
 
+DEBUG:bool=False
+
+def pdebug(*args,**kwargs):
+  if DEBUG:
+    print(*args, file=sys.stderr, **kwargs, flush=True)
+
 def pipenames(a)->FileNames:
   """ Return file names of in.pipe, out.pip and log """
   workdir=a.workdir if a.workdir is not None else \
-    join(gettempdir(),"litrepl_"+sha256(getcwd().encode('utf-8')).hexdigest()[:6])
+    join(gettempdir(),f"litrepl_{getuid()}_"+
+         sha256(getcwd().encode('utf-8')).hexdigest()[:6])
+  pdebug(f"Workdir: {workdir}")
   makedirs(workdir, exist_ok=True)
   return FileNames(workdir, join(workdir,"_in.pipe"), join(workdir,"_out.pipt"),
                    join(workdir,"_pid.txt"))

### 2023-03-11 python/litrepl/base.py
litrepl: Add status command

@@ -33,13 +33,13 @@ def pdebug(*args,**kwargs):
 
 def pipenames(a)->FileNames:
   """ Return file names of in.pipe, out.pip and log """
-  workdir=a.workdir if a.workdir is not None else \
+  auxdir=a.auxdir if a.auxdir is not None else \
     join(gettempdir(),f"litrepl_{getuid()}_"+
          sha256(getcwd().encode('utf-8')).hexdigest()[:6])
-  pdebug(f"Workdir: {workdir}")
-  makedirs(workdir, exist_ok=True)
-  return FileNames(workdir, join(workdir,"_in.pipe"), join(workdir,"_out.pipt"),
-                   join(workdir,"_pid.txt"))
+  pdebug(f"Auxdir: {auxdir}")
+  makedirs(auxdir, exist_ok=True)
+  return FileNames(auxdir, join(auxdir,"_in.pipe"), join(auxdir,"_out.pipe"),
+                   join(auxdir,"_pid.txt"))
 
 def fork_python(a, name):
   assert 'python' in name
@@ -90,7 +90,7 @@ def fork_ipython(a, name):
 
 def start_(a, fork_handler:Callable[...,None])->None:
   """ Starts the background Python interpreter. Kill an existing interpreter if
-  any. Creates files `_inp.pipe`, `_out.pipt`, `_pid.txt`."""
+  any. Creates files `_inp.pipe`, `_out.pipe`, `_pid.txt`."""
   wd,inp,outp,pid=astuple(pipenames(a))
   if isfile(pid):
     system(f'kill "$(cat {pid})" >/dev/null 2>&1')

### 2023-03-11 sh/mkpdflatex.sh
env: Add mkpdflatex.sh script; Support SVG images

@@ -1,0 +1,27 @@
+#!/bin/sh
+
+set -e -x
+
+CWD="$LITREPL_ROOT"
+TEX="$1"
+shift
+test -f "$TEX"
+
+D=$CWD/_pdflatex
+mkdir $D 2>/dev/null || true
+
+TEXminted=$(dirname "$TEX")/$(basename "$TEX" .tex).minted.tex
+trap "rm $TEXminted" 0 1 2 9
+
+(
+cd $(dirname "$TEX")
+sed "s@usepackage{minted}@usepackage[outputdir=$D]{minted}@" \
+  "$(basename $TEX)" > "$(basename $TEXminted)"
+pdflatex \
+  --shell-escape \
+  --output-directory="$D" \
+  "$(basename $TEXminted)" "$@"
+)
+
+cp $D/$(basename "$TEXminted" .tex).pdf \
+   $(dirname "$TEX")/$(basename "$TEX" .tex).pdf

### 2023-03-11 setup.py
setup.py: not relying on LITREPL_ROOT

@@ -15,7 +15,7 @@ except Exception:
     from subprocess import check_output
     import sys
     LITREPL_REVISION=check_output(['git', 'rev-parse', 'HEAD'],
-                                  cwd=environ['LITREPL_ROOT']).decode().strip()
+                                  cwd=dirname(__file__)).decode().strip()
   except Exception:
     warning("Couldn't use `git rev-parse`, no revision metadata will be set")
     LITREPL_REVISION=None

### 2023-03-11 doc/example.tex
Update docs

@@
 
-LitREPL for latex recognizes \texttt{lcode} environments as code and
-\texttt{lresult} as result sections. The tag names is currently hardcoded into
-the simple parser the tool is using, so we need to additionally introduce it to
-the Latex system. Here we do it in a most simple way.
+LitREPL treats \texttt{lcode} environment as code sections and \texttt{lresult}
+environment as result sections. The names are currently hardcoded into the
+simplified LitREPL parser. Wrapping it in other tags is not allowed.
+
+LaTeX does not know anything about these environments by default, we need to
+introduce them to be able to compile the document using e.g. \texttt{pdflatex}.
 
@@
 Executable section is the text between the \texttt{lcode} begin/end tags.
-Putting the cursor on it and typing the \texttt{:LitEval1} executes it in the
-background Python interpreter.
 
@@
 
-\texttt{lresult} tags next to the executable section mark the result section.
-The output of the executable section will be pasted here. The
-original content of the section will be replaced.
+Putting the cursor on it and typing the \texttt{:LitEval1} runs the code in the
+background Python interpreter.
+
+\texttt{lresult} begin/end tags mark the result section.  LitREPL replaces its
+content with the above code section's execution result.
 
@@
 
-Commented \texttt{lresult}/\texttt{lnoresult} tags also marks result sections.
-This way we could customise the Latex markup for every particular section.
+LitREPL recognizes \texttt{l[no]code}/\texttt{l[no]result} comments as
+code/result section markers. This way we can use Python to produce LaTeX markup
+as output.
 
-\begin{lcode}
+%lcode
 print("Hi!")
-\end{lcode}
+%lnocode
 
@@
 
-Additionally, VimREPL for Latex recognises \texttt{linline} tags for which it
-prints its first argument and pastes the result in place of the second argument.
+Additionally, VimREPL recognises \texttt{linline} 2-argument tags. The first
+arguement is treaten as a Python printable expression. The second arguemnt is to
+be replaced with the printing output. In our simplified definition, we simply
+ignore the first argument and paste the second to the LaTeX processor as-is.
 

### 2024-02-09 python/litrepl/__init__.py
Update version.txt -> semver.txt

@@ -11,7 +11,7 @@ from .eval import *
 from .base import *
 
 # Determine the package version using the following source priorities:
-# 1) version.txt+git 2) version.txt 3) version.py
+# 1) semver.txt+git 2) semver.txt 3) version.py
 __revision__:Optional[str]
 try:
   __revision__=environ["LITREPL_REVISION"]
@@ -31,7 +31,7 @@ except Exception:
 
 __semver__:Optional[str]
 try:
-  __semver__=open(join(environ['LITREPL_ROOT'],'version.txt')).read().strip()
+  __semver__=open(join(environ['LITREPL_ROOT'],'semver.txt')).read().strip()
 except Exception:
     try:
       from litrepl.semver import __semver__ as __sv__

### 2024-02-09 setup.py
Fix a typo

@@ -25,7 +25,7 @@ LITREPL_SEMVER:Optional[str]
 try:
   LITREPL_SEMVER=open(join(dirname(__file__),'semver.txt')).read().strip()
 except Exception:
-  warning("Couldn't read 'version.txt', no metadata will be set")
+  warning("Couldn't read 'semver.txt', no metadata will be set")
   LITREPL_SEMVER=None
 
 

### 2024-02-10 setup.py
Add Pypi description

@@ -39,12 +39,17 @@ if LITREPL_SEMVER:
     f.write("# AUTOGENERATED by setup.exe!\n")
     f.write(f"__semver__ = '{LITREPL_SEMVER}'\n")
 
+with open("README.md", "r") as f:
+  long_description = f.read()
+
 setup(
   name="litrepl",
   zip_safe=False, # https://mypy.readthedocs.io/en/latest/installed_packages.html
   version=LITREPL_SEMVER if LITREPL_SEMVER else "999",
   package_dir={'':'python'},
   packages=find_packages(where='python'),
+  long_description=long_description,
+  long_description_content_type="text/markdown",
   install_requires=['lark', 'psutil'],
   scripts=['./python/bin/litrepl'],
   python_requires='>=3.6',

### 2024-07-18 python/litrepl/eval.py
Revert sigint handler udpate

@@ -56,17 +56,15 @@ def with_sigmask(signals=None):
 
 @contextmanager
 def with_sigint(a:LitreplArgs, fns:FileNames):
-  ipid=int(open(fns.pidf).read())
   def _handler(signum,frame):
-    if a.propagate_sigint:
-      pdebug(f"Sending SIGINT to {ipid}")
-      os.kill(ipid,SIGINT)
-    else:
-      pdebug(f"Ignoring SIGINT")
+    pdebug(f"Sending SIGINT to {ipid}")
+    os.kill(ipid,SIGINT)
+  ipid=int(open(fns.pidf).read())
   prev=None
   try:
     with with_sigmask():
-      prev=signal(SIGINT,_handler)
+      if a.propagate_sigint:
+        prev=signal(SIGINT,_handler)
     yield
   finally:
     with with_sigmask():

### 2024-07-18 sh/airepl.sh
Add airepl.sh

@@ -1,0 +1,30 @@
+#!/bin/sh
+
+echo "WARNING: THIS SCRIPT EXECUTES AI-GENERATED PROGRAMS!" >&2
+
+litrepl restart ai
+(
+echo "/model \"~/.local/share/nomic.ai/GPT4All/Meta-Llama-3-8B-Instruct.Q4_0.gguf\""
+echo "Please be very brief! How can I convert 'input.gif' to 'output.webm' using ffmpeg oneliner?"
+) | tee >(cat 1>&2) \
+  | litrepl eval-code ai \
+  | tee >(cat 1>&2) \
+  | sed -n 's/^`\?\(ffmpeg[^`]*\)`\?$/\1/p' \
+  >_cmd.sh
+
+while ! (
+        rm output.webm 2>/dev/null ; \
+        test -s _cmd.sh || { echo "Please suggest an ffmpeg oneliner!" >&2 ; exit 1; } && \
+        sh _cmd.sh && \
+        test -s output.webm || { echo "OUTPUT FILE IS EMPTY!" >&2 ; exit 1; } ; \
+        ) 2>_err.txt ; do
+  (
+  echo $'Nope! The output was:\n\n'
+  cat _err.txt | tail
+  echo $'\n\nPlease make another attempt!'
+  ) | tee >(cat 1>&2) \
+    | litrepl eval-code ai \
+    | tee >(cat 1>&2) \
+    | sed -n 's/^`\?\(ffmpeg[^`]*\)`\?$/\1/p' \
+    > _cmd.sh
+done

### 2024-07-18 sh/test.sh
Fix empty filetype bug; Version 3.2.1

@@ -681,6 +681,19 @@ grep -q -v '1234' file.md
 )}
 #}}}
 
+test_vim_lstatus() {( #{{{
+mktest "_test_vim_lstatus"
+runlitrepl start
+
+runvim >_vim.log 2>&1 <<"EOF"
+:let g:litrepl_errfile = '_litrepl.err'
+:LStatus
+:wqa!
+EOF
+grep -q -v -i error _litrepl.err
+)}
+#}}}
+
 test_foreground() {( #{{{
 mktest "_test_foreground"
 runlitrepl start
@@ -779,6 +792,7 @@ tests() {
   echo test_vim_leval_cursor
   echo test_vim_leval_explicit
   echo test_vim_lmon
+  echo test_vim_lstatus
   echo test_foreground
   echo test_status
   echo test_interrupt

### 2024-07-18 sh/airepl.sh
Update airepl.sh; Update README.md

@@ -1,6 +1,11 @@
 #!/bin/sh
 
-echo "WARNING: THIS SCRIPT EXECUTES AI-GENERATED PROGRAMS!" >&2
+(
+echo 'WARNING: THIS SCRIPT EXECUTES AI-GENERATED PROGRAMS!'
+echo 'ARE YOU USING SYSTEM-LEVEL CONTAINER?'
+echo '(Press Enter to continue; Press CTRL+C to cancel)'
+) >&2
+read
 
 litrepl restart ai
 (

### 2024-07-18 sh/test.sh
Remove incorrect grep -v from tests; Change LEvalAbove into LEval above etc.

@@ -528,7 +528,7 @@ echo $?>ret.txt
 test `cat ret.txt` = "123"
 grep -q "^before-exit" out.md
 grep -q "123" out.md
-grep -q -v "^after-exit" out.md
+not grep -q "^after-exit" out.md
 )} #}}}
 
 test_exception_errcode() {( #{{{
@@ -560,7 +560,7 @@ echo $?>ret.txt
 test `cat ret.txt` = "123"
 grep -q "^before-exception" out.md
 grep -q "123" out.md
-grep -q -v "^after-exception" out.md
+not grep -q "^after-exception" out.md
 runlitrepl stop
 )} #}}}
 
@@ -622,7 +622,7 @@ runvim file.md >_vim.log 2>&1 <<EOF
 :LEval
 :wqa!
 EOF
-grep -q -v '^result-1' file.md
+not grep -q '^result-1' file.md
 grep -q '^result-2' file.md
 )}
 #}}}
@@ -631,7 +631,7 @@ test_vim_leval_explicit() {( #{{{
 mktest "_test_vim_leval_explicit"
 runlitrepl start
 
-cat >file.md <<"EOF"
+cat >source.md <<"EOF"
 ``` python
 print("result-1")
 ```
@@ -645,14 +645,51 @@ print("result-2")
 
 ``` result
 ```
+
+``` python
+print("result-3")
+```
+
+``` result
+```
 EOF
 
-runvim file.md >_vim.log 2>&1 <<"EOF"
+runvim >_vim.log 2>&1 <<"EOF"
+:e! source.md
 :LEval 1
-:wq!
+:w! file1.md
+
+:e! source.md
+:LEval all
+:w! file-all.md
+
+:e! source.md
+9G
+:LEval above
+:w! file-above.md
+
+:e! source.md
+9G
+:LEval below
+:w! file-below.md
+:qa!
 EOF
-grep -q -v '^result-1' file.md
-grep -q '^result-2' file.md
+
+not grep -q '^result-1' file1.md
+grep -q '^result-2' file1.md
+not grep -q '^result-3' file1.md
+
+grep -q '^result-1' file-all.md
+grep -q '^result-2' file-all.md
+grep -q '^result-3' file-all.md
+
+grep -q '^result-1' file-above.md
+grep -q '^result-2' file-above.md
+not grep -q '^result-3' file-above.md
+
+not grep -q '^result-1' file-below.md
[…]

### 2024-07-18 sh/screencast.sh
Add screencast.sh

@@ -1,0 +1,16 @@
+#!/bin/sh
+
+PIDS=''
+trap "kill \$PIDS" SIGINT
+st -t litrepl-demo -e vim-demo -X _test/_test1.md &
+PIDS="$PIDS $!"
+sleep 1
+while ! wmctrl -i -r $(wmctrl -l | grep -v grep | grep litrepl-demo | awk '{print $1}') -e 0,100,100,546,382 ; do echo -n . ; sleep 0.5 ; done
+# peek &
+peek --no-headerbar &
+sleep 1
+off=`expr 372 - 346 + 6`
+h2=`expr 100 + $off`
+while ! wmctrl -i -r $(wmctrl -l | grep -v grep | grep Peek | awk '{print $1}') -e 0,100,$h2,546,374 ; do echo -n . ; sleep 0.5 ; done
+PIDS="$PIDS $!"
+wait $PIDS

### 2024-07-19 python/litrepl/base.py
Handle invalid interpreter nicely

@@ -28,7 +28,7 @@ from .types import (PrepInfo, RunResult, NSec, FileName, SecRec,
                     FileNames, IType, Settings, CursorPos, ReadResult, SType,
                     LitreplArgs)
 from .eval import (process, pstderr, rresultLoad, rresultSave, processAdapt,
-                   processCont, interpExitCode)
+                   processCont, interpExitCode, readipid)
 from .utils import(unindent, indent, escape, fillspaces, fmterror,
                    cursor_within, nlines, wraplong, blind_unlink)
 
@@ -94,7 +94,10 @@ def settings(fns:FileNames)->Optional[Settings]:
   """ Determines the session settings. Currently just finds out the type of the
   interpreter. """
   try:
-    pid=int(open(fns.pidf).read())
+    pid=readipid(fns)
+    if pid is None:
+      pdebug(f"could not determine pid of an interpreter")
+      return None
     p=Process(pid)
     cmd=p.cmdline()
     itype=None
@@ -111,9 +114,6 @@ def settings(fns:FileNames)->Optional[Settings]:
       assert False, f"Unknown interpreter {cmd}"
     pdebug(f"interpreter pid {pid} cmd '{cmd}' leads to type '{itype}'")
     return Settings(itype,pattern1,pattern2)
-  except FileNotFoundError:
-    pdebug(f"could not determine pid of an interpreter")
-    return None
   except NoSuchProcess as p:
     pdebug(f"could not determine process of an interpreter ({p})")
     return None
@@ -507,8 +507,12 @@ def eval_section_(a:LitreplArgs, tree, secrec:SecRec, interrupt:bool=False)->int
     if not ss:
       return (fns,None)
     if interrupt:
-      pdebug("Sending SIGINT to the interpreter")
-      os.kill(int(open(fns.pidf).read()),SIGINT)
+      ipid=readipid(fns)
+      if ipid is not None:
+        pdebug("Sending SIGINT to the interpreter")
+        os.kill(ipid,SIGINT)
+      else:
+        pdebug("Failed to determine interpreter pid, not sending SIGINT")
     return fns,ss
 
   def _checkecode(fns,pending:bool):

### 2024-07-19 python/litrepl/eval.py

@@ -54,17 +54,26 @@ def with_sigmask(signals=None):
     if old is not None:
       pthread_sigmask(SIG_SETMASK,old)
 
+def readipid(fns:FileNames)->Optional[int]:
+  try:
+    return int(open(fns.pidf).read())
+  except (ValueError,FileNotFoundError):
+    return None
+
 @contextmanager
 def with_sigint(a:LitreplArgs, fns:FileNames):
   def _handler(signum,frame):
     pdebug(f"Sending SIGINT to {ipid}")
     os.kill(ipid,SIGINT)
-  ipid=int(open(fns.pidf).read())
+  ipid=readipid(fns)
   prev=None
   try:
     with with_sigmask():
-      if a.propagate_sigint:
-        prev=signal(SIGINT,_handler)
+      if ipid is not None:
+        if a.propagate_sigint:
+          prev=signal(SIGINT,_handler)
+      else:
+        pdebug(f"Failed to read pid: not installing SIGINT handler")
     yield
   finally:
     with with_sigmask():
@@ -226,10 +235,10 @@ def process(a:LitreplArgs,fns:FileNames, ss:Settings, lines:str)->Tuple[str,RunR
 
 def interpIsRunning(fns:FileNames)->bool:
   try:
-    ipid=int(open(fns.pidf).read())
+    ipid=readipid(fns)
+    if ipid is None:
+      return False
     os.kill(ipid, 0)
-  except FileNotFoundError:
-    return False
   except OSError as err:
     if err.errno == ESRCH:
       return False
@@ -248,7 +257,7 @@ def interpExitCode(fns:FileNames,poll_sec=0.5,poll_attempts=4,undefined=-1)->Opt
     else:
       try:
         return int(open(fns.ecodef).read())
-      except FileNotFoundError:
+      except (ValueError,FileNotFoundError):
         pass
     poll_attempts-=1
     if poll_attempts<=0:
@@ -303,7 +312,7 @@ def processAsync(fns:FileNames, ss:Settings, code:str)->RunResult:
   """ Send `code` to the interpreter and fork the response reader. The output
   file is locked and its name is saved into the resulting `RunResult` object.
   """
-  wd,inp,outp,pidf,_=astuple(fns)
+  wd,inp,outp,_,_=astuple(fns)
   codehash=abs(hash(code))
   fname=join(wd,f"litrepl_eval_{codehash}.txt")
   pdebug(f"processAsync starting via {fname}")

### 2024-07-19 sh/test.sh

@@ -807,6 +807,19 @@ cat out1.md | runlitrepl \
 grep -q 'KeyboardInterrupt' out2.md
 )} #}}}
 
+test_invalid_interpreter() {( #{{{
+mktest "_test_invalid_interpreter"
+cat >source.md <<"EOF"
+```python
+3+2
+```
+```result
+```
+EOF
+cat source.md | runlitrepl --python-interpreter=non-existent-ipython eval-sections >out1.md
+grep -q 'Interpreter exited with code: 127' out1.md
+)} #}}}
+
 die() {
   echo "$@" >&2
   exit 1
@@ -839,6 +852,7 @@ tests() {
   echo test_foreground
   echo test_status
   echo test_interrupt
+  echo test_invalid_interpreter
 }
 
 runlitrepl() {

### 2024-07-19 doc/example.tex
Update README.md and examples
Update examples

@@
 \documentclass{article}
+\usepackage{verbatim}
+\usepackage{mdframed}
+\usepackage{minted}
+\renewcommand{\MintedPygmentize}{pygmentize}
 \usepackage[utf8]{inputenc}
+
+% LitREPL-compatible environment for code snippets
+\newenvironment{python}
+  {\VerbatimEnvironment
+   \begin{minted}[breaklines,fontsize=\footnotesize]{python}}
+  {\end{minted}}
+\BeforeBeginEnvironment{python}{
+  \begin{mdframed}[nobreak=true,frametitle=\tiny{Python}]}
+\AfterEndEnvironment{python}{\end{mdframed}}
+
+% LitREPL-compatible environment for code results
+\newenvironment{result}{\verbatim}{\endverbatim}
+\BeforeBeginEnvironment{result}{
+  \begin{mdframed}[frametitle=\tiny{Result}]\footnotesize}
+\AfterEndEnvironment{result}{\end{mdframed}}
+
+% LitREPL-compatible command for inline code results
+\newcommand{\linline}[2]{#2}
+
 \begin{document}
 
-LitREPL treats \texttt{lcode} environment as code sections and \texttt{lresult}
-environment as result sections. The names are currently hardcoded into the
-simplified LitREPL parser. Wrapping it in other tags is not allowed.
+\section{Preamble}
 
-LaTeX does not know anything about these environments by default, we need to
-introduce them to be able to compile the document using e.g. \texttt{pdflatex}.
+Litrepl uses its own internal parser to determine LaTeX-looking tags defining
+code and result sections. As one option, it expects \texttt{python} environment
+to mark code and \texttt{result} wrapping results. Both tags need to be
+introduced to LaTeX so it doesn't get confused. The following preamble sets both
+environments to be rendered as framed boxes of fixed-width text with proper
+highlighting:
 
-\newenvironment{lcode}{\begin{texttt}}{\end{texttt}}
-\newenvironment{lresult}{\begin{texttt}}{\end{texttt}}
-\newcommand{\linline}[2]{#2}
+\begin{verbatim}
+\newenvironment{python}
+  {\VerbatimEnvironment
+   \begin{minted}[breaklines,fontsize=\footnotesize]{python}}
+  {\end{minted}}
+\BeforeBeginEnvironment{python}{
+  \begin{mdframed}[nobreak=true,frametitle=\tiny{Python}]}
+\AfterEndEnvironment{python}{\end{mdframed}}
 
-Executable section is the text between the \texttt{lcode} begin/end tags.
+\newenvironment{result}{\verbatim}{\endverbatim}
+\BeforeBeginEnvironment{result}{
+  \begin{mdframed}[frametitle=\tiny{Result}]\footnotesize}
+\AfterEndEnvironment{result}{\end{mdframed}}
+\end{verbatim}
 
-\begin{lcode}
+\section{Basic evaluation}
+
+\begin{verbatim}
+\begin{python}
 W='Hello, World!'
 print(W)
-\end{lcode}
+\end{python}
+\end{verbatim}
 
-Putting the cursor on it and typing the \texttt{:LitEval1} runs the code in the
+\begin{python}
+W='Hello, World!'
+print(W)
+\end{python}
+
+Putting the cursor on it and typing the \texttt{:LEval} runs the code in the
 background Python interpreter.
 
-\texttt{lresult} begin/end tags mark the result section.  LitREPL replaces its
+\texttt{result} begin/end tags mark the result section.  LitREPL replaces its
 content with the above code section's execution result.
 
-\begin{lresult}
+\begin{verbatim}
+\begin{result}
 Hello, World!
-\end{lresult}
+\end{result}
+\end{verbatim}
 
-LitREPL recognizes \texttt{l[no]code}/\texttt{l[no]result} comments as
-code/result section markers. This way we can use Python to produce LaTeX markup
-as output.
+\begin{result}
+Hello, World!
+\end{result}
 
-%lcode
-print("Hi!")
-%lnocode
+\section{Producing LaTeX}
 
-%lresult
-Hi!
-%lnoresult
+LitREPL also recognizes \texttt{result}/\texttt{noresult} comments as result
+section markers. With their help we can directly produce LaTeX markup as output:
+
+\begin{verbatim}
+\begin{python}
+print("\\textbf{Hi!}")
+\end{python}
+
+%result
+\textbf{Hi!}
+%noresult
+\end{verbatim}
+
+\begin{python}
+print("\\textbf{Hi!}")
+\end{python}
+
+%result
+\textbf{Hi!}
+%noresult
+
+\section{Inline output}
 
@@
 arguement is treaten as a Python printable expression. The second arguemnt is to
-be replaced with the printing output. In our simplified definition, we simply
-ignore the first argument and paste the second to the LaTeX processor as-is.
+be replaced with its value.
 
-\linline{W}{Hello, World!}
+The value of \texttt{W} is happen to be: \linline{W}{Hello, World!}
 

### 2024-07-19 doc/example.tex
Update examples

@@
 
-Litrepl uses its own internal parser to determine LaTeX-looking tags defining
-code and result sections. As one option, it expects \texttt{python} environment
-to mark code and \texttt{result} wrapping results. Both tags need to be
-introduced to LaTeX so it doesn't get confused. The following preamble sets both
-environments to be rendered as framed boxes of fixed-width text with proper
+Literpl employs its own internal parser to identify LaTeX-like tags that define
+code and result sections. One option is to use the \texttt{python} environment
+to mark code and wrap results with \texttt{result}. Both environments need to be
+defined in LaTeX so it doesn't get confused. The following preamble sets up both
+environments to render as framed boxes of fixed-width text with proper
 highlighting:
@@
 
+Here is how we can use the environments we just introduced.
+
 \begin{verbatim}
@@
 \end{python}
-
-Putting the cursor on it and typing the \texttt{:LEval} runs the code in the
-background Python interpreter.
-
-\texttt{result} begin/end tags mark the result section.  LitREPL replaces its
-content with the above code section's execution result.
 
@@
 
-LitREPL also recognizes \texttt{result}/\texttt{noresult} comments as result
-section markers. With their help we can directly produce LaTeX markup as output:
+Literpl also recognizes \texttt{result} and \texttt{noresult} comments, which
+serve as markers for result sections. By using these comments, we can generate
+LaTeX markup directly as output.
 
@@
 
-Additionally, VimREPL recognises \texttt{linline} 2-argument tags. The first
-arguement is treaten as a Python printable expression. The second arguemnt is to
-be replaced with its value.
+Furthermore, Literpl recognizes \texttt{linline} tags with two arguments. The
+first argument is treated as a Python printable expression, while the second
+argument will be replaced with its evaluated value.
 

### 2024-07-19 python/litrepl/base.py
Cereful try-finally handling; Version 3.3.2

@@ -28,7 +28,7 @@ from .types import (PrepInfo, RunResult, NSec, FileName, SecRec,
                     FileNames, IType, Settings, CursorPos, ReadResult, SType,
                     LitreplArgs)
 from .eval import (process, pstderr, rresultLoad, rresultSave, processAdapt,
-                   processCont, interpExitCode, readipid)
+                   processCont, interpExitCode, readipid, with_parent_finally)
 from .utils import(unindent, indent, escape, fillspaces, fmterror,
                    cursor_within, nlines, wraplong, blind_unlink)
 
@@ -38,15 +38,6 @@ def pdebug(*args,**kwargs):
   if DEBUG:
     print(f"[{time():14.3f},{getpid()}]", *args, file=sys.stderr, **kwargs, flush=True)
 
-PID:int=getpid()
-
-@contextmanager
-def with_parent_finally(_handler):
-  try:
-    yield
-  finally:
-    if PID==getpid():
-      _handler()
 
 def defauxdir(suffix:Optional[str]=None)->str:
   """ Generate the default name of working directory. """
@@ -592,17 +583,16 @@ def eval_section_(a:LitreplArgs, tree, secrec:SecRec, interrupt:bool=False)->int
       self._print(f"{bmarker}{tree.children[1].children[0].value}{emarker}")
 
   def _finally():
-    if PID==getpid():
-      if a.map_cursor:
-        cl=a.map_cursor[0] # cursor line
-        for threshold,diff in sorted(ledder.items()):
-          if cl>threshold:
-            cl=max(threshold,cl+diff)
-        with open(a.map_cursor_output,"w") as f:
-          f.write(str(cl))
-      if a.foreground:
-        for st in stypes:
-          stop(a,st)
+    if a.map_cursor:
+      cl=a.map_cursor[0] # cursor line
+      for threshold,diff in sorted(ledder.items()):
+        if cl>threshold:
+          cl=max(threshold,cl+diff)
+      with open(a.map_cursor_output,"w") as f:
+        f.write(str(cl))
+    if a.foreground:
+      for st in stypes:
+        stop(a,st)
 
   with with_parent_finally(_finally):
     C().visit(tree)

### 2024-07-19 python/litrepl/eval.py

@@ -40,6 +40,15 @@ def pusererror(fname,err)->None:
 SIGMASK_NESTED=False
 
 @contextmanager
+def with_parent_finally(_handler):
+  pid:int=getpid()
+  try:
+    yield # Might for a child process
+  finally:
+    if pid==getpid():
+      _handler()
+
+@contextmanager
 def with_sigmask(signals=None):
   global SIGMASK_NESTED
   assert not SIGMASK_NESTED
@@ -67,7 +76,11 @@ def with_sigint(a:LitreplArgs, fns:FileNames):
     os.kill(ipid,SIGINT)
   ipid=readipid(fns)
   prev=None
-  try:
+  def _finally():
+    with with_sigmask():
+      if prev is not None:
+        signal(SIGINT,prev)
+  with with_parent_finally(_finally):
     with with_sigmask():
       if ipid is not None:
         if a.propagate_sigint:
@@ -75,10 +88,6 @@ def with_sigint(a:LitreplArgs, fns:FileNames):
       else:
         pdebug(f"Failed to read pid: not installing SIGINT handler")
     yield
-  finally:
-    with with_sigmask():
-      if prev is not None:
-        signal(SIGINT,prev)
 
 @contextmanager
 def with_alarm(timeout_sec:float):
@@ -89,20 +98,21 @@ def with_alarm(timeout_sec:float):
     raise TimeoutError()
 
   prev=None
-  try:
-    if timeout_sec>0 and timeout_sec<float('inf'):
-      with with_sigmask():
-        prev=signal(SIGALRM,_handler)
-        setitimer(ITIMER_REAL,timeout_sec)
-      # pdebug(f"Alarm {timeout_sec} set")
-    yield
-  finally:
+  def _finally():
     if timeout_sec>0 and timeout_sec<float('inf'):
       # pdebug(f"Alarm {timeout_sec} cleaning")
       with with_sigmask():
         if prev is not None:
           setitimer(ITIMER_REAL,0)
           signal(SIGALRM,prev)
+
+  with with_parent_finally(_finally):
+    if timeout_sec>0 and timeout_sec<float('inf'):
+      with with_sigmask():
+        prev=signal(SIGALRM,_handler)
+        setitimer(ITIMER_REAL,timeout_sec)
+      # pdebug(f"Alarm {timeout_sec} set")
+    yield
 
 
 def merge_basic2(acc,r,x)->Tuple[bytes,int]:
[…]

### 2024-07-20 python/litrepl/base.py
Fix broken pipe vs exit 127 race; Version 3.3.3

@@ -26,7 +26,7 @@ from contextlib import contextmanager
 
 from .types import (PrepInfo, RunResult, NSec, FileName, SecRec,
                     FileNames, IType, Settings, CursorPos, ReadResult, SType,
-                    LitreplArgs)
+                    LitreplArgs, ECode, ECODE_OK, ECODE_RUNNING)
 from .eval import (process, pstderr, rresultLoad, rresultSave, processAdapt,
                    processCont, interpExitCode, readipid, with_parent_finally)
 from .utils import(unindent, indent, escape, fillspaces, fmterror,
@@ -478,15 +478,15 @@ def eval_code_(a:LitreplArgs,
   res=rresultSave(pptext,runr) if rr.timeout else pptext
   return res,rr
 
-def eval_section_(a:LitreplArgs, tree, secrec:SecRec, interrupt:bool=False)->int:
+def eval_section_(a:LitreplArgs, tree, secrec:SecRec, interrupt:bool=False)->ECode:
   """ Evaluate code sections of the parsed `tree`, as specified in the `secrec`
   request.  """
   nsecs=secrec.nsecs
-  ssrc:Dict[int,str]={}    # Section sources
-  sres:Dict[int,str]={}    # Section results
-  ledder:Dict[int,int]={}  # Facility to restore the cursor
-  stypes:Set[SType]=set()  # Section types we have run
-  ecode:int=0              # Combined exit code
+  ssrc:Dict[int,str]={}     # Section sources: sec.num -> code
+  sres:Dict[int,str]={}     # Section results: sec.num -> result
+  ledder:Dict[int,int]={}   # Facility to restore the cursor: line -> offset
+  ecodes:Dict[int,ECode]={} # Exit codes: sec.num -> exitcode
+  stypes:Set[SType]=set()   # Section types we have run
 
   def _getinterp(bmarker:str)->Tuple[FileNames,Optional[Settings]]:
     st=bmarker2st(bmarker)
@@ -506,19 +506,19 @@ def eval_section_(a:LitreplArgs, tree, secrec:SecRec, interrupt:bool=False)->int
         pdebug("Failed to determine interpreter pid, not sending SIGINT")
     return fns,ss
 
-  def _checkecode(fns,pending:bool):
-    nonlocal ecode
-    if ecode==0:
-      ec=interpExitCode(fns,undefined=200)
-      pdebug(f"interpreter exit code: {ec}")
[…]

### 2024-07-20 python/litrepl/eval.py

@@ -19,9 +19,11 @@ from functools import partial
 from argparse import ArgumentParser
 from contextlib import contextmanager
 from errno import ESRCH
-from signal import pthread_sigmask, valid_signals, SIG_BLOCK, SIG_UNBLOCK, SIG_SETMASK
+from signal import (pthread_sigmask, valid_signals, SIG_BLOCK, SIG_UNBLOCK,
+                    SIG_SETMASK)
 
-from .types import LitreplArgs, Settings, RunResult, ReadResult, FileNames
+from .types import (LitreplArgs, Settings, RunResult, ReadResult, FileNames,
+                    ECode, ECODE_OK, ECODE_RUNNING, ECODE_UNDEFINED)
 from .utils import blind_unlink
 
 def pstderr(*args,**kwargs):
@@ -254,16 +256,16 @@ def interpIsRunning(fns:FileNames)->bool:
       return False
   return True
 
-def interpExitCode(fns:FileNames,poll_sec=0.5,poll_attempts=4,undefined=-1)->Optional[int]:
-  """ Returns:
-    * <int>: interpreter exited with this exit code
-    * None: interpreter is still running
-    * undefined: unable to determine the code
-  """
+def interpExitCode(fns:FileNames,
+                   poll_sec=0.5,
+                   poll_attempts=4,
+                   undefined=255)->ECode:
+  """ Determines retcode of the interpreter. Polls it a little if it looks
+  crashed."""
   ecode:int|None=None
   while ecode is None:
     if interpIsRunning(fns):
-      return None
+      return ECODE_RUNNING
     else:
       try:
         return int(open(fns.ecodef).read())
@@ -273,7 +275,7 @@ def interpExitCode(fns:FileNames,poll_sec=0.5,poll_attempts=4,undefined=-1)->Opt
     if poll_attempts<=0:
       break
     sleep(poll_sec)
-  return undefined
+  return ECODE_UNDEFINED
 
 @contextmanager
 def with_fd(name:str, flags:int, open_timeout_sec=float('inf')):
@@ -348,6 +350,7 @@ def processAsync(fns:FileNames, ss:Settings, code:str)->RunResult:
                 pdebug("processAsync reader interact finish")
               except BrokenPipeError:
                 pdebug("processAsync catches Broken Pipe error")
+                os.write(fo,"<BrokenPipe>\n".encode())
                 raise
         pdebug("Exiting!")
         exit(0)

### 2024-07-20 python/litrepl/types.py

@@ -5,6 +5,12 @@ from enum import Enum
 FileName=str
 LitreplArgs=Any
 
+# Process exit code or None (running)
+ECode=Optional[int]
+ECODE_RUNNING=None
+ECODE_OK=0
+ECODE_UNDEFINED=255
+
 class SType(Enum):
   """ Code section types """
   SPython = 0

### 2024-07-20 sh/test.sh

@@ -808,6 +808,8 @@ grep -q 'KeyboardInterrupt' out2.md
 )} #}}}
 
 test_invalid_interpreter() {( #{{{
+# Exact result messages might start a race (exit code X VS broken pipe) That is
+# why we put tow sections.
 mktest "_test_invalid_interpreter"
 cat >source.md <<"EOF"
 ```python
@@ -815,8 +817,13 @@ cat >source.md <<"EOF"
 ```
 ```result
 ```
+```python
+5+6
+```
+```result
+```
 EOF
-cat source.md | runlitrepl --python-interpreter=non-existent-ipython eval-sections >out1.md
+cat source.md | runlitrepl --python-interpreter=non-existent-ipython eval-sections >out1.md || true
 grep -q 'Interpreter exited with code: 127' out1.md
 )} #}}}
 

### 2024-07-20 sh/test.sh
Update help formatting; Remove eval-section; Version 3.3.4

@@ -752,7 +752,11 @@ print(session)
 EOF
 cat source.md | runlitrepl --filetype=markdown eval-sections '0,2' >out1.md
 grep -q "^common-session" out1.md
-cat source.md | runlitrepl --foreground --filetype=markdown eval-sections '1,2' >out2.md
+cat source.md | runlitrepl \
+  --foreground \
+  --exception-exit=123 \
+  --filetype=markdown \
+  eval-sections '1,2' >out2.md
 grep -q "^foreground" out2.md
 cat source.md | runlitrepl --filetype=markdown eval-sections '2' >out3.md
 grep -q "^common-session" out3.md

### 2024-07-21 python/litrepl/base.py
Section variables draft

@@ -3,8 +3,9 @@ import sys
 import fcntl
 import re
 
-from copy import deepcopy
-from typing import List, Optional, Tuple, Set, Dict, Callable, Any
+from re import search, match as re_match, compile as re_compile
+from copy import copy, deepcopy
+from typing import List, Optional, Tuple, Set, Dict, Callable, Any, Iterable
 from select import select
 from os import environ, system, isatty, getpid, unlink, getpgid, setpgid
 from lark import Lark, Visitor, Transformer, Token, Tree
@@ -26,7 +27,7 @@ from contextlib import contextmanager
 
 from .types import (PrepInfo, RunResult, NSec, FileName, SecRec,
                     FileNames, IType, Settings, CursorPos, ReadResult, SType,
-                    LitreplArgs, ECode, ECODE_OK, ECODE_RUNNING)
+                    LitreplArgs, ECode, ECODE_OK, ECODE_RUNNING, SECVAR_RE)
 from .eval import (process, pstderr, rresultLoad, rresultSave, processAdapt,
                    processCont, interpExitCode, readipid, with_parent_finally)
 from .utils import(unindent, indent, escape, fillspaces, fmterror,
@@ -478,10 +479,16 @@ def eval_code_(a:LitreplArgs,
   res=rresultSave(pptext,runr) if rr.timeout else pptext
   return res,rr
 
-def eval_section_(a:LitreplArgs, tree, secrec:SecRec, interrupt:bool=False)->ECode:
-  """ Evaluate code sections of the parsed `tree`, as specified in the `secrec`
+def secvar_matches(code:str)->Iterable[Tuple[str,int]]:
+  for secvar in re.findall(SECVAR_RE,code):
+    idx=int(''.join([c for c in secvar if c.isdigit()]))
+    yield str(secvar),idx
+
+
+def eval_section_(a:LitreplArgs, tree, sr:SecRec, interrupt:bool=False)->ECode:
+  """ Evaluate code sections of the parsed `tree`, as specified in the `sr`
   request.  """
-  nsecs=secrec.nsecs
+  nsecs=sr.nsecs
   ssrc:Dict[int,str]={}     # Section sources: sec.num -> code
   sres:Dict[int,str]={}     # Section results: sec.num -> result
   ledder:Dict[int,int]={}   # Facility to restore the cursor: line -> offset
[…]

### 2024-07-21 python/litrepl/eval.py

@@ -6,7 +6,7 @@ import fcntl
 from fcntl import LOCK_NB,LOCK_UN,LOCK_EX
 from copy import deepcopy
 from typing import List, Optional, Tuple, Set, Dict, Callable
-from re import search, match as re_match
+from re import search, match as re_match, compile as re_compile
 from select import select
 from os import environ, system, getpid, unlink
 from lark import Lark, Visitor, Transformer, Token, Tree
@@ -207,7 +207,7 @@ TIMEOUT_SEC=3
 
 def mkre(prompt:str):
   """ Matches the shortest string followed by a sequence of prompts """
-  return re.compile(f"((.(?!{prompt}))*(.(?={prompt}))?)({prompt})+".encode('utf-8'),
+  return re_compile(f"((.(?!{prompt}))*(.(?={prompt}))?)({prompt})+".encode('utf-8'),
                     re.MULTILINE|re.DOTALL)
 
 
@@ -410,7 +410,7 @@ def processAdapt(a:LitreplArgs,
   rr=processCont(a,fns,ss,runr,timeout=timeout)
   return rr,runr
 
-PRESULT_RE=re.compile(r"(.*)\[BG:([a-zA-Z0-9_\/\.-]+)\]\n.*",
+PRESULT_RE=re_compile(r"(.*)\[BG:([a-zA-Z0-9_\/\.-]+)\]\n.*",
                       re.A|re.MULTILINE|re.DOTALL)
 
 def rresultLoad(text:str)->Tuple[str,Optional[RunResult]]:

### 2024-07-21 python/litrepl/types.py

@@ -1,4 +1,6 @@
+import re
 from typing import (Any, Set, List, Dict, Tuple, Callable, Optional)
+from re import compile as re_compile
 from dataclasses import dataclass
 from enum import Enum
 
@@ -43,14 +45,15 @@ CursorPos=Tuple[int,int]
 class PrepInfo:
   """ Results of the document preprocessing """
   nsec:NSec                         # Total number of code sections
-  cursors:Dict[CursorPos,NSec]      # Sections, rbesolved from the cursor position
+  cursors:Dict[CursorPos,NSec]      # Sections, resolved from cursor positions
   pending:Dict[NSec,RunResult]      # Async job markers
+  results:Dict[NSec,str]            # Results
 
 @dataclass
 class SecRec:
   """ Request for section evaluation """
   nsecs:Set[NSec]                   # Sections to evaluate
-  pending:Dict[NSec,RunResult]      # Contexts of already running sections
+  preproc:PrepInfo                  # Results of preprocessing
 
 @dataclass
 class FileNames:
@@ -68,4 +71,6 @@ class Settings:
   itype:IType
   pattern1:Tuple[str,str]           # Request-response pair 1
   pattern2:Tuple[str,str]           # Request-response pair 2
+
+SECVAR_RE = re_compile("> *R[0-9]+ *<",flags=re.MULTILINE|re.A)
 

### 2024-07-21 python/litrepl/base.py
Define evaluation state data structure

@@ -25,9 +25,9 @@ from textwrap import dedent
 from subprocess import check_output, DEVNULL, CalledProcessError
 from contextlib import contextmanager
 
-from .types import (PrepInfo, RunResult, NSec, FileName, SecRec,
-                    FileNames, IType, Settings, CursorPos, ReadResult, SType,
-                    LitreplArgs, ECode, ECODE_OK, ECODE_RUNNING, SECVAR_RE)
+from .types import (PrepInfo, RunResult, NSec, FileName, SecRec, FileNames,
+                    IType, Settings, CursorPos, ReadResult, SType, LitreplArgs,
+                    EvalState, ECode, ECODE_OK, ECODE_RUNNING, SECVAR_RE)
 from .eval import (process, pstderr, rresultLoad, rresultSave, processAdapt,
                    processCont, interpExitCode, readipid, with_parent_finally)
 from .utils import(unindent, indent, escape, fillspaces, fmterror,
@@ -464,8 +464,8 @@ def eval_code_(a:LitreplArgs,
                ss:Settings,
                code:str,
                runr:Optional[RunResult]=None) -> Tuple[str,ReadResult]:
-  """ Start or complete the code snippet evaluation process.  `RunResult` may
-  contain the existing runner's context.
+  """ Start or complete the snippet evaluation process. `runr`
+  contains the already existing runner's context, if any.
 
   The function returns either the evaluation result or the running context
   encoded in the result for later reference.
@@ -489,15 +489,11 @@ def eval_section_(a:LitreplArgs, tree, sr:SecRec, interrupt:bool=False)->ECode:
   """ Evaluate code sections of the parsed `tree`, as specified in the `sr`
   request.  """
   nsecs=sr.nsecs
-  ssrc:Dict[int,str]={}     # Section sources: sec.num -> code
-  sres:Dict[int,str]={}     # Section results: sec.num -> result
-  ledder:Dict[int,int]={}   # Facility to restore the cursor: line -> offset
-  ecodes:Dict[int,ECode]={} # Exit codes: sec.num -> exitcode
-  stypes:Set[SType]=set()   # Section types we have run
+  es=EvalState()
 
   def _getinterp(bmarker:str)->Tuple[FileNames,Optional[Settings]]:
     st=bmarker2st(bmarker)
-    stypes.add(st)
[…]

### 2024-07-21 python/litrepl/types.py

@@ -74,3 +74,17 @@ class Settings:
 
 SECVAR_RE = re_compile("> *R[0-9]+ *<",flags=re.MULTILINE|re.A)
 
+
+@dataclass
+class EvalState:
+  """ Interpreter state, tracking evaluation of document sections """
+  sres:Dict[int,str]                # Section results: sec.num -> result
+  ledder:Dict[int,int]              # Facility to restore the cursor: line -> offset
+  ecodes:Dict[int,ECode]            # Exit codes: sec.num -> exitcode
+  stypes:Set[SType]                 # Section types we have already run
+
+  def __init__(self):
+    self.sres,self.ledder,self.ecodes,self.stypes={},{},{},set()
+
+
+

### 2024-07-21 python/litrepl/base.py
Propagate evluation state strough the code

@@ -270,34 +270,33 @@ def start(a:LitreplArgs, st:SType)->int:
     raise ValueError(f"Unsupported section type: {st}")
 
 
-def code_preprocess_ipython(a:LitreplArgs, code:str) -> str:
+def code_preprocess_ipython(a:LitreplArgs, es:EvalState, code:str) -> str:
   # IPython seems to not echo the terminating cpaste pattern into the output
   # which is good.
   paste_pattern='12341234213423'
   return (f'\n%cpaste -q -s {paste_pattern}\n{code}\n{paste_pattern}\n')
+def code_preprocess_python(a:LitreplArgs, es:EvalState, code:str) -> str:
+  return fillspaces(code, '# spaces')
+def code_preprocess_gpt4allcli(a:LitreplArgs, es:EvalState, code:str) -> str:
+  return code + "/ask\n"
 def result_postprocess_ipython(a:LitreplArgs, text:str) -> str:
   # A workaround for https://github.com/ipython/ipython/issues/13622
   r=re.compile('ERROR! Session/line number was not unique in database. '
                'History logging moved to new session [0-9]+\\n')
   return re.sub(r,'',text)
-
-def code_preprocess_python(a:LitreplArgs, code:str) -> str:
-  return fillspaces(code, '# spaces')
-def code_preprocess_gpt4allcli(a:LitreplArgs, code:str) -> str:
-  return code + "/ask\n"
 def result_postprocess_python(a:LitreplArgs, text:str) -> str:
   return text
 def result_postprocess_gpt4allcli(a:LitreplArgs, text:str) -> str:
   return text.strip()+"\n"
 
 
-def code_preprocess(a:LitreplArgs, ss:Settings, code:str) -> str:
+def code_preprocess(a:LitreplArgs, ss:Settings, es:EvalState, code:str) -> str:
   if (ss.itype is None) or ss.itype == IType.IPython:
-    return code_preprocess_ipython(a,code)
+    return code_preprocess_ipython(a,es,code)
   elif ss.itype == IType.Python:
-    return code_preprocess_python(a,code)
+    return code_preprocess_python(a,es,code)
   elif ss.itype == IType.GPT4AllCli:
-    return code_preprocess_gpt4allcli(a,code)
+    return code_preprocess_gpt4allcli(a,es,code)
   else:
     raise ValueError(fmterror(f'''
       Interpreter type {ss.itype} is not supported for pre-processing. Did you
[…]

### 2024-07-21 python/litrepl/base.py
Move code preprocessing to a gpt4all-local handler

@@ -278,6 +278,13 @@ def code_preprocess_ipython(a:LitreplArgs, es:EvalState, code:str) -> str:
 def code_preprocess_python(a:LitreplArgs, es:EvalState, code:str) -> str:
   return fillspaces(code, '# spaces')
 def code_preprocess_gpt4allcli(a:LitreplArgs, es:EvalState, code:str) -> str:
+  for secvar,ref in secvar_matches(copy(code)):
+    code=code.replace(
+      secvar,
+      es.sres.get(
+        ref,es.sr.preproc.results.get(ref,'<invalid section variable>')
+      )
+    )
   return code + "/ask\n"
 def result_postprocess_ipython(a:LitreplArgs, text:str) -> str:
   # A workaround for https://github.com/ipython/ipython/issues/13622
@@ -489,7 +496,7 @@ def eval_section_(a:LitreplArgs, tree, sr:SecRec, interrupt:bool=False)->ECode:
   """ Evaluate code sections of the parsed `tree`, as specified in the `sr`
   request.  """
   nsecs=sr.nsecs
-  es=EvalState()
+  es=EvalState(sr)
 
   def _getinterp(bmarker:str)->Tuple[FileNames,Optional[Settings]]:
     st=bmarker2st(bmarker)
@@ -543,11 +550,6 @@ def eval_section_(a:LitreplArgs, tree, sr:SecRec, interrupt:bool=False)->ECode:
       self._print(f"{bmarker}{t}{emarker}")
       bm,em=tree.children[0].meta,tree.children[2].meta
       code=unindent(bm.column-1,t)
-      for secvar,ref in secvar_matches(copy(code)):
-        code=code.replace(
-          secvar,
-          es.sres.get(ref,sr.preproc.results.get(ref,'<invalid section variable>'))
-        )
       if self.nsec in nsecs:
         fns,ss=_getinterp(bmarker)
         rr=None
@@ -767,7 +769,7 @@ def status_verbose(a:LitreplArgs,sts:List[SType],version:str)->int:
         print(f"{st2name(st)} pending section {nsec} reader: ?")
     if st==SType.SPython:
       ss=settings(fns)
-      es=EvalState()
+      es=EvalState(SecRec.empty())
       try:
         assert ss is not None
         interpreter_path=eval_code(a,fns,ss,es,

### 2024-12-27 python/litrepl/base.py
Improve tex section parsing

@@ -321,9 +321,9 @@ def grammar_(a:LitreplArgs, filetype:str)->Tuple[LarkGrammar,Symbols]:
               | resultsec -> e_ocodesection
               | ignoresec -> e_comsection
       ignoresec.2 : ignorebegin ignoretext ignoreend
-      codesec.1 : codebegin ctext codeend
+      codesec.1 : codebegin codetext codeend
                      | comcodebegin codesectext comcodeend
-      resultsec.1 : resultbegin ctext resultend
+      resultsec.1 : resultbegin resulttext resultend
                      | comresultbegin vertext comresultend
       codebegin : /{symbols_md.codebegin}/
       codeend : /{symbols_md.codeend}/
@@ -338,7 +338,8 @@ def grammar_(a:LitreplArgs, filetype:str)->Tuple[LarkGrammar,Symbols]:
       ignorebegin : /{symbols_md.ignorebegin}/
       ignoreend : /{symbols_md.ignoreend}/
       topleveltext : /(.(?!{toplevel_markers_markdown}))*./s
-      ctext : /(.(?!{symbols_md.resultend}|{symbols_md.codeend}))*./s
+      codetext : /(.(?!{symbols_md.codeend}))*./s
+      resulttext : /(.(?!{symbols_md.resultend}))*./s
       ignoretext : /(.(?!{symbols_md.ignoreend}))*./s
       vertext : /(.(?!{symbols_md.comresultend}))*./s
       codesectext : /(.(?!{symbols_md.comcodeend}))*./s
@@ -358,10 +359,10 @@ def grammar_(a:LitreplArgs, filetype:str)->Tuple[LarkGrammar,Symbols]:
               | inlinecodesec -> e_inline
               | ignoresec -> e_comment
       ignoresec.2 : ignorebegin ignoretext ignoreend
-      codesec.1 : codebegin innertext codeend
-                | comcodebegin innertext comcodeend
-      resultsec.1 : resultbegin innertext resultend
-                  | comresultbegin innertext comresultend
+      codesec.1 : codebegin codetext codeend
+                | comcodebegin comcodetext comcodeend
+      resultsec.1 : resultbegin resulttext resultend
+                  | comresultbegin comresulttext comresultend
[…]

### 2024-12-27 sh/runtests.sh

@@ -427,6 +427,11 @@ print("AB")
 \end{zzz}
 \begin{result}
 \end{result}
+%python
+print('\end{python}')
+%nopython
+%result
+%noresult
 EOF
 cat source.tex | runlitrepl --filetype=latex parse-print >out.tex
 diff -u source.tex out.tex
@@ -494,6 +499,12 @@ print("AB")
 \begin{result}
 AB
 \end{result}
+%python
+print('\end{python}')
+%nopython
+%result
+\end{python}
+%noresult
 EOF
 )} #}}}
 

### 2024-12-27 env.sh
Update env

@@ -1,4 +1,5 @@
 CWD=$(pwd)
+export PROJECT_ROOT=$CWD
 export PATH=\
 $CWD/sh:\
 $CWD/python/bin:\

### 2024-12-27 sh/diffchanges.sh
Version 3.11.0

@@ -3,9 +3,9 @@
 git log --oneline | grep 'Version [0-9]\+\.[0-9]\+\.[0-9]\+' >_versions.txt
 HASH_LAST=`git rev-parse HEAD`
 HASH_PREV=`head -n 1 _versions.txt | awk '{print $1}' `
-# HASH_LAST=`head -n 1 _versions.txt | awk '{print $1}' `
-# HASH_PREV=`head -n 2 _versions.txt | tail -n 1 | awk '{print $1}' `
-git diff $HASH_PREV..$HASH_LAST -- . ':(exclude)README.md' ':(exclude)semver.txt' >_lastdiff.diff
+
+# git log $HASH_PREV..$HASH_LAST --pretty=format:"%h %s" --patch -- python ':(exclude)README.md' ':(exclude)semver.txt' >_lastdiff.diff
+git log $HASH_PREV..$HASH_LAST --pretty=format:"%h %s" -- python ':(exclude)README.md' ':(exclude)semver.txt' >_lastdiff.diff
 
 cat >_mkchangelog.ai <<"EOF"
 You task is to check the diff file and generate a text briefly describing the

### 2024-12-27 paper/paper.tex
Update paper; Add tex version

\documentclass[a4paper,12pt,twocolumn]{article}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{svg}
\usepackage{natbib}

\geometry{margin=1in}

\fancypagestyle{plain}{
  \fancyhf{}
  \lhead{Sergei Mironov}
  \rhead{Litrepl}
  \rfoot{\thepage}
}

\pagestyle{plain}

\title{Litrepl: Literate Paper Processor Promoting Transparency More Than Reproducibility}
\author{Sergei Mironov \\
        \texttt{sergei.v.mironov@proton.me}}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
Litrepl is a Python text processor for recognizing and evaluation code sections
in Markdown or LaTeX documents. Litrepl is designed as a middleware aimed at
separating text editors from the programming language interpreter management
logic thus contributing to the diversity of both. In this role, Litrepl can
become a component of modular "UNIX Way" interactive development or typesetting
environments.
\end{abstract}

\section{Statement of Need}

The concept of Literate Programming, initially proposed by Donald Knuth,
revolves around explaining to human beings what we want a computer to do. This
approach is embodied in tools like WEB, CWEB, and NOWEB, which focus on
non-interactive, ahead-of-time compilation. The process includes two primary
modes of operation: weaving, which creates human-readable documentation, and
tangling, which generates machine-executable code. Over time, the concept has
evolved, showing a trend towards simplification \cite{Knuth1984lp,
Ramsey1994lps}.

Concurrently, another concept of human-computer interaction called the
Read-Evaluate-Print Loop (REPL) gained traction, notably within the LISP
community and through the APL language designed for mathematical computations.
The combination of a command-line interface and a language interpreter enables
incremental and interactive programming, allowing users to directly modify the
interpreter state. By maintaining human involvement in the loop, this approach
facilitates human thought processes \cite{McCarthy1959recfun, Iverson1962apl}.

A significant milestone in this area was the development of the IPython project,
which later evolved into the Jupyter Project. The creators introduced a new
document format called the Notebook, characterized by a series of logical
sections of various types, including text and code, which could directly
interact with programming language interpreters. This interactive communication,
akin to REPL style programming, allows the creation of well-structured documents
suitable for presentations and sharing. The concept underpinning these
developments is termed Literate Computing, which includes goals of spanning a
wide audience range, boosting reproducibility, and fostering collaboration. To
achieve these objectives, several technical decisions were made, notably the
introduction of bidirectional communication—between the computational core,
known as the Jupyter Kernel, and the Notebook acting as a client, along with
another layer of client-server communication between the Notebook web server and
the user’s web browser \cite{Perez2007IPython, Granger2021litcomp,
Kluyver2016jupnb}.

4. While we recognize the importance of various goals within the Literate
Computing framework, we argue that reproducibility is paramount. Successfully
addressing reproducibility would alone suffice to enhance communication over
time among research project participants and significantly expand the audience.
However, as it became clear (\cite{Dolstra2010, Vallet2022}), this challenge
extends far beyond the scope of a human-computer interaction system, and even
beyond the typical boundaries of software distribution management for a
particular programming language. A comprehensive solution to the software
deployment problem should operate at the entire operating system level.

5. Consequently, we propose focusing on transparency in human-computer
interaction rather than on reproducibility. We present "Litrepl," a tool
designed to integrate the REPL interactive programming style into existing
editors in a straightforward yet potent manner.

6. First, we utilize simple bidirectional text streams for inter-process
communication with language interpreters to evaluate code. Second, we advocate
for the reuse of existing text document formats. In both the Markdown and LaTeX
formats that we have implemented, simplified parsers are utilized to distinguish
code and result sections from the rest of the document. As of now, we support
Python and Shell interpreter families, as well as a custom AI communication
interpreter. Finally, we strive to leverage operating system facilities as much
as possible.

\section{How it works}

\begin{figure*}[t]
  \centering
  \includesvg[inkscapelatex=false,width=0.8\textwidth]{pic}
  \caption{Caption describing the image shown in the figure.}
  \label{fig:example}
\end{figure*}

\subsection{Inter-process communication}

When the file appears on the Litrepl's stdin stream, it uses Earley parser to
reliably recognize code and result sections.

\subsection{Parsing and Evaluation}

Details and discussions about the second subsection.

\subsection{Session Management}

Details and discussions about the second subsection.

\section{Conclusion}

The tool is implemented in Python in about 2K lines of code according to the LOC
metric, and has only two Python dependencies so far, at the cost of the
dependency on the operating system intefaces for which we choose POSIX as a
wide-spread openly available standard.

\section*{References}
\bibliographystyle{plain} % Set the bibliography style
\bibliography{paper} % Include the .bib file named 'paper'

\end{document}

### 2024-12-28 python/litrepl/main.py
Make eval-sections the default command

@@ -126,18 +126,20 @@ def make_parser():
   sps.add_parser('parse-print',
     help=dedent('''Parse and print the input file back (diagnostics).'''))
   evalsec=sps.add_parser('eval-sections',
-    help=dedent('''Parse stdin, evaluate the sepcified sections (by default - all
-    available sections), print the resulting file to stdout.'''))
-  evalsec.add_argument('locs',type=str,metavar='LOCS',default='0..$',help=LOCSHELP,nargs='?')
+    help=dedent('''Parse stdin, evaluate the specified sections (by default -
+    all available sections), print the resulting file to stdout.'''))
+  evalsec.add_argument('locs',type=str,metavar='LOCS',default='0..$',
+    help=LOCSHELP,nargs='?')
   _with_type(sps.add_parser('eval-code', help='Evaluate the code snippet.'))
-  _with_type(sps.add_parser('repl', help='Connect to the background terminal using GNU socat.'))
+  _with_type(sps.add_parser('repl',
+    help='Connect to the background terminal using GNU socat.'))
   interrupt=sps.add_parser('interrupt',
     help='Send SIGINT to the background interpreter.')
   interrupt.add_argument('locs',metavar='LOCS',default='0..$',help=LOCSHELP,nargs='?')
   regexp=sps.add_parser('print-regexp',
     help='Print regexp matching start of code sections for the given file type.')
-  regexp.add_argument('format',metavar='STR',default='vim',
-    help=dedent('''Regexp format to print: 'vim' or 'lark'. Defaults to 'vim'''),nargs='?')
+  regexp.add_argument('format',metavar='STR',default='vim',nargs='?',
+    help=dedent('''Regexp format to print: 'vim' or 'lark'. Defaults to 'vim'''))
   regexp.add_argument('--tty',action='store_true',
     help='Read intput document from stdin (required to get per-section status).')
   grammar=sps.add_parser('print-grammar',
@@ -149,7 +151,10 @@ def make_parser():
 AP=make_parser()
 
 def main(args=None):
-  a=AP.parse_args(args or sys.argv[1:])
+  args=args or sys.argv[1:]
+  a=AP.parse_args(args)
+  if a.command is None:
+    a=AP.parse_args(args+['eval-sections'])
 
   if a.debug>0:
     litrepl.eval.DEBUG=True
@@ -297,7 +302,7 @@ def main(args=None):
     g=parse_(a).grammar
     print(g)
   else:
[…]

### 2024-12-28 python/litrepl/base.py
Make section address start from 1 rahter than from 0

@@ -642,7 +642,7 @@ def solve_sloc(s:str, tree:LarkTree)->SecRec:
       return tree[0]
     def s_const_num(self,tree):
       self.q+=1
-      nknown[self.q]=int(tree[0].children[0].value)
+      nknown[self.q]=int(tree[0].children[0].value)-1
       return int(self.q)
     def s_const_last(self,tree):
       return int(lastq)

### 2024-12-28 sh/runtests.sh

@@ -438,7 +438,7 @@ diff -u source.tex out.tex
 cat source.tex | runlitrepl \
   --python-markers=python,zzz \
   --filetype=latex \
-  eval-sections '0..$' >out.tex
+  eval-sections '1..$' >out.tex
 diff -u out.tex - <<"EOF"
 % == Ignore non-tags ==
 \newcommand{\linline}[2]{#2}
@@ -525,7 +525,7 @@ EOF
 cat source.tex | runlitrepl --filetype=latex parse
 cat source.tex | runlitrepl --filetype=latex parse-print >out.tex
 diff -u source.tex out.tex
-cat source.tex | runlitrepl --filetype=latex eval-sections '0..$' >out.tex
+cat source.tex | runlitrepl --filetype=latex eval-sections '1..$' >out.tex
 diff -u out.tex - <<"EOF"
 \linline{"A"+"B"}
 {AB}\linline{"C"+"D"}{CD}
@@ -551,7 +551,7 @@ for i in tqdm(range(100)):
 EOF
 cat source.md | runlitrepl --filetype=markdown parse-print >out.md
 diff -u source.md out.md
-cat source.md | runlitrepl --filetype=markdown eval-sections '0..$' >out.md
+cat source.md | runlitrepl --filetype=markdown eval-sections '1..$' >out.md
 test "$(cat out.md | grep '100%' | wc -l)" = "1"
 runlitrepl stop
 )} #}}}
@@ -699,7 +699,7 @@ print("after-exit")
 EOF
 (
 set +e
-cat source.md | runlitrepl --filetype=markdown eval-sections '0..$' >out.md 2>/dev/null
+cat source.md | runlitrepl --filetype=markdown eval-sections '1..$' >out.md 2>/dev/null
 echo $?>ret.txt
 )
 test `cat ret.txt` = "123"
@@ -731,7 +731,7 @@ print("after-exception")
 EOF
 (
 set +e
-cat source.md | runlitrepl --filetype=markdown eval-sections '0..$' >out.md 2>/dev/null
+cat source.md | runlitrepl --filetype=markdown eval-sections '1..$' >out.md 2>/dev/null
 echo $?>ret.txt
 )
 test `cat ret.txt` = "123"
@@ -756,7 +756,7 @@ _=system("echo 2")
 ``` result
 ```
 EOF
-cat in.md | runlitrepl --debug=0 --filetype=markdown eval-sections '0..$' >out.md
[…]

### 2024-12-28 paper/paper.tex
Update env; Update paper

@@
 \usepackage{svg}
-\usepackage{natbib}
+
+\usepackage{biblatex}
+\addbibresource{paper.bib}
 
@@
 }
+
+\newcommand{\ls}{\begin{itemize}\item}
+\newcommand{\li}{\item}
+\renewcommand{\le}{\end{itemize}}
+
+\newcommand{\es}{\begin{enumerate}\item}
+\newcommand{\ei}{\item}
+\newcommand{\ee}{\end{enumerate}}
 
@@
 interpreter state. By maintaining human involvement in the loop, this approach
-facilitates human thought processes \cite{McCarthy1959recfun, Iverson1962apl}.
+facilitates human thought processes \cite{Spence1975apl, McCarthy1959recfun,
+Iverson1962apl}.
 
@@
 Python and Shell interpreter families, as well as a custom AI communication
-interpreter. Finally, we strive to leverage operating system facilities as much
-as possible.
+interpreter. Finally, we strive to leverage POSIX \cite{POSIX2024} operating
+system facilities as much as possible.
 
 \section{How it works}
+
+Litrepl is implemented in Python as a command-line text processor utility. Its
+primary function is to take a text document as input, process it according to
+specified command line arguments and environment settings, and then output the
+resulting document through its standard output.
+
+The operation of Litrepl is best illustrated through the following example:
+
+\begin{verbatim}
+$ cat input.tex
+\begin{python}
+print("Hello")
+\end{python}
+\begin{result}
+\end{result}
+
+$ cat input.tex | litrepl
+\begin{python}
+print("Hello")
+\end{python}
+\begin{result}
+Hello
+\end{result}
+\end{verbatim}
+
+In this example, the input document \verb|input.tex| contains a Python code
+block that outputs the string "Hello". When the document is piped through
+Litrepl, it processes the code block and inserts the result directly into the
+document where specified.
 
@@
   \caption{Caption describing the image shown in the figure.}
-  \label{fig:example}
+  \label{fig:resource-allocation}
 \end{figure*}
 
-\subsection{Inter-process communication}
+\subsection{Interfacing Interpreters}
 
-When the file appears on the Litrepl's stdin stream, it uses Earley parser to
-reliably recognize code and result sections.
+Litrepl communicates with interpreters using two uni-directional text streams:
+one for writing input and another for reading outputs. To establish effective
+communication, the interpreter should conform to the following general
+assumptions:
 
-\subsection{Parsing and Evaluation}
+\begin{itemize}
+  \item Synchronous single-user mode, which is implemented in most interpreters.
+  \item A capability to disable command line prompts. Litrepl relies on the echo
+        response, as described below, rather than on prompt detection.
+  \item The presence of an echo command or equivalent. The interpreter must be
+        able to echo an argument string provided by the user in response to the
+        echo command.
+\end{itemize}
+
+In Litrepl, these details are hardcoded for several prominent interpreter
+families, which we refer to as \textbf{interpreter classes}. At the time of
+writing, Litrepl supports three classes: \textbf{python}, \textbf{sh}, and
+\textbf{ai}. Using command line arguments, users can configure how to map code
+section labels to the correct class and specify which interpreter command to
+execute to start a session for each class.
+
+\subsection{Session Management}
 
@@
 
-\subsection{Session Management}
+1. In order to start interpreter sessions, Litrepl allocates resources as shown
+on the Figure \ref{fig:resource-allocation}. First, the auxilary directory is
+created using the name, computed from the current interpreter class, working
+directory and the name of a system folder for temporary files.  Next, the two
+POSIX pipes, in.pipe and out.pipe, are created. Finally, the configured
+interpreter is started with the pipes attached.
+
+\subsection{Parsing and Evaluation}
 
@@
 \section*{References}
-\bibliographystyle{plain} % Set the bibliography style
-\bibliography{paper} % Include the .bib file named 'paper'
+\printbibliography
 

### 2024-12-28 paper/paper.tex
Update paper

@@
 \usepackage{svg}
+\usepackage{comment}
 
@@
 
-The operation of Litrepl is best illustrated through the following example:
+The operation of Litrepl is best illustrated through the example below. Consider
+the document named \verb|input.tex|:
 
+\begin{comment}
+\begin{sh}
+echo '\begin{verbatim}'
+echo '$ cat input.tex'
+cat input.tex
+echo '\end{verbatim}'
+\end{sh}
+\end{comment}
+%lresult
 \begin{verbatim}
@@
 \begin{python}
-print("Hello")
+import sys
+print(f"Hello, {sys.platform}!")
 \end{python}
@@
 \end{result}
+\end{verbatim}
+%lnoresult
 
+This document contains a Python code section and an empty result section marked
+with the corresponding Latex environment tags. To "execute" the document we
+pipe it though the Litrepl processor as follows:
+
+\begin{comment}
+\begin{sh}
+echo '\begin{verbatim}'
+echo '$ cat input.tex | litrepl'
+echo "sys.platform='linux'" | litrepl repl python >/dev/null
+cat input.tex | litrepl
+echo '\end{verbatim}'
+\end{sh}
+\end{comment}
+%lresult
+\begin{verbatim}
 $ cat input.tex | litrepl
 \begin{python}
-print("Hello")
+import sys
+print(f"Hello, {sys.platform}!")
 \end{python}
 \begin{result}
-Hello
+Hello, linux!
 \end{result}
 \end{verbatim}
+%lnoresult
 
-In this example, the input document \verb|input.tex| contains a Python code
-block that outputs the string "Hello". When the document is piped through
-Litrepl, it processes the code block and inserts the result directly into the
-document where specified.
+Now we can see the expected greeting of the author's operating system platform.
+The side-effect of this execution is the started session of the python
+interpreter which is now running in the background. We can modify its state by
+adding more section to the document and executing them selectively or e.g. by
+accessing \verb|litrepl repl python| terminal. So for example, setting
+\verb|sys.platform| to another value and re-evaluating the document would yield
+a different greeting.
 

### 2024-12-29 python/litrepl/eval.py
Modify async tag pattern

@@ -339,7 +339,7 @@ def processAsync(fns:FileNames, ss:Interpreter, code:str)->RunResult:
   """
   wd,inp,outp,_,_,_=astuple(fns)
   codehash=abs(hash(code))
-  fname=join(wd,f"litrepl_eval_{codehash}.txt")
+  fname=join(wd,f"partial_{codehash}.txt")
   pdebug(f"processAsync starting via {fname}")
   with with_locked_fd(fname,CREATE_WRONLY_EMPTY,LOCK_NONBLOCKING) as fo:
     if fo is not None:
@@ -399,7 +399,6 @@ def processCont(fns:FileNames,
         pdebug(f"processCont final readout start")
         res=readout(fdr,prompt=mkre(p2[1]),merge=merge_rn2)
         pdebug(f"processCont final readout finish")
-        # res=res+f"\nDBG Obtained from:{runr.fname}\n"
         rr=ReadResult(res,False) # Return final result
         remove_silent(runr.fname)
         pdebug(f"processCont unlinked {runr.fname}")
@@ -429,7 +428,7 @@ def processAdapt(fns:FileNames,
   rr=processCont(fns,ss,runr,timeout=timeout,propagate_sigint=propagate_sigint)
   return rr,runr
 
-PRESULT_RE=re_compile(r"(.*)\[BG:([a-zA-Z0-9_\/\.-]+)\]\n.*",
+PRESULT_RE=re_compile(r"(.*)\[LR:([a-zA-Z0-9_\/\.-]+)\]\s*",
                       re.A|re.MULTILINE|re.DOTALL)
 
 def rresultLoad(text:str)->Tuple[str,Optional[RunResult]]:
@@ -443,7 +442,7 @@ def rresultSave(text:str, presult:RunResult)->str:
   """ Saves uncompleted RunRestult into the response text in order to load and
   check it later. """
   sep='\n' if text and text[-1]!='\n' else ''
-  return (text+f"{sep}[BG:{presult.fname}]\n")
+  return (text+f"{sep}[LR:{presult.fname}]\n")
 
 
 def interp_code_preprocess(a:LitreplArgs, ss:Interpreter, es:EvalState, code:str) -> str:

### 2024-12-29 sh/runtests.sh

@@ -573,7 +573,7 @@ EOF
 cat source.md | runlitrepl --filetype=markdown --timeout=1,inf eval-sections >out1.md
 cat out1.md | runlitrepl --filetype=markdown --timeout=inf,1 --pending-exitcode=33 eval-sections >out2.md ||
 test "$?" = "33"
-grep -q 'BG' out2.md
+grep -q 'LR' out2.md
 runlitrepl stop
 )} #}}}
 
@@ -1106,7 +1106,7 @@ cat source.md | runlitrepl \
   --filetype=markdown \
   --timeout=0,inf \
   eval-sections '1..$' >out1.md
-grep -q 'BG' out1.md
+grep -q 'LR' out1.md
 sleep 1 # IPython seems to die without this delay
 cat out1.md | runlitrepl \
   --filetype=markdown \

### 2024-12-29 python/litrepl/base.py
Modify and unify hashing algorithm

@@ -20,7 +20,6 @@ from argparse import ArgumentParser
 from collections import defaultdict
 from os import makedirs, getuid, getcwd, WEXITSTATUS, remove
 from tempfile import gettempdir
-from hashlib import sha256
 from psutil import Process, NoSuchProcess
 from textwrap import dedent
 from subprocess import check_output, DEVNULL, CalledProcessError
@@ -34,7 +33,7 @@ from .eval import (process, pstderr, rresultLoad, rresultSave, processAdapt,
                    processCont, interpExitCode, readipid, with_parent_finally,
                    with_fd, read_nonblock, eval_code, eval_code_)
 from .utils import(unindent, indent, escape, fillspaces, fmterror,
-                   cursor_within, nlines, wraplong, remove_silent)
+                   cursor_within, nlines, wraplong, remove_silent, hashdigest)
 
 from .interpreters.python import PythonInterpreter
 from .interpreters.ipython import IPythonInterpreter
@@ -64,9 +63,7 @@ def defauxdir(st:SType, suffix:Optional[str]=None)->str:
   """ Calculate the default aux. directory name. """
   suffix_=f"{suffix}_" if suffix is not None else ""
   return join(gettempdir(),
-              (f"litrepl_{getuid()}_"+suffix_+
-               sha256(getcwd().encode('utf-8')).hexdigest()[:6]),
-              st2name(st))
+              f"litrepl_{getuid()}_{suffix}{hashdigest(getcwd())}", st2name(st))
 
 def st2name(st:SType)->str:
   """ Return string representation of the code section type """

### 2024-12-29 python/litrepl/eval.py

@@ -24,7 +24,7 @@ from signal import (pthread_sigmask, valid_signals, SIG_BLOCK, SIG_UNBLOCK,
 
 from .types import (LitreplArgs, RunResult, ReadResult, FileNames, EvalState,
                     ECode, ECODE_OK, ECODE_RUNNING, ECODE_UNDEFINED)
-from .utils import remove_silent, wraplong
+from .utils import remove_silent, wraplong, hashdigest
 
 def pstderr(*args,**kwargs):
   print(*args, file=sys.stderr, **kwargs, flush=True)
@@ -338,7 +338,7 @@ def processAsync(fns:FileNames, ss:Interpreter, code:str)->RunResult:
   file is locked and its name is saved into the resulting `RunResult` object.
   """
   wd,inp,outp,_,_,_=astuple(fns)
-  codehash=abs(hash(code))
+  codehash=hashdigest(code)
   fname=join(wd,f"partial_{codehash}.txt")
   pdebug(f"processAsync starting via {fname}")
   with with_locked_fd(fname,CREATE_WRONLY_EMPTY,LOCK_NONBLOCKING) as fo:

### 2024-12-29 python/litrepl/utils.py

@@ -2,6 +2,7 @@ from textwrap import dedent, wrap
 from re import match as re_match, compile as re_compile
 from typing import Iterable,List
 from os import unlink, system
+from hashlib import sha256
 from .types import CursorPos
 
 DEBUG:bool=False
@@ -126,4 +127,7 @@ def runsocat(fns, hint=None):
     hint=SOCAT_HINT
   print(hint,end='')
   system(f"socat - 'PIPE:{fns.outp},flock-ex-nb=1!!PIPE:{fns.inp},flock-ex-nb=1'")
+
+def hashdigest(x)->str:
+  return sha256(str(hash(x)).encode('utf-8')).hexdigest()[:7]
 

### 2024-12-29 python/litrepl/eval.py
Modify async tag pattern

@@ -428,8 +428,8 @@ def processAdapt(fns:FileNames,
   rr=processCont(fns,ss,runr,timeout=timeout,propagate_sigint=propagate_sigint)
   return rr,runr
 
-PRESULT_RE=re_compile(r"(.*)\[LR:([a-zA-Z0-9_\/\.-]+)\]\s*",
-                      re.A|re.MULTILINE|re.DOTALL)
+PRESULT_RE=re_compile(r"(.*)\[LR:([a-zA-Z0-9_\/\.-]+)\]\s*$",
+                      re.A|re.DOTALL)
 
 def rresultLoad(text:str)->Tuple[str,Optional[RunResult]]:
   m=re_match(PRESULT_RE,text)

### 2024-12-29 paper/paper.tex
Update paper

@@
 
-Details and discussions about the second subsection.
+To initiate interpreter sessions, Litrepl allocates resources as depicted in
+Figure \ref{fig:resource-allocation}. The process begins by creating an
+auxiliary directory. The name of this directory is derived from the combination
+of the current interpreter class name, the Litrepl working directory, and a
+system-designated folder for temporary files. Subsequently, two POSIX
+pipes—\texttt{in.pipe} and \texttt{out.pipe} are established. The configured
+interpreter is then launched with these pipes attached, and the interpreter's
+process identifier is recorded in a dedicated file.
 
-1. In order to start interpreter sessions, Litrepl allocates resources as shown
-on the Figure \ref{fig:resource-allocation}. First, the auxilary directory is
-created using the name, computed from the current interpreter class, working
-directory and the name of a system folder for temporary files.  Next, the two
-POSIX pipes, in.pipe and out.pipe, are created. Finally, the configured
-interpreter is started with the pipes attached.
+The initialization routine incorporates additional open calls on the pipes to
+prevent premature termination of the interpreter due to detachments. This
+ensures the interpreter remains active and responsive.
+
+When an evaluation request is received, Litrepl calculates the name of the
+anticipated response storage file and forwards the request to the interpreter.
+It concurrently initiates a response reader process. The response storage is
+secured using soft POSIX file locks, enabling the main process to ascertain when
+the response is ready. In instances where the response exceeds the configured
+time to completion, Litrepl provides a partial result. This partial result can
+be updated with complete details in subsequent calls. The format for a partial
+result section appears as follows:
+% Displaying a verbatim box above two columns in a two-column paper
+
+\begin{comment}
+\begin{sh}
+echo -n '\begin' ; echo '{verbatim}'
+echo '\begin{result}'
+litrepl --python-auxdir=/tmp/litrepl/python restart python
+{
+cat <<EOF
+import time
+print('... some output ...')
+time.sleep(9999)
+EOF
+}|litrepl --python-auxdir=/tmp/litrepl/python --timeout=1,inf eval-code python
+echo '\end{result}'
+echo -n '\end'; echo '{verbatim}'
+\end{sh}
+\end{comment}
+
+\begin{figure*}[t]
+  \centering
+  \begin{minipage}{\textwidth}
+    \centering
+    %lresult
+    \begin{verbatim}
+    \begin{result}
+    ... some output ...
+    [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
+    \end{result}
+    \end{verbatim}
+    %lnoresult
+  \end{minipage}
+  \caption{Example output in a verbatim format displayed above two columns.}
+  \label{fig:verbatim_example}
+\end{figure*}
+
+% The figure* environment is used to span two columns, while a minipage is
+% used to contain the verbatim listing. This code creates a box above the
+% two-column layout with a caption explaining the contents.
 

### 2024-12-29 paper/paper.tex
Update paper

@@
 \usepackage{comment}
+\usepackage{float}
 
@@
 \section{Statement of Need}
+
+\begin{figure*}[!hbt]
+  \centering
+  \includesvg[inkscapelatex=false,width=0.8\textwidth]{pic}
+  \caption{Caption describing the image shown in the figure.}
+  \label{fig:resource-allocation}
+\end{figure*}
 
@@
 import sys
-print(f"Hello, {sys.platform}!")
+print(f"I use {sys.platform} btw!")
 \end{python}
@@
 import sys
-print(f"Hello, {sys.platform}!")
+print(f"I use {sys.platform} btw!")
 \end{python}
 \begin{result}
-Hello, linux!
+I use linux btw!
 \end{result}
@@
 
-Now we can see the expected greeting of the author's operating system platform.
-The side-effect of this execution is the started session of the python
-interpreter which is now running in the background. We can modify its state by
-adding more section to the document and executing them selectively or e.g. by
-accessing \verb|litrepl repl python| terminal. So for example, setting
-\verb|sys.platform| to another value and re-evaluating the document would yield
-a different greeting.
-
-\begin{figure*}[t]
-  \centering
-  \includesvg[inkscapelatex=false,width=0.8\textwidth]{pic}
-  \caption{Caption describing the image shown in the figure.}
-  \label{fig:resource-allocation}
-\end{figure*}
+Now we can see the expected statement about the author's operating system.  The
+side-effect of this execution is the started session of the python interpreter
+which is now running in the background. We can modify its state by adding more
+section to the document and executing them selectively or e.g. by accessing
+\verb|litrepl repl python| terminal. So for example, setting \verb|sys.platform|
+to another value and re-evaluating the document would yield a different
+statement.
 
@@
 
+\begin{comment}
+\begin{sh}
+echo -n '\begin' ; echo '{verbatim}'
+echo '\begin{result}'
+litrepl --python-auxdir=/tmp/litrepl/python restart python
+{
+cat <<EOF
+import time
+print('... some output ...')
+time.sleep(9999)
+EOF
+}|litrepl --python-auxdir=/tmp/litrepl/python --timeout=1,inf eval-code python
+echo '\end{result}'
+echo -n '\end'; echo '{verbatim}'
+\end{sh}
+\end{comment}
+
+\begin{figure*}[hbt!]
+  \centering
+  %lresult
+  \begin{verbatim}
+  \begin{result}
+  ... some output ...
+  [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
+  \end{result}
+  \end{verbatim}
+  %lnoresult
+  \caption{Example output in a verbatim format displayed above two columns.}
+  \label{fig:verbatim_example}
+\end{figure*}
+
 To initiate interpreter sessions, Litrepl allocates resources as depicted in
@@
 result section appears as follows:
-% Displaying a verbatim box above two columns in a two-column paper
-
-\begin{comment}
-\begin{sh}
-echo -n '\begin' ; echo '{verbatim}'
-echo '\begin{result}'
-litrepl --python-auxdir=/tmp/litrepl/python restart python
-{
-cat <<EOF
-import time
-print('... some output ...')
-time.sleep(9999)
-EOF
-}|litrepl --python-auxdir=/tmp/litrepl/python --timeout=1,inf eval-code python
-echo '\end{result}'
-echo -n '\end'; echo '{verbatim}'
-\end{sh}
-\end{comment}
-
-\begin{figure*}[t]
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    %lresult
-    \begin{verbatim}
-    \begin{result}
-    ... some output ...
-    [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
-    \end{result}
-    \end{verbatim}
-    %lnoresult
-  \end{minipage}
-  \caption{Example output in a verbatim format displayed above two columns.}
-  \label{fig:verbatim_example}
-\end{figure*}
-
-% The figure* environment is used to span two columns, while a minipage is
-% used to contain the verbatim listing. This code creates a box above the
-% two-column layout with a caption explaining the contents.
 
@@
 
-Details and discussions about the second subsection.
+\begin{figure*}[hbt!]
+  \centering
+  \begin{verbatim}
+  document ::= (code_section | result_section | ignore_section | text)*
+  code_section ::= code_bmarker code_section_text code_emarker
+  result_section ::= result_bmarker result_section_text result_emarker
+  ignore_section ::= ignore_bmarker ignore_section_text ignore_emarker
+  \end{verbatim}
+  \caption{Example output in a verbatim format displayed above two columns.}
+  \label{fig:document_representation}
+\end{figure*}
+
+To efficiently detect sections within a document, Litrepl employs simplified
+versions of each document markup grammar, which it defines and parses. In this
+approach, documents are abstracted as a straightforward sequence comprising
+code, result, and text sections. Additionally, Litrepl identifies ignore blocks,
+which serve as block comments. Any sections encapsulated within these ignore
+blocks are excluded from parsing and evaluation. The structure of any document,
+therefore, can be articulated using the following informal grammar:
+
+The specific markers that denote each block type are determined by the document
+format and configuration settings. For example, in a LaTeX document containing
+Python code, Litrepl recognizes markers such as \verb|\begin|\verb|{python}| or
+\verb|% |\verb|python| as code section begin markers, and
+\verb|\end|\verb|{python}| or \verb|% |\verb|nopython| as end markers. These
+latter variants are transparent to LaTeX, facilitating the inclusion of LaTeX
+markup seamlessly.
+
+This parsing strategy allows Litrepl to navigate and process documents with
+precision, ensuring that relevant sections are evaluated while non-relevant
+parts are effectively ignored. By integrating document format-specific markers,
+Litrepl maintains flexibility and adaptability in diverse document scenarios.
 

### 2024-12-29 paper/paper.tex
Update paper

@@
 \usepackage{float}
+\usepackage[font=small,skip=0pt]{caption}
 
@@
   \centering
+  \begin{minipage}{\textwidth}
   %lresult
@@
   %lnoresult
+  \end{minipage}
   \caption{Example output in a verbatim format displayed above two columns.}
@@
   \centering
+  \begin{minipage}{\textwidth}
   \begin{verbatim}
-  document ::= (code_section | result_section | ignore_section | text)*
-  code_section ::= code_bmarker code_section_text code_emarker
-  result_section ::= result_bmarker result_section_text result_emarker
-  ignore_section ::= ignore_bmarker ignore_section_text ignore_emarker
+  document       ::= (code | result | ignore | text)*
+  code           ::= (code-normal | code-comment)
+  result         ::= (result-normal | result-comment)
+  code-normal    ::= "\begin{MARKER}" text "\end{MARKER}"
+  code-comment   ::= "% MARKER" text "% noMARKER"
+  result-normal  ::= "\begin{result}" text "\end{result}"
+  result-comment ::= "% result" text "% noresult"
+  ignore         ::= "% ignore" text "% noignore"
+  text           ::= ...
   \end{verbatim}
-  \caption{Example output in a verbatim format displayed above two columns.}
+  \end{minipage}
+  \caption{An illustrative grammar template for LaTeX documents where marker
+  serves as a parameter configured via command-line arguments for each supported
+  interpreter class.}
   \label{fig:document_representation}
@@
 
-This parsing strategy allows Litrepl to navigate and process documents with
-precision, ensuring that relevant sections are evaluated while non-relevant
-parts are effectively ignored. By integrating document format-specific markers,
-Litrepl maintains flexibility and adaptability in diverse document scenarios.
+The template grammars similar to the example are hardcoded for Markdown and
+LaTeX document formats. In each run, Litrepl compiles them to the document
+parser based on its configuration options using the Lark\cite{Lark} parsing
+library.
+
+\section{Discussion}
+
+TODO
 
@@
 
-\section*{References}
 \printbibliography

### 2024-12-30 paper/paper.tex
Update paper

@@
 In Litrepl, these details are hardcoded for several prominent interpreter
-families, which we refer to as \textbf{interpreter classes}. At the time of
-writing, Litrepl supports three classes: \textbf{python}, \textbf{sh}, and
+families, which we refer to as \textit{interpreter classes}. At the time of
+writing, Litrepl supports three such classes: \textbf{python}, \textbf{sh}, and
 \textbf{ai}. Using command line arguments, users can configure how to map code
@@
   \caption{Example output in a verbatim format displayed above two columns.}
-  \label{fig:verbatim_example}
+  \label{fig:partial-result}
 \end{figure*}
 
-To initiate interpreter sessions, Litrepl allocates resources as depicted in
-Figure \ref{fig:resource-allocation}. The process begins by creating an
-auxiliary directory. The name of this directory is derived from the combination
-of the current interpreter class name, the Litrepl working directory, and a
-system-designated folder for temporary files. Subsequently, two POSIX
-pipes—\texttt{in.pipe} and \texttt{out.pipe} are established. The configured
-interpreter is then launched with these pipes attached, and the interpreter's
-process identifier is recorded in a dedicated file.
+\begin{comment}
+%lai
+Lets rewrite the "Session management" section. Please consider the following
+points, check the grammar and combine them into a linked narrative. Please
+generate it using Latex markup. Wrap your own comments in Latex line comments.
+Please don't generate figures. Just text.
 
-The initialization routine incorporates additional open calls on the pipes to
-prevent premature termination of the interpreter due to detachments. This
-ensures the interpreter remains active and responsive.
+----
 
-When an evaluation request is received, Litrepl calculates the name of the
-anticipated response storage file and forwards the request to the interpreter.
-It concurrently initiates a response reader process. The response storage is
-secured using soft POSIX file locks, enabling the main process to ascertain when
-the response is ready. In instances where the response exceeds the configured
-time to completion, Litrepl provides a partial result. This partial result can
-be updated with complete details in subsequent calls. The format for a partial
-result section appears as follows:
+1. As mentioned earlier, the necessary side-effect of code section evaluation,
+which makes read-eval loop possible - is the the interpreter session left
+running in the background. The resources required to manage this mode of
+operation are shown on the Figure \ref{fig:resource-allocation}. All of them are
+visible as files residing in some auxiliary directory.
+
+2. If not set via arguments or environment, the path to this directory
+is calculated out of interpreter class name, current working directory and the
+operating system temporary folder name.
+
+3. The code to be evaluated determines (via hashing) the name of the response file,
+containing interpreter response for this code.
+
+4. Also, aux. dir contains a couple of pipes providing an access to the
+interpreter standard input and ouput streams.
+
+5. Finally, the PID of the interpreter process is also stored in a file.
+
+6. When evaluating, Litrepl forks a response reader which keeps it soft-locked
+until the interpreter signals the completion by answering to the echo probe. If
+the response takes longer than configured, Litrepl outputs a pratial response
+result which can be read and checked for completion during next runs.
+
+7. Litrepl provides \textbf{start}, \textbf{stop}, \textbf{restart} commands
+accepting the name of interpreter class to act on to start, stop or restart
+interpreter sessions.
+
+----
+
+%lnoai
+\end{comment}
+
+%lresult
+Litrepl's ability to maintain interpreter sessions in the background is crucial
+for enabling a Read-Eval-Print Loop (REPL) environment. These resources, shown
+in Figure \ref{fig:resource-allocation}, are stored as files within an auxiliary
+directory.
+
+If not specified by command-line arguments or environment variables, the
+directory path is automatically derived from the interpreter class name, the
+current working directory, and the OS's temporary file location.
+
+The auxiliary directory includes two POSIX pipes for interpreter I/O and a file
+recording the running interpreter's process ID, aiding session management.
+
+When a code section is evaluated, Litrepl assigns a response file name derived
+from hashing the code. This response file stores the output from the
+interpreter.
+
+During evaluation, Litrepl spawns a response reader process with a soft lock,
+active until the interpreter completes and responds to an echo probe. If the
+response exceeds the configured duration, Litrepl outputs a partial result tag,
+which is recognized and reevaluated in subsequent runs. Figure
+\ref{fig:partial-result} shows an example partial result section.
+
+Litrepl provides \textbf{start}, \textbf{stop}, and \textbf{restart} commands to
+control background sessions. The \textbf{interrupt} command sends an
+interruption signal to the interpreter. Finally, the \textbf{repl} command
+establishes direct communication with the interpreter, allowing manual
+inspection of its state.
+%lnoresult
 
@@
 
+%lignore
 \begin{figure*}[hbt!]
@@
 \end{figure*}
+%lnoignore
 

### 2024-12-31 paper/paper.tex
Update paper

@@
 
-To efficiently detect sections within a document, Litrepl employs simplified
-versions of each document markup grammar, which it defines and parses. In this
-approach, documents are abstracted as a straightforward sequence comprising
-code, result, and text sections. Additionally, Litrepl identifies ignore blocks,
-which serve as block comments. Any sections encapsulated within these ignore
-blocks are excluded from parsing and evaluation. The structure of any document,
-therefore, can be articulated using the following informal grammar:
+Litrepl abstracts documents as a straightforward sequence comprising code,
+result, and text sections. Additionally, Litrepl identifies ignore blocks, which
+act as comments that prevent enclosed sections from being evaluated.
 
-The specific markers that denote each block type are determined by the document
-format and configuration settings. For example, in a LaTeX document containing
-Python code, Litrepl recognizes markers such as \verb|\begin|\verb|{python}| or
-\verb|% |\verb|python| as code section begin markers, and
-\verb|\end|\verb|{python}| or \verb|% |\verb|nopython| as end markers. These
-latter variants are transparent to LaTeX, facilitating the inclusion of LaTeX
-markup seamlessly.
-
-The template grammars similar to the example are hardcoded for Markdown and
-LaTeX document formats. In each run, Litrepl compiles them to the document
-parser based on its configuration options using the Lark\cite{Lark} parsing
-library.
+Template grammars similar to the illustrative example above are encoded for
+Markdown and LaTeX formats. Before each run, Litrepl calls Lark\cite{Lark} to
+compile a customized parser and uses it to access the sections.
 

### 2025-01-01 python/litrepl/base.py
Update lignore pattern

@@ -293,8 +293,8 @@ class SymbolsLatex(Symbols):
     self.comcodeend = comcodeend_re
     self.comresultbegin = r"\%[ ]*l?result"
     self.comresultend = r"\%[ ]*l?noresult"
-    self.ignorebegin = r"\%lignore"
-    self.ignoreend = r"\%lnoignore"
+    self.ignorebegin = r"\%[ ]*lignore"
+    self.ignoreend = r"\%[ ]*lnoignore"
     self.inlinemarker = r"\\l[a-zA-Z0-9]*inline"
 
 OBR="{"

### 2025-01-01 paper/paper.tex
Update paper

@@
 
-Litrepl is implemented in Python as a command-line text processor utility. Its
-primary function is to take a text document as input, process it according to
-specified command line arguments and environment settings, and then output the
-resulting document through its standard output.
+Litrepl is implemented as a command-line text processor utility. Its primary
+function is to take a text document as input, process it according to specified
+command line arguments and environment settings, and then output the resulting
+document through its standard output.
 
@@
 Litrepl's ability to maintain interpreter sessions in the background is crucial
-for enabling a Read-Eval-Print Loop (REPL) environment. These resources, shown
-in Figure \ref{fig:resource-allocation}, are stored as files within an auxiliary
-directory.
+for enabling a Read-Eval-Print Loop (REPL) environment. The associated
+resources, shown in Figure \ref{fig:resource-allocation}, are stored as files
+within an auxiliary directory.
 
@@
 
-%lignore
+%% ignore %%
 \begin{figure*}[hbt!]
@@
   interpreter class.}
-  \label{fig:document_representation}
+  \label{fig:document-representation}
 \end{figure*}
-%lnoignore
+%% noignore %%
 
@@
 
-Template grammars similar to the illustrative example above are encoded for
-Markdown and LaTeX formats. Before each run, Litrepl calls Lark\cite{Lark} to
-compile a customized parser and uses it to access the sections.
+Template grammars similar to the illustrative example in Figure
+\ref{fig:document-representation} are encoded for Markdown and LaTeX formats.
+Before each run, Litrepl calls Lark\cite{Lark} to compile a customized parser
+and uses it to access the sections.
+
+Evaluation results are encoded back into the result sections, and the entire
+document is printed. At this stage, certain conditions can be optionally
+checked. First, setting \verb|--pending-exitcode| to a non-zero value instructs
+Litrepl to report an error if a section takes longer than the timeout to
+evaluate. Second, setting \verb|--exception-exitcode| directs Litrepl to detect
+Python exceptions. Lastly, \verb|--irreproducible-exitcode| triggers an error if
+the evaluation result doesn't match the text initially present in the result
+section.
+
+The last option implements the only formal check for aiding reproducibility that
+Litrepl provides.
 
@@
 
-TODO
+\begin{comment}
+% lai
+
+Lets write the "Discussion" section. of the paper. Please check the grammar of
+the following points and make a linked narrative out of them. Please use Latex
+markup.
+
+---
+
+The technical decision to abstract interpreters using text streams has its
+positive and negative sides. While simplicity is the obvious bonus, the negatvie
+consequences can be groouped into (1) No parallel evaluation at the
+communication level. Th interpreter gets locked until it finishes evaluation of
+one snippet before it can proceed to another one. (2) Transferrable data type is
+limited to text-only streams.
+
+We argue that the lack of parallel execution at the communication level can be
+compensated by using interpreter-specific parallelism where and if it is
+supported. For example, Python programs can use all sort of subprocess
+utilities while shell programs have full access to shell job control.
+
+Restricting streams to text-only data type seems to be a more funcdamental
+limitation. Litrepl lifts this limitation by supporting text-only document
+formats. Both Latex and Markdown have means of including rich data but they
+don't encode these data directly. Instead they rely on side channels: file
+system or network for referring such objects. Litrepl shares both pros such as
+good representation in well-known version control systems, and cons, e.g.
+requirement to organize such a side-channel data transfer explicitly.
+
+Another controversial technical solution - to transfer the whole document at
+input and output - can have performance penalties. In our experience the system
+works sufficienlty good for few-thousand line documents, but for larger
+documents delays might become uncomfortable even on modern computers. Still we
+prefer to keep this interface to make editor integration as simple as possible:
+typical plugin can just pipe the whole document through the tool using few lines
+of code.
+
+In addition, we allow a more performant integrations which take responsibility
+for pre-parsing and piping only the relevant parts of the document. For these
+approach, Litrepl offers \verb|print-regexp| command printing the anchor regexp
+in a few common formats.
+% lnoai
+\end{comment}
+
+% lresult
+The technical decision to abstract interpreters using text streams comes with
+both advantages and disadvantages. A key advantage is simplicity. However, there
+are notable negative aspects. Firstly, there is no parallel evaluation at the
+communication level, meaning the interpreter is locked until it completes the
+evaluation of one snippet before proceeding to the next. Secondly, the
+transferable data type is restricted to text-only streams.
+
+We argue that the lack of parallel execution at the communication level can be
+mitigated using interpreter-specific parallelism, where supported. For instance,
+Python programs can utilize various subprocess utilities, while shell programs
+have full access to shell job control.
+
+The restriction to text-only data types presents a more fundamental limitation.
+Litrepl lifts this restriction by supporting text-only document formats. Both
+LaTeX and Markdown incorporate rich data without encoding it directly, instead
+relying on side channels, such as the file system or network, to refer to such
+objects. Consequently, Litrepl shares the advantages of good representation in
+well-known version control systems, and the disadvantages, such as the need to
+explicitly organize side-channel data transfer.
+
+Another controversial technical decision is transferring the entire document as
+input and output, which can negatively impact performance. Our experience shows
+that the system performs adequately for documents of a few thousand lines.
+However, larger documents may experience uncomfortable delays, even on modern
+computers. Despite this, we choose to maintain this interface because it
+simplifies editor integration. A typical plugin can pipe the whole document
+through the tool using just a few lines of code.
+
+Moreover, we support more performance-oriented integrations that handle
+pre-parsing and piping only relevant document parts. For these approaches,
+Litrepl offers the \verb|print-regexp| command, which outputs the anchor regexp
+in several common formats.
+% lnoresult
 

### 2025-01-01 paper/paper.tex
Update paper

@@
 \begin{abstract}
-Litrepl is a Python text processor for recognizing and evaluation code sections
-in Markdown or LaTeX documents. Litrepl is designed as a middleware aimed at
+Litrepl is a text processor tool for recognizing and evaluating code sections in
+Markdown or LaTeX documents. Litrepl is designed as middleware aimed at
 separating text editors from the programming language interpreter management
-logic thus contributing to the diversity of both. In this role, Litrepl can
-become a component of modular "UNIX Way" interactive development or typesetting
-environments.
+logic, thus contributing to the diversity of both.
 \end{abstract}

### 2025-01-01 python/litrepl/base.py
Add more Latex lignore options

@@ -293,8 +293,8 @@ class SymbolsLatex(Symbols):
     self.comcodeend = comcodeend_re
     self.comresultbegin = r"\%[ ]*l?result"
     self.comresultend = r"\%[ ]*l?noresult"
-    self.ignorebegin = r"\%[ ]*lignore"
-    self.ignoreend = r"\%[ ]*lnoignore"
+    self.ignorebegin = r"\%[ ]*lignore|\%[ ]*l?ignore[ ]*\%"
+    self.ignoreend = r"\%[ ]*lnoignore|\%[ ]*l?noignore[ ]*\%"
     self.inlinemarker = r"\\l[a-zA-Z0-9]*inline"
 
 OBR="{"

### 2025-01-01 paper/paper.tex
Update paper

@@
 \begin{abstract}
-Litrepl is a text processor tool for recognizing and evaluating code sections in
-Markdown or LaTeX documents. Litrepl is designed as middleware aimed at
-separating text editors from the programming language interpreter management
-logic, thus contributing to the diversity of both.
+
+Litrepl is a simple and lightweight text processing tool designed to recognize
+and evaluate code sections within Markdown or LaTeX documents. This
+functionality is useful for both batch document section evaluation and
+interactive coding within a text editor, provided a straightforward integration
+is established.
+
 \end{abstract}

### 2025-01-01 paper/paper.tex
Update paper

@@
 \newcommand{\ee}{\end{enumerate}}
+\newcommand{\Latex}{\LaTeX\ }
 
@@
 Litrepl is a simple and lightweight text processing tool designed to recognize
-and evaluate code sections within Markdown or LaTeX documents. This
+and evaluate code sections within Markdown or \Latex documents. This
 functionality is useful for both batch document section evaluation and
@@
 
-Concurrently, another concept of human-computer interaction called the
-Read-Evaluate-Print Loop (REPL) gained traction, notably within the LISP
-community and through the APL language designed for mathematical computations.
+Concurrently, a concept of human-computer interaction called the
+Read-Evaluate-Print Loop or ''REPL'' gained traction, notably within the LISP
+an APL communities.\cite{Spence1975apl, McCarthy1959recfun, Iverson1962apl}
+
 The combination of a command-line interface and a language interpreter enables
@@
 interpreter state. By maintaining human involvement in the loop, this approach
-facilitates human thought processes \cite{Spence1975apl, McCarthy1959recfun,
-Iverson1962apl}.
+is believed to facilitate human thought processes.
 
-A significant milestone in this area was the development of the IPython project,
-which later evolved into the Jupyter Project. The creators introduced a new
-document format called the Notebook, characterized by a series of logical
-sections of various types, including text and code, which could directly
-interact with programming language interpreters. This interactive communication,
-akin to REPL style programming, allows the creation of well-structured documents
-suitable for presentations and sharing. The concept underpinning these
-developments is termed Literate Computing, which includes goals of spanning a
-wide audience range, boosting reproducibility, and fostering collaboration. To
-achieve these objectives, several technical decisions were made, notably the
-introduction of bidirectional communication—between the computational core,
-known as the Jupyter Kernel, and the Notebook acting as a client, along with
-another layer of client-server communication between the Notebook web server and
-the user’s web browser \cite{Perez2007IPython, Granger2021litcomp,
-Kluyver2016jupnb}.
+A significant milestone in this area became the IPython terminal, which later
+evolved into the Jupyter Project. The creators introduced a new document format
+called the Notebook, characterized by a series of logical sections of various
+types, including text and code, which could directly interact with programming
+language interpreters. This interactive communication, akin to REPL style
+programming, allows the creation of well-structured documents suitable for
+presentations and sharing. The concept underpinning these developments is termed
+Literate Computing, which includes goals of spanning a wide audience range,
+boosting reproducibility, and fostering collaboration. To achieve these
+objectives, several technical decisions were made, notably the introduction of
+bidirectional communication—between the computational core, known as the Jupyter
+Kernel, and the Notebook acting as a client, along with another layer of
+client-server communication between the Notebook web server and the user’s web
+browser \cite{Perez2007IPython, Granger2021litcomp, Kluyver2016jupnb}.
 
-4. While we recognize the importance of various goals within the Literate
-Computing framework, we argue that reproducibility is paramount. Successfully
-addressing reproducibility would alone suffice to enhance communication over
-time among research project participants and significantly expand the audience.
-However, as it became clear (\cite{Dolstra2010, Vallet2022}), this challenge
-extends far beyond the scope of a human-computer interaction system, and even
-beyond the typical boundaries of software distribution management for a
-particular programming language. A comprehensive solution to the software
-deployment problem should operate at the entire operating system level.
+While we recognize the importance of these goals within the Literate Computing
+framework, we argue that reproducibility is paramount. Successfully addressing
+reproducibility would alone suffice to enhance communication over time among
+research project participants and significantly expand the audience.  However,
+as it became clear (\cite{Dolstra2010, Vallet2022}), this challenge extends far
+beyond the scope of a human-computer interaction system, and even beyond the
+typical boundaries of software distribution management for a particular
+programming language. A comprehensive solution to the software deployment
+problem should operate at the entire operating system level.
 
-5. Consequently, we propose focusing on transparency in human-computer
-interaction rather than on reproducibility. We present "Litrepl," a tool
-designed to integrate the REPL interactive programming style into existing
-editors in a straightforward yet potent manner.
+After understanding the scale and complexity of the reproducibility problem, we
+suggest to change the focus of human-computer interaction towards simplicity and
+transparency. We saw an opportunity to implement a tool named "Litrepl", which
+would offer the same familiar REPL-style editing, be compatible with existing
+text editors, thus maintaining familiar editing practices, and have a gradually
+smaller codebase.
 
-6. First, we utilize simple bidirectional text streams for inter-process
-communication with language interpreters to evaluate code. Second, we advocate
-for the reuse of existing text document formats. In both the Markdown and LaTeX
-formats that we have implemented, simplified parsers are utilized to distinguish
-code and result sections from the rest of the document. As of now, we support
-Python and Shell interpreter families, as well as a custom AI communication
-interpreter. Finally, we strive to leverage POSIX \cite{POSIX2024} operating
-system facilities as much as possible.
+We accomplish this through the following steps. First, we utilize simple
+bidirectional text streams for inter-process communication with language
+interpreters to evaluate code. Second, we advocate for the reuse of existing
+text document formats. In both the Markdown and \Latex evaluators that we have
+implemented, simplified parsers are utilized to distinguish code and result
+sections from the rest of the document. As of now, we support Python and Shell
+interpreter families, as well as a custom AI communication interpreter. Finally,
+we strive to leverage POSIX \cite{POSIX2024} operating system facilities as much
+as possible.
 
@@
 This document contains a Python code section and an empty result section marked
-with the corresponding Latex environment tags. To "execute" the document we
+with the corresponding \Latex environment tags. To "execute" the document we
 pipe it though the Litrepl processor as follows:
@@
   \end{minipage}
-  \caption{An illustrative grammar template for LaTeX documents where marker
+  \caption{An illustrative grammar template for \Latex documents where marker
   serves as a parameter configured via command-line arguments for each supported
@@
 Template grammars similar to the illustrative example in Figure
-\ref{fig:document-representation} are encoded for Markdown and LaTeX formats.
+\ref{fig:document-representation} are encoded for Markdown and \Latex formats.
 Before each run, Litrepl calls Lark\cite{Lark} to compile a customized parser
@@
 Litrepl lifts this restriction by supporting text-only document formats. Both
-LaTeX and Markdown incorporate rich data without encoding it directly, instead
+\Latex and Markdown incorporate rich data without encoding it directly, instead
 relying on side channels, such as the file system or network, to refer to such

### 2025-01-02 paper/paper.tex
Update paper

@@
-\documentclass[a4paper,12pt,twocolumn]{article}
+\documentclass[letterpaper,12pt,twocolumn]{article}
 
@@
 tangling, which generates machine-executable code. Over time, the concept has
-evolved, showing a trend towards simplification \cite{Knuth1984lp,
-Ramsey1994lps}.
+evolved, showing a trend towards simplification \textcite{Knuth1984lp},
+\textcite{Ramsey1994lps}.
 
 Concurrently, a concept of human-computer interaction called the
-Read-Evaluate-Print Loop or ''REPL'' gained traction, notably within the LISP
-an APL communities.\cite{Spence1975apl, McCarthy1959recfun, Iverson1962apl}
+Read-Evaluate-Print Loop or ''REPL'' gained traction, notably within the LISP an
+APL communities.\textcite{Spence1975apl, McCarthy1959recfun, Iverson1962apl}
 
@@
 client-server communication between the Notebook web server and the user’s web
-browser \cite{Perez2007IPython, Granger2021litcomp, Kluyver2016jupnb}.
+browser \textcite{Perez2007IPython, Granger2021litcomp, Kluyver2016jupnb}.
 
@@
 research project participants and significantly expand the audience.  However,
-as it became clear (\cite{Dolstra2010, Vallet2022}), this challenge extends far
-beyond the scope of a human-computer interaction system, and even beyond the
+as it became clear (\textcite{Dolstra2010, Vallet2022}), this challenge extends
+far beyond the scope of a human-computer interaction system, and even beyond the
 typical boundaries of software distribution management for a particular
@@
 interpreter families, as well as a custom AI communication interpreter. Finally,
-we strive to leverage POSIX \cite{POSIX2024} operating system facilities as much
-as possible.
+we strive to leverage POSIX \textcite{POSIX2024} operating system facilities as
+much as possible.
 
@@
 \ref{fig:document-representation} are encoded for Markdown and \Latex formats.
-Before each run, Litrepl calls Lark\cite{Lark} to compile a customized parser
-and uses it to access the sections.
+Before each run, Litrepl calls Lark\textcite{Lark} to compile a customized
+parser and uses it to access the sections.
 

### 2025-01-02 paper/paper.tex
Update paper

@@
 
-The concept of Literate Programming, initially proposed by Donald Knuth,
-revolves around explaining to human beings what we want a computer to do. This
-approach is embodied in tools like WEB, CWEB, and NOWEB, which focus on
-non-interactive, ahead-of-time compilation. The process includes two primary
-modes of operation: weaving, which creates human-readable documentation, and
-tangling, which generates machine-executable code. Over time, the concept has
-evolved, showing a trend towards simplification \textcite{Knuth1984lp},
-\textcite{Ramsey1994lps}.
+The concept of \textit{Literate Programming} was formulated by Donald Knuth,
+suggesting a shift in focus from writing code to explaining to human beings what
+we want a computer to do (\textcite{Knuth1984lp}). This approach is embodied in
+the WEB system and its derived family of tools, which translate combined code
+and narrative documents into human-readable documentation and machine-executable
+code. Over time, the concept has evolved, showing a trend towards simplification
+(\textcite{Ramsey1994lps}).
 
-Concurrently, a concept of human-computer interaction called the
-Read-Evaluate-Print Loop or ''REPL'' gained traction, notably within the LISP an
-APL communities.\textcite{Spence1975apl, McCarthy1959recfun, Iverson1962apl}
+Concurrently, a concept of human-computer interaction often called the
+\textit{Read-Evaluate-Print Loop} or ''REPL'' gained traction, notably within
+the LISP an APL communities (\textcite{Spence1975apl},
+\textcite{McCarthy1959recfun}, \textcite{Iverson1962apl}).
 
@@
 interpreter state. By maintaining human involvement in the loop, this approach
-is believed to facilitate human thought processes.
+is believed to facilitate human thought processes (\textcite{Granger2021litcomp}).
 
-A significant milestone in this area became the IPython terminal, which later
-evolved into the Jupyter Project. The creators introduced a new document format
-called the Notebook, characterized by a series of logical sections of various
-types, including text and code, which could directly interact with programming
-language interpreters. This interactive communication, akin to REPL style
-programming, allows the creation of well-structured documents suitable for
+A significant milestone in this area became the IPython interpreter
+(\textcite{Perez2007IPython}), which later evolved into the Jupyter Project. The
+creators introduced a new document format called the Jupyter Notebook
+(\textcite{Kluyver2016jupnb}), characterized by a series of logical sections of
+various types, including text and code, which could directly interact with
+programming language interpreters. This interactive communication, akin to REPL
+style programming, allows the creation of well-structured documents suitable for
 presentations and sharing. The concept underpinning these developments is termed
-Literate Computing, which includes goals of spanning a wide audience range,
-boosting reproducibility, and fostering collaboration. To achieve these
-objectives, several technical decisions were made, notably the introduction of
-bidirectional communication—between the computational core, known as the Jupyter
-Kernel, and the Notebook acting as a client, along with another layer of
-client-server communication between the Notebook web server and the user’s web
-browser \textcite{Perez2007IPython, Granger2021litcomp, Kluyver2016jupnb}.
+\textit{Literate Computing}\cite{Perez2015blog}, which includes goals of
+spanning a wide audience range, boosting reproducibility, and fostering
+collaboration. To achieve these objectives, several technical decisions were
+made, notably the introduction of bidirectional communication between the
+computational core, known as the Jupyter Kernel, and the Notebook web-based
+renderer, along with another layer of client-server communication between the
+renderer and the user’s web browser.
 
-While we recognize the importance of these goals within the Literate Computing
-framework, we argue that reproducibility is paramount. Successfully addressing
-reproducibility would alone suffice to enhance communication over time among
-research project participants and significantly expand the audience.  However,
-as it became clear (\textcite{Dolstra2010, Vallet2022}), this challenge extends
-far beyond the scope of a human-computer interaction system, and even beyond the
-typical boundaries of software distribution management for a particular
+While we recognize the importance of all goals within the Literate Computing
+framework, we argue that the goal of reproducibility is paramount. Successfully
+addressing reproducibility would alone suffice to enhance communication over the
+time and space separated researchers and significantly expand the audience.
+However, as it became clear (\textcite{Dolstra2010}), this challenge extends
+beyond the scope of a single human-computer interaction system, and even beyond
+the typical boundaries of software distribution management for a particular
 programming language. A comprehensive solution to the software deployment
@@
 
-After understanding the scale and complexity of the reproducibility problem, we
-suggest to change the focus of human-computer interaction towards simplicity and
-transparency. We saw an opportunity to implement a tool named "Litrepl", which
-would offer the same familiar REPL-style editing, be compatible with existing
-text editors, thus maintaining familiar editing practices, and have a gradually
-smaller codebase.
+Following \textcite{Vallet2022}, we suggest changing the focus of human-computer
+interaction towards simplicity and transparency. We saw an opportunity to
+implement a tool that would offer familiar REPL-style editing, be compatible
+with existing code editors and text formats, thus maintaining familiar editing
+practices, and have a gradually smaller codebase.
 
-We accomplish this through the following steps. First, we utilize simple
-bidirectional text streams for inter-process communication with language
-interpreters to evaluate code. Second, we advocate for the reuse of existing
-text document formats. In both the Markdown and \Latex evaluators that we have
-implemented, simplified parsers are utilized to distinguish code and result
-sections from the rest of the document. As of now, we support Python and Shell
-interpreter families, as well as a custom AI communication interpreter. Finally,
-we strive to leverage POSIX \textcite{POSIX2024} operating system facilities as
-much as possible.
+We introduce \textit{Litrepl}, a processor designed to achieve these goals
+through several means. First, we employ straightforward bidirectional text
+streams for inter-process communication with language interpreters to evaluate
+code.  Second, we advocate for the reuse of existing text document formats. In
+both the Markdown and \Latex evaluators that we have implemented, simplified
+parsers are utilized to distinguish code and result sections from the rest of
+the document.  As of now, we support Python and Shell interpreter families, as
+well as a custom large language model communication interpreter. Finally, we
+strive to leverage POSIX\cite{POSIX2024} operating system facilities as much as
+possible.
 

### 2025-01-02 paper/paper.tex
Update paper

@@
   \includesvg[inkscapelatex=false,width=0.8\textwidth]{pic}
-  \caption{Caption describing the image shown in the figure.}
+  \caption{Litrepl resource allocation diagram. Hash \textbf{A} is computed
+  based on the Litrepl working directory and the interpreter class. Hash
+  \textbf{B} is computed based on the contents of the code section.}
   \label{fig:resource-allocation}
@@
 metric, and has only two Python dependencies so far, at the cost of the
-dependency on the operating system intefaces for which we choose POSIX as a
-wide-spread openly available standard.
+dependency on the operating system interfaces for which we choose POSIX as a
+widespread openly available standard. Needless to say, we used Litrepl to
+evaluate and verify the examples presented in this document.
 

### 2025-01-02 paper/paper.tex
Update paper

@@
 
-The tool is implemented in Python in about 2K lines of code according to the LOC
+The tool is implemented in Python in under 2K lines of code according to the LOC
 metric, and has only two Python dependencies so far, at the cost of the
-dependency on the operating system interfaces for which we choose POSIX as a
-widespread openly available standard. Needless to say, we used Litrepl to
-evaluate and verify the examples presented in this document.
+dependency on the POSIX operating system interfaces. Needless to say, we used
+Litrepl to evaluate and verify the examples presented in this document.
 

### 2025-01-02 paper/paper.tex
Update paper

@@
 \usepackage{hyperref}
-\usepackage{fancyhdr}
+% \usepackage{fancyhdr}
 \usepackage{svg}
@@
 
-\fancypagestyle{plain}{
-  \fancyhf{}
-  \lhead{Sergei Mironov}
-  \rhead{Litrepl}
-  \rfoot{\thepage}
-}
+% \fancypagestyle{plain}{
+%   \fancyhf{}
+%   \lhead{Sergei Mironov}
+%   \rhead{Litrepl}
+%   \rfoot{\thepage}
+% }
 
@@
 we want a computer to do (\textcite{Knuth1984lp}). This approach is embodied in
-the WEB system and its derived family of tools, which translate combined code
+the WEB system and its descendant family of tools, which translate combined code
 and narrative documents into human-readable documentation and machine-executable
@@
 
-A significant milestone in this area became the IPython interpreter
+A significant milestone in this field was the IPython interpreter
 (\textcite{Perez2007IPython}), which later evolved into the Jupyter Project. The
@@
 renderer, along with another layer of client-server communication between the
-renderer and the user’s web browser.
+web-server and the user’s web browser.
 
@@
 framework, we argue that the goal of reproducibility is paramount. Successfully
-addressing reproducibility would alone suffice to enhance communication over the
-time and space separated researchers and significantly expand the audience.
-However, as it became clear (\textcite{Dolstra2010}), this challenge extends
-beyond the scope of a single human-computer interaction system, and even beyond
-the typical boundaries of software distribution management for a particular
-programming language. A comprehensive solution to the software deployment
-problem should operate at the entire operating system level.
+addressing reproducibility would alone suffice to enhance communication among
+time-separated and space-separated researchers and significantly expand the
+audience. However, as it became clear (\textcite{Dolstra2010}), this challenge
+extends beyond the scope of a single human-computer interaction system, and even
+beyond the typical boundaries of software distribution management for a
+particular programming language. A comprehensive solution to the software
+deployment problem should operate at the entire operating system level.
 
@@
 interaction towards simplicity and transparency. We saw an opportunity to
-implement a tool that would offer familiar REPL-style editing, be compatible
-with existing code editors and text formats, thus maintaining familiar editing
-practices, and have a gradually smaller codebase.
+implement a tool that would offer REPL-style editing, be compatible with
+existing code editors and text formats, thus maintaining familiar editing
+practices, and have a significantly smaller codebase.
 
-We introduce \textit{Litrepl}, a processor designed to achieve these goals
-through several means. First, we employ straightforward bidirectional text
-streams for inter-process communication with language interpreters to evaluate
-code.  Second, we advocate for the reuse of existing text document formats. In
-both the Markdown and \Latex evaluators that we have implemented, simplified
-parsers are utilized to distinguish code and result sections from the rest of
-the document.  As of now, we support Python and Shell interpreter families, as
-well as a custom large language model communication interpreter. Finally, we
-strive to leverage POSIX\cite{POSIX2024} operating system facilities as much as
-possible.
+We introduce \textit{Litrepl}, a text processor that employs the following
+approaches: first, utilizing straightforward bidirectional text streams for
+inter-process communication with language interpreters to evaluate code; second,
+advocating for the reuse of existing text document formats. In both the Markdown
+and \LaTeX\ evaluators we have implemented, simplified parsers are used to
+distinguish code and result sections from the rest of the document. As of now,
+we support Python and Shell interpreter families, as well as a custom large
+language model communication interpreter. Finally, we strive to leverage
+POSIX\cite{POSIX2024} operating system facilities as much as possible.
 
@@
 
-Litrepl is implemented as a command-line text processor utility. Its primary
-function is to take a text document as input, process it according to specified
-command line arguments and environment settings, and then output the resulting
-document through its standard output.
+Litrepl is implemented as a command-line text utility. Its primary function is
+to take a text document as input, process it according to specified command-line
+arguments and environment settings, and then output the resulting document
+through its standard output.
 

### 2025-01-02 paper/paper.tex
Update paper

@@
 
-Litrepl is a simple and lightweight text processing tool designed to recognize
-and evaluate code sections within Markdown or \Latex documents. This
-functionality is useful for both batch document section evaluation and
-interactive coding within a text editor, provided a straightforward integration
-is established.
+Litrepl is a lightweight text processing tool designed to recognize and evaluate
+code sections within Markdown or \Latex documents. This functionality is useful
+for both batch document section evaluation and interactive coding within a text
+editor, provided a straightforward integration is established.
 
@@
   \end{minipage}
-  \caption{Example output in a verbatim format displayed above two columns.}
+  \caption{Partial result section with a continuation tag.}
   \label{fig:partial-result}
@@
 \ref{fig:document-representation} are encoded for Markdown and \Latex formats.
-Before each run, Litrepl calls Lark\textcite{Lark} to compile a customized
+Before each run, Litrepl calls Lark (\textcite{Lark}) to compile a customized
 parser and uses it to access the sections.
@@
 Litrepl lifts this restriction by supporting text-only document formats. Both
-\Latex and Markdown incorporate rich data without encoding it directly, instead
-relying on side channels, such as the file system or network, to refer to such
-objects. Consequently, Litrepl shares the advantages of good representation in
-well-known version control systems, and the disadvantages, such as the need to
-explicitly organize side-channel data transfer.
+\Latex and Markdown incorporate non-text data without encoding it directly,
+instead relying on references and side channels, such as the file system or
+network resource identifiers. Consequently, Litrepl shares, for example, the
+benefits of human-readable representation in version control systems and the
+penalties, such as the need to explicitly organize side-channel data transfer.
 
@@
 
-Moreover, we support more performance-oriented integrations that handle
-pre-parsing and piping only relevant document parts. For these approaches,
-Litrepl offers the \verb|print-regexp| command, which outputs the anchor regexp
-in several common formats.
+A more performance-oriented integrations can make pre-parsing and pipe only
+relevant document parts. For these approaches, Litrepl offers the
+\verb|print-regexp| command, which outputs the anchor regexp in several common
+formats.
 % lnoresult

### 2025-01-04 env.sh
Remove pylightnix submodule

@@ -5,14 +5,11 @@ $CWD/sh:\
 $CWD/python/bin:\
 $PATH
 export LITREPL_ROOT="$CWD"
-export PYLIGHTNIX_ROOT="$CWD/_pylightnix"
 export PYTHONPATH=\
 $CWD/python:\
-$CWD/modules/pylightnix/src:\
 $PYTHONPATH
 export MYPYPATH=\
 $CWD/python:\
-$CWD/modules/pylightnix/src
 alias ipython="sh $CWD/ipython.sh"
 
 unset vim
@@ -20,7 +17,6 @@ unset VIMRUNTIME
 
 python3 -c 'import lark' || echo "Warning: 'lark' python library not found" >&2
 python3 -c 'import dataclasses' || echo "Warning: 'dataclasses' python library not found" >&2
-# python -c 'import pylightnix' || echo "Warning: 'lark' python library not found" >&2
 litrepl --version >/dev/null || echo "litrepl is not in PATH" >&2
 socat -V >/dev/null || echo "GNU socat not found" >&2
 

### 2025-01-04 python/litrepl/base.py
Remove underscores in auxdir file names

@@ -111,8 +111,8 @@ def pipenames(a:LitreplArgs, st:SType)->FileNames:
   the current directory. """
   auxdir=st2auxdir(a,st)
   return FileNames(auxdir,
-                   join(auxdir,"_in.pipe"),join(auxdir,"_out.pipe"),
-                   join(auxdir,"_pid.txt"),join(auxdir,"_ecode.txt"),join(auxdir,"_emsg.txt"))
+                   join(auxdir,"in.pipe"),join(auxdir,"out.pipe"),
+                   join(auxdir,"pid.txt"),join(auxdir,"ecode.txt"),join(auxdir,"emsg.txt"))
 
 def attach(fns:FileNames, st:SType|None=None)->Optional[Interpreter]:
   """ Attach to the interpreter associated with the given pipe filenames. """
@@ -158,7 +158,7 @@ def write_child_pid(pidf,pid):
 
 def start_(a:LitreplArgs, interpreter:str, i:Interpreter)->int:
   """ Starts the background Python interpreter. Kill an existing interpreter if
-  any. Creates files `_inp.pipe`, `_out.pipe`, `_pid.txt`, etc."""
+  any. Creates files `inp.pipe`, `out.pipe`, `pid.txt`, etc."""
   fns=i.fns
   makedirs(fns.wd, exist_ok=True)
   try:

### 2025-01-05 paper/paper.tex
Update paper

@@
 suggesting a shift in focus from writing code to explaining to human beings what
-we want a computer to do (\textcite{Knuth1984lp}). This approach is embodied in
-the WEB system and its descendant family of tools, which translate combined code
-and narrative documents into human-readable documentation and machine-executable
-code. Over time, the concept has evolved, showing a trend towards simplification
-(\textcite{Ramsey1994lps}).
+we want a computer to do. This approach is embodied in the WEB
+system\cite{Knuth1984lp} and its descendant family of tools, whose name refers
+to a text document format containing the "network" of code sections interleaved
+with text.
+
+The system could both render such text into human-readable documentation and
+compile machine-executable code. Over time, this concept has evolved, showing a
+trend towards simplification (\textcite{Ramsey1994lps}).
 
@@
 A significant milestone in this field was the IPython interpreter
-(\textcite{Perez2007IPython}), which later evolved into the Jupyter Project. The
+(\textcite{Perez2007IPython}), which later evolved into the Jupyter Project. Its
 creators introduced a new document format called the Jupyter Notebook
@@
 While we recognize the importance of all goals within the Literate Computing
-framework, we argue that the goal of reproducibility is paramount. Successfully
-addressing reproducibility would alone suffice to enhance communication among
+framework, we think that the goal of reproducibility is more important than
+others. Addressing it alone would suffice to enhance communication among
 time-separated and space-separated researchers and significantly expand the
@@
 particular programming language. A comprehensive solution to the software
-deployment problem should operate at the entire operating system level.
+deployment problem operates at the entire operating system level.
 
@@
 existing code editors and text formats, thus maintaining familiar editing
-practices, and have a significantly smaller codebase.
+practices, contain no important hidden state variables and have a significantly
+smaller codebase.
 
@@
 families, which we refer to as \textit{interpreter classes}. At the time of
-writing, Litrepl supports three such classes: \textbf{python}, \textbf{sh}, and
-\textbf{ai}. Using command line arguments, users can configure how to map code
-section labels to the correct class and specify which interpreter command to
-execute to start a session for each class.
+writing, Litrepl supports three such classes: \texttt{python}, \texttt{sh}, and
+\texttt{ai}. Using these names in command line arguments, users can configure
+how to map code section labels to the correct class and specify which
+interpreter command to execute to start a session for each class. For example,
+to select a Bourne-Again Shell interpreter as \verb|sh|, we add the
+\verb|--sh-interpreter=/usr/bin/bash| argument assuming that this binary is
+present in the system.
+%lnoresult
 
@@
   \end{minipage}
-  \caption{Partial result section with a continuation tag.}
+  \caption{Example partial result section that ends with a continuation tag,
+  where \texttt{LR:} is a distinguishable prefix, and the rest is the filename
+  storing a response of the interpreter to the section being evaluated.}
   \label{fig:partial-result}
@@
 
-Litrepl provides \textbf{start}, \textbf{stop}, and \textbf{restart} commands to
-control background sessions. The \textbf{interrupt} command sends an
-interruption signal to the interpreter. Finally, the \textbf{repl} command
-establishes direct communication with the interpreter, allowing manual
-inspection of its state.
+As already mentioned, most of the resources are exposed to user via the
+filesystem. The only added hidden state in the entire system resides in the
+reader process, in a small finite state machine that handles special ASCII
+characters used by some applications for text animation.
+
+Litrepl provides \textbf{start}, \textbf{stop}, \textbf{restart} and
+\textbf{status} commands to control background sessions, so, for example
+
+\begin{verbatim}
+$ litrepl restart python
+\end{verbatim}
+
+stops the Python interpreter if it was running and starts a new instance of it.
+The \textbf{interrupt} command sends an interruption signal to the interpreter.
+Finally, the \textbf{repl} command establishes direct communication with the
+interpreter, allowing manual inspection of its state. The \textbf{help} command
+prints the detailed description of each command and the configuration
+arguments available.
 %lnoresult
@@
 
-Evaluation results are encoded back into the result sections, and the entire
+Evaluation results are written back into the result sections, and the entire
 document is printed. At this stage, certain conditions can be optionally
@@
 both advantages and disadvantages. A key advantage is simplicity. However, there
-are notable negative aspects. Firstly, there is no parallel evaluation at the
+are notable negative aspects. First, there is no parallel evaluation at the
 communication level, meaning the interpreter is locked until it completes the
-evaluation of one snippet before proceeding to the next. Secondly, the
+evaluation of one snippet before proceeding to the next. Second, the
 transferable data type is restricted to text-only streams.

### 2025-01-05 paper/paper.tex
Update paper

@@
 existing code editors and text formats, thus maintaining familiar editing
-practices, contain no important hidden state variables and have a significantly
+practices, contain only a few hidden state variables, and have a significantly
 smaller codebase.
@@
 present in the system.
-%lnoresult
 

### 2025-01-05 paper/paper.tex
Update paper

@@
 During evaluation, Litrepl spawns a response reader process with a soft lock,
-active until the interpreter completes and responds to an echo probe. If the
-response exceeds the configured duration, Litrepl outputs a partial result tag,
-which is recognized and reevaluated in subsequent runs.  Figure
+active until the interpreter completes and responds to an echo probe. The state
+machine that operates the probe is the only added hidden state in the entire
+system.
+
+If the response exceeds the configured duration, Litrepl outputs a partial
+result tag, which is recognized and reevaluated in subsequent runs.  Figure
 \ref{fig:partial-result} shows an example partial result section.
-
-As already mentioned, most of the resources are exposed to user via the
-filesystem. The only added hidden state in the entire system resides in the
-reader process, in a small finite state machine that handles special ASCII
-characters used by some applications for text animation.
 

### 2025-01-06 paper/tex/paper.tex
Update paper.tex

--- 2025-01-05 paper/paper.tex
@@
 \usepackage[font=small,skip=0pt]{caption}
-
 \usepackage{biblatex}
 \addbibresource{paper.bib}
+\usepackage[para,bottom]{footmisc}
 
@@
 
-We introduce \textit{Litrepl}, a text processor that employs the following
-approaches: first, utilizing straightforward bidirectional text streams for
-inter-process communication with language interpreters to evaluate code; second,
-advocating for the reuse of existing text document formats. In both the Markdown
-and \LaTeX\ evaluators we have implemented, simplified parsers are used to
-distinguish code and result sections from the rest of the document. As of now,
-we support Python and Shell interpreter families, as well as a custom large
-language model communication interpreter. Finally, we strive to leverage
-POSIX\cite{POSIX2024} operating system facilities as much as possible.
+We introduce
+\textit{Litrepl}\footnote{\url{https://github.com/sergei-mironov/litrepl}},
+a text processor that employs the following approaches: first, utilizing
+straightforward bidirectional text streams for inter-process communication with
+language interpreters to evaluate code; second, advocating for the reuse of
+existing text document formats. In both the Markdown and \Latex evaluators we
+have implemented, simplified parsers are used to distinguish code and result
+sections from the rest of the document.  As of now, we support Python and Shell
+interpreter families, as well as a custom large language model communication
+interpreter. Finally, we strive to leverage POSIX\cite{POSIX2024} operating
+system facilities as much as possible.
 

### 2025-01-06 paper/tex/paper.tex
Update paper

@@
 for both batch document section evaluation and interactive coding within a text
-editor, provided a straightforward integration is established.
+editor, provided a straightforward integration is established. Inspired by
+Project Jupyter, it aims to facilitate the creation of research documents. In
+the light of recent developments in software deployment, however, we have
+shifted our focus from ad-hoc reproducibility to enhancing transparency in
+communication with programming language interpreters, by either eliminating or
+clearly exposing mutable states within the communication process.
 

### 2025-01-07 paper/tex/paper.tex
Update papers

--- 2025-01-06 paper/tex/paper.tex
@@
 for both batch document section evaluation and interactive coding within a text
-editor, provided a straightforward integration is established.
+editor, provided a straightforward integration is established. Inspired by
+Project Jupyter, Litrepl aims to facilitate the creation of research documents.
+In the light of recent developments in software deployment, however, we have
+shifted our focus from informal reproducibility to enhancing transparency in
+communication with programming language interpreters, by either eliminating or
+clearly exposing mutable states within the communication process.
 

### 2025-01-18 paper/tex/paper.tex
Update papers

@@
 \begin{sh}
+litrepl restart python
 echo '\begin{verbatim}'
@@
   ... some output ...
-  [LR:/tmp/litrepl/python/partial_7b81e1e.txt]
+  [LR:/tmp/litrepl/python/partial_c335adc.txt]
   \end{result}
@@
 
-\begin{comment}
-% lai
-
-Lets write the "Discussion" section. of the paper. Please check the grammar of
-the following points and make a linked narrative out of them. Please use Latex
-markup.
-
----
-
-The technical decision to abstract interpreters using text streams has its
-positive and negative sides. While simplicity is the obvious bonus, the negatvie
-consequences can be groouped into (1) No parallel evaluation at the
-communication level. Th interpreter gets locked until it finishes evaluation of
-one snippet before it can proceed to another one. (2) Transferrable data type is
-limited to text-only streams.
-
-We argue that the lack of parallel execution at the communication level can be
-compensated by using interpreter-specific parallelism where and if it is
-supported. For example, Python programs can use all sort of subprocess
-utilities while shell programs have full access to shell job control.
-
-Restricting streams to text-only data type seems to be a more funcdamental
-limitation. Litrepl lifts this limitation by supporting text-only document
-formats. Both Latex and Markdown have means of including rich data but they
-don't encode these data directly. Instead they rely on side channels: file
-system or network for referring such objects. Litrepl shares both pros such as
-good representation in well-known version control systems, and cons, e.g.
-requirement to organize such a side-channel data transfer explicitly.
-
-Another controversial technical solution - to transfer the whole document at
-input and output - can have performance penalties. In our experience the system
-works sufficienlty good for few-thousand line documents, but for larger
-documents delays might become uncomfortable even on modern computers. Still we
-prefer to keep this interface to make editor integration as simple as possible:
-typical plugin can just pipe the whole document through the tool using few lines
-of code.
-
-In addition, we allow a more performant integrations which take responsibility
-for pre-parsing and piping only the relevant parts of the document. For these
-approach, Litrepl offers \verb|print-regexp| command printing the anchor regexp
-in a few common formats.
-% lnoai
-\end{comment}
-
-% lresult
 The technical decision to abstract interpreters using text streams comes with
@@
 formats.
-% lnoresult
 
